---
title: "FUNWAVE-TVD beach_2d_radiation diagnostic plots"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M UTC', tz = 'UTC')`"
output:
  html_document:
    toc: true
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(terra)

# The report is knitted from report/, while results/ is at repository level.
repo_dir <- normalizePath("..", mustWork = TRUE)
data_dir <- if (dir.exists(file.path(repo_dir, "results"))) {
  file.path(repo_dir, "results")
} else {
  file.path(repo_dir, "example_results")
}

dx <- 2
dy <- 2
jet <- grDevices::colorRampPalette(c("#00007F", "blue", "#00FFFF", "yellow", "red", "#7F0000"))(128)

available_numbers <- sub("^eta_", "", list.files(data_dir, pattern = "^eta_[0-9]{5}$"))
stopifnot(length(available_numbers) > 0)
instant_number <- tail(sort(available_numbers), 1)

field_path <- function(name, number = instant_number) {
  file.path(data_dir, paste0(name, "_", number))
}

read_field <- function(name, number = instant_number, required = TRUE) {
  path <- field_path(name, number)
  if (!file.exists(path)) {
    if (required) stop("Required FUNWAVE field is absent: ", basename(path))
    return(NULL)
  }
  as.matrix(utils::read.table(path, header = FALSE))
}

# FUNWAVE ASCII arrays have row 1 at y = 0. terra stores row 1 at ymax,
# so reverse rows and flatten row-by-row to retain the MATLAB pcolor orientation.
as_funwave_rast <- function(z, dx = 2, dy = 2) {
  r <- terra::rast(nrows = nrow(z), ncols = ncol(z),
                   xmin = 0, xmax = ncol(z) * dx,
                   ymin = 0, ymax = nrow(z) * dy)
  terra::values(r) <- as.vector(t(z[nrow(z):1, , drop = FALSE]))
  r
}

plot_field <- function(z, main, units, xlim = c(300, 500), ylim = c(0, 90)) {
  r <- as_funwave_rast(z, dx, dy)
  terra::plot(r, col = jet, main = main, xlim = xlim, ylim = ylim,
              axes = TRUE, plg = list(title = units, cex = 0.8),
              xlab = "x (m)", ylab = "y (m)")
}

eta <- read_field("eta", instant_number)
mask <- read_field("mask", required = FALSE)
brk_x <- read_field("BrkSrcX", required = FALSE)
frc_x <- read_field("FrcInsX", required = FALSE)
```

## Provenance and input data

This is an R translation of the MATLAB files placed in `reference/`:
`plot_frc_brk.m`, `plot_balance.m` and `plot_vertical_profile.m`, from the
FUNWAVE-TVD `beach_2d_radiation` example. It uses base R for plotting and
`terra` to preserve the spatial-grid orientation.

The report reads the latest workflow's `results/` directory when available;
otherwise it uses the included `example_results/` files. The example files are
the 30-second GitHub Actions test output and are not intended as a physical
forecast. The workflow now runs a 300-second diagnostic case, which passes the
180-second steady-state threshold and generates the averaged radiation terms.

```{r run-summary}
meta_path <- file.path(data_dir, "run-metadata.json")
if (file.exists(meta_path)) {
  cat(paste(readLines(meta_path), collapse = "\n"))
}
```

## Instantaneous elevation, breaking and friction

This reproduces the three-panel layout in `plot_frc_brk.m`. The source MATLAB
example uses output number 21; this report detects and uses the latest available
instantaneous output (`r instant_number`).

```{r instantaneous-plots, fig.width=13, fig.height=4.8}
old_par <- par(no.readonly = TRUE)
on.exit(par(old_par), add = TRUE)
layout(matrix(1:3, nrow = 1))
plot_field(eta, "instantaneous eta", "eta (m)")

if (is.null(brk_x)) {
  plot.new(); title("breaking stress in x")
  text(0.5, 0.5, "BrkSrcX output is not available")
} else {
  plot_field(brk_x, "instantaneous breaking stress in x", "(h + eta) Rbx (m2/s2)")
}

if (is.null(frc_x)) {
  plot.new(); title("friction stress in x")
  text(0.5, 0.5, "FrcInsX output is not available")
} else {
  plot_field(frc_x, "instantaneous friction stress in x", "-Cd uU (m2/s2)")
}
```

## Averaged momentum-balance terms

`plot_balance.m` and the averaged lower panel in `plot_frc_brk.m` require
`PgrdX`, `FRCX`, `DxSxx`, `DySxy`, `DxUUH`, `DyUVH`, `PgrdY`, `FRCY`,
`DySyy`, `DxSxy`, `DyVVH`, `DxUVH` and `BrkDissX`. The compact test case does
not request these fields, so the section activates automatically only when all
required outputs are present.

```{r balance, fig.width=10, fig.height=7}
required_x <- c("PgrdX", "FRCX", "DxSxx", "DySxy", "DxUUH", "DyUVH")
required_y <- c("PgrdY", "FRCY", "DySyy", "DxSxy", "DyVVH", "DxUVH")
balance_numbers <- sub("^PgrdX_", "", list.files(data_dir, pattern = "^PgrdX_[0-9]{5}$"))
balance_number <- if (length(balance_numbers)) tail(sort(balance_numbers), 1) else NA_character_
available_balance <- !is.na(balance_number) && all(vapply(c(required_x, required_y), function(nm)
  file.exists(field_path(nm, balance_number)), logical(1)))

if (!available_balance) {
  cat("Averaged momentum-balance fields were not written by this run.\n")
} else {
  ny <- 25
  x_terms <- lapply(required_x, read_field, number = balance_number)
  names(x_terms) <- required_x
  y_terms <- lapply(required_y, read_field, number = balance_number)
  names(y_terms) <- required_y
  old_par <- par(no.readonly = TRUE); on.exit(par(old_par), add = TRUE)
  par(mfrow = c(2, 1), mar = c(4, 4, 2, 1))
  matplot(vapply(x_terms, function(z) z[ny, ], numeric(ncol(eta))), type = "l",
          lty = c(1, 1, 2, 1, 1, 2), col = c("red", "black", "blue", "blue", "cyan4", "cyan4"),
          xlab = "grid point", ylab = "m2/s2")
  legend("topleft", legend = names(x_terms), lty = c(1, 1, 2, 1, 1, 2),
         col = c("red", "black", "blue", "blue", "cyan4", "cyan4"), bty = "n")
  matplot(vapply(y_terms, function(z) z[ny, ], numeric(ncol(eta))), type = "l",
          lty = c(1, 1, 2, 1, 2, 1), col = c("red", "black", "blue", "blue", "cyan4", "cyan4"),
          xlab = "grid point", ylab = "m2/s2")
  legend("topleft", legend = names(y_terms), lty = c(1, 1, 2, 1, 2, 1),
         col = c("red", "black", "blue", "blue", "cyan4", "cyan4"), bty = "n")
}
```

## Vertical velocity reconstruction

The MATLAB `plot_vertical_profile.m` reconstructs vertical profiles using
`dep.out`, `Ax`, `Ay`, `Bx` and `By`. It performs the calculation but does not
actually issue a plotting command. The same reconstruction can be made in R
when these diagnostic fields are enabled in FUNWAVE; they are not in the
compact test output.

```{r vertical-profile, fig.width=7, fig.height=5}
vertical_fields <- c("dep.out", "Ax", "Ay", "Bx", "By", "u", "v", "eta")
has_vertical <- file.exists(file.path(data_dir, "dep.out")) &&
  all(vapply(vertical_fields[-1], function(nm) file.exists(field_path(nm)), logical(1)))

if (!has_vertical) {
  cat("Vertical-profile diagnostics were not written by this run.\n")
} else {
  dep <- as.matrix(utils::read.table(file.path(data_dir, "dep.out"), header = FALSE))
  ax <- read_field("Ax"); ay <- read_field("Ay")
  bx <- read_field("Bx"); by <- read_field("By")
  u <- read_field("u"); v <- read_field("v")
  z_levels <- seq(0, -1, length.out = 10)
  iy <- ceiling(nrow(dep) / 2); ix <- ceiling(ncol(dep) / 2)
  z <- -dep[iy, ix] * z_levels
  za <- -0.5528 * dep[iy, ix] + 0.4472 * eta[iy, ix]
  U <- u[iy, ix] + (za - z) * ax[iy, ix] + 0.5 * (za^2 - z^2) * bx[iy, ix]
  V <- v[iy, ix] + (za - z) * ay[iy, ix] + 0.5 * (za^2 - z^2) * by[iy, ix]
  matplot(cbind(U, V), z, type = "l", lty = 1, col = c("#2166AC", "#B2182B"),
          xlab = "velocity (m/s)", ylab = "z (m)")
  legend("bottomright", c("U", "V"), col = c("#2166AC", "#B2182B"), lty = 1, bty = "n")
}
```
