---
title: "FUNWAVE-TVD daily test forecast"
date: "`r format(Sys.time(), '%Y-%m-%d %H:%M UTC', tz = 'UTC')`"
output:
  html_document:
    self_contained: true
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(jsonlite)
library(ggplot2)

# rmarkdown renders this file with report/ as its working directory.
# Resolve the repository-level results directory rather than assuming cwd.
results_dir <- normalizePath("../results", mustWork = TRUE)
metadata <- fromJSON(file.path(results_dir, "run-metadata.json"))
eta_files <- list.files(results_dir, pattern = "^eta_", full.names = TRUE)
stopifnot(length(eta_files) > 0)
eta_file <- tail(sort(eta_files), 1)
eta <- as.matrix(read.table(eta_file, header = FALSE))
```

This page is generated automatically after the daily FUNWAVE-TVD test. Full
model files are retained as a GitHub Actions artifact for seven days; only this
small HTML report is published.

## Simulation log

```{r metadata}
knitr::kable(data.frame(
  Item = c("Model", "Test case", "Started (UTC)", "Finished (UTC)", "MPI processes", "Files written"),
  Value = c(metadata$model, metadata$test_case, metadata$started_at_utc,
            metadata$finished_at_utc, metadata$mpi_processes,
            paste(metadata$result_files, collapse = ", "))
))
```

## Final free-surface elevation

The plot reads the final `eta_*` field written by FUNWAVE-TVD. This example
uses a deliberately short 30-second, two-process test case; replace the case
preparation script when operational forcing is available.

```{r eta-plot, fig.width=9, fig.height=4.2}
eta_df <- expand.grid(x = seq_len(ncol(eta)), y = seq_len(nrow(eta)))
eta_df$eta_m <- as.vector(eta)

ggplot(eta_df, aes(x, y, fill = eta_m)) +
  geom_raster() +
  coord_equal(expand = FALSE) +
  scale_fill_gradient2(low = "#2166ac", mid = "white", high = "#b2182b",
                       midpoint = 0, name = "eta (m)") +
  labs(x = "Grid x", y = "Grid y") +
  theme_minimal(base_size = 12)
```
