---
title: "Port Fairy near-real-time FUNWAVE test"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 6)

case_dir <- normalizePath("../port_fairy/output", mustWork = TRUE)
results_dir <- file.path(case_dir, "results")
forcing <- read.csv(file.path(case_dir, "latest_buoy_forcing.csv"),
                    check.names = FALSE)
grid <- read.csv(file.path(case_dir, "grid_metadata.csv"), check.names = FALSE)

read_field <- function(prefix, directory = results_dir) {
  candidates <- list.files(directory, pattern = paste0("^", prefix, "_[0-9]{5}$"),
                           full.names = TRUE)
  if (!length(candidates)) return(NULL)
  number <- as.integer(sub(paste0("^", prefix, "_"), "", basename(candidates)))
  as.matrix(utils::read.table(candidates[which.max(number)], header = FALSE))
}

eta <- read_field("eta")
hmax <- read_field("Hmax")
waveheight <- read_field("WaveHeight")
```

## Latest forcing

This five-minute test uses a TMA/JONSWAP-style irregular wave source derived
from the newest usable IMOS Spotter observation. It is a model demonstration,
not a validated harbour forecast.

```{r forcing}
knitr::kable(forcing, digits = 2)
```

The positive model x-axis has a true-north bearing of
`r round(grid$model_x_bearing_deg_true, 1)` degrees and runs from the buoy-side
source edge toward the coast. `ThetaPeak` is therefore already expressed in
the rotated FUNWAVE coordinate system.

## Latest free-surface elevation

```{r eta, fig.cap='Latest modelled water-surface elevation. x = 0 is the buoy-side source edge; x increases toward the coast.'}
if (is.null(eta)) {
  plot.new(); text(0.5, 0.5, "No eta output was produced.")
} else {
  x <- (seq_len(nrow(eta)) - 1) * grid$dx_m
  y <- (seq_len(ncol(eta)) - 1) * grid$dy_m
  image(x, y, eta, asp = 1, col = hcl.colors(40, "Blue-Red 3"),
        xlab = "Model x: buoy-side edge to coast (m)",
        ylab = "Rotated model y (m)")
  contour(x, y, eta, add = TRUE, drawlabels = FALSE, col = "grey20")
}
```

## Maximum wave diagnostic

```{r hmax, fig.cap='Maximum-wave or wave-height diagnostic available at the end of the 5-minute test.'}
z <- if (!is.null(waveheight)) waveheight else hmax
if (is.null(z)) {
  plot.new(); text(0.5, 0.5, "No WaveHeight/Hmax output was produced.")
} else {
  x <- (seq_len(nrow(z)) - 1) * grid$dx_m
  y <- (seq_len(ncol(z)) - 1) * grid$dy_m
  image(x, y, z, asp = 1, col = hcl.colors(40, "YlOrRd", rev = TRUE),
        xlab = "Model x: buoy-side edge to coast (m)",
        ylab = "Rotated model y (m)")
}
```

## Provenance

```{r provenance}
knitr::kable(grid)
```
