---
title: "Port Fairy near-real-time FUNWAVE test"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 11, fig.height = 6)

suppressPackageStartupMessages({
  library(terra)
  library(leaflet)
})

case_dir <- normalizePath("../port_fairy/output", mustWork = TRUE)
# The normal forecast writes support files under site/assets. Historical runs
# set PORT_FAIRY_SITE_DIR to an isolated report folder so their maps/assets do
# not overwrite the live report's equivalents.
report_site_dir <- normalizePath(
  file.path("..", Sys.getenv("PORT_FAIRY_SITE_DIR", unset = "site")),
  mustWork = FALSE
)
results_dir <- file.path(case_dir, "results")
forcing <- read.csv(file.path(case_dir, "latest_buoy_forcing.csv"),
                    check.names = FALSE)
grid <- read.csv(file.path(case_dir, "grid_metadata.csv"), check.names = FALSE)
if (nrow(forcing) != 1 || nrow(grid) != 1) {
  stop("Expected one row in forcing and grid metadata files.")
}
if (!is.finite(grid$dx_m[1]) || !is.finite(grid$dy_m[1]) ||
    !is.finite(grid$total_time_s[1]) || !is.finite(grid$plot_intv_s[1]) ||
    grid$dx_m[1] <= 0 || grid$dy_m[1] <= 0 ||
    grid$total_time_s[1] <= 0 || grid$plot_intv_s[1] <= 0) {
  stop("Grid metadata must contain positive dx_m, dy_m, total_time_s and plot_intv_s values.")
}
metres_label <- function(x) {
  # Strip decimal zeroes, not meaningful integer zeroes: the previous helper
  # turned 200 into "2", which mislabelled the Umean/Vmean window as 2 s.
  out <- formatC(as.numeric(x), format = "f", digits = 6)
  out <- sub("0+$", "", out)
  sub("\\.$", "", out)
}
public_output_dx_m <- if ("public_output_dx_m" %in% names(grid) &&
                          is.finite(grid$public_output_dx_m[1])) {
  grid$public_output_dx_m[1]
} else {
  max(10, grid$dx_m[1])
}
public_output_dy_m <- if ("public_output_dy_m" %in% names(grid) &&
                          is.finite(grid$public_output_dy_m[1])) {
  grid$public_output_dy_m[1]
} else {
  public_output_dx_m
}
public_output_factor_x <- public_output_dx_m / grid$dx_m[1]
public_output_factor_y <- public_output_dy_m / grid$dy_m[1]
if (public_output_factor_x < 1 || public_output_factor_y < 1 ||
    abs(public_output_factor_x - round(public_output_factor_x)) > 1e-8 ||
    abs(public_output_factor_y - round(public_output_factor_y)) > 1e-8) {
  stop("The requested public-output x/y resolutions must be integer multiples of the model x/y resolutions.")
}
public_output_factor_x <- as.integer(round(public_output_factor_x))
public_output_factor_y <- as.integer(round(public_output_factor_y))
grid_tag <- if ("grid_file_tag" %in% names(grid) &&
                nzchar(as.character(grid$grid_file_tag[1]))) {
  as.character(grid$grid_file_tag[1])
} else if (abs(grid$dx_m[1] - grid$dy_m[1]) < 1e-8) {
  paste0(metres_label(grid$dx_m[1]), "m")
} else {
  paste0(metres_label(grid$dx_m[1]), "m_x_", metres_label(grid$dy_m[1]), "m_y")
}
public_grid_tag <- paste0(metres_label(public_output_dx_m), "m x ",
                          metres_label(public_output_dy_m), "m")
simulation_minutes <- grid$total_time_s[1] / 60
final_sixth_start_s <- (5 / 6) * grid$total_time_s[1]
elapsed_clock <- function(elapsed_s) {
  sprintf("%02.0f:%02.0f", floor(elapsed_s / 60), round(elapsed_s %% 60))
}
simulation_label <- paste0(sprintf("%.0f", simulation_minutes), "-minute")
final_sixth_start_label <- elapsed_clock(final_sixth_start_s)
simulation_end_label <- elapsed_clock(grid$total_time_s[1])
eta_display_limits <- c(-5, 5)
eta_max_display_limits <- c(0, 5)
hsig_display_limits <- c(0, 6)
ygnbu_palette <- RColorBrewer::brewer.pal(9, "YlGnBu")
# Station 1 is the buoy/offshore end. Use the perceptually ordered Magma
# sequence, from black-purple offshore to pale yellow at the beach end.
# These fixed Magma samples avoid adding another R package.
station_colours <- c("#000004", "#51127C", "#B5367A", "#FB8861", "#FCFDBF")
eta_animation_palette <- grDevices::colorRampPalette(ygnbu_palette)(100)

pretty_field_name <- function(x) {
  x <- gsub("_", " ", x)
  x <- tools::toTitleCase(x)
  x <- gsub("\\bFunwave\\b", "FUNWAVE", x)
  x <- gsub("\\bHs\\b", "Hs", x)
  x <- gsub("\\bTp\\b", "Tp", x)
  x <- gsub("\\bMglob\\b", "Mglob", x)
  x <- gsub("\\bNglob\\b", "Nglob", x)
  x <- gsub("\\bDx\\b", "DX", x)
  x <- gsub("\\bDy\\b", "DY", x)
  x <- gsub("\\bWk\\b", "WK", x)
  x
}

transpose_summary <- function(x, digits = 2) {
  x <- as.data.frame(x, check.names = FALSE, stringsAsFactors = FALSE)
  if (nrow(x) != 1) stop("transpose_summary() expects a one-row data frame.")
  data.frame(
    Item = pretty_field_name(names(x)),
    Value = vapply(x[1, ], function(value) {
      if (is.numeric(value)) format(value, digits = digits, trim = TRUE, scientific = FALSE) else as.character(value)
    }, character(1)),
    check.names = FALSE
  )
}

report_table <- function(x, digits = 2, summary = NULL) {
  table_html <- paste(
    knitr::kable(x, format = "html", digits = digits, escape = TRUE,
                 table.attr = 'class="pf-report-table"'),
    collapse = "\n"
  )
  if (is.null(summary)) {
    summary <- paste0("Show ", nrow(as.data.frame(x)), "-row table")
  }
  knitr::asis_output(paste0(
    '<details class="pf-report-table-details"><summary>',
    htmltools::htmlEscape(summary),
    '</summary>', table_html, '</details>'
  ))
}

# The buoys report UTC. Show the animation using Port Fairy's local civil time
# while retaining the model's elapsed time from that observation.
port_fairy_tz <- "Australia/Melbourne"
buoy_time_utc <- as.POSIXct(
  sub(" UTC$", "", forcing$time_utc[1]),
  format = "%Y-%m-%d %H:%M:%S", tz = "UTC"
)
if (is.na(buoy_time_utc)) stop("Could not parse the buoy observation time.")

# FUNWAVE writes text grids as Nglob rows (model y) by Mglob columns (model x).
# R's image() expects z[x, y], so transpose every model output exactly once.
as_model_xy <- function(z, label = "model field") {
  z <- as.matrix(z)
  if (all(dim(z) == c(grid$Nglob, grid$Mglob))) return(t(z))
  if (all(dim(z) == c(grid$Mglob, grid$Nglob))) return(z)
  stop(sprintf("%s has %.0f x %.0f values; expected %.0f x %.0f (y x x).",
               label, nrow(z), ncol(z), grid$Nglob, grid$Mglob))
}

read_text_model <- function(path, label = basename(path)) {
  as_model_xy(utils::read.table(path, header = FALSE), label)
}

read_field <- function(prefix, directory = results_dir) {
  candidates <- list.files(directory, pattern = paste0("^", prefix, "_[0-9]{5}$"),
                           full.names = TRUE)
  if (!length(candidates)) return(NULL)
  number <- as.integer(sub(paste0("^", prefix, "_"), "", basename(candidates)))
  read_text_model(candidates[which.max(number)], prefix)
}


read_all_fields <- function(prefix, directory = results_dir) {
  candidates <- list.files(directory, pattern = paste0("^", prefix, "_[0-9]{5}$"),
                           full.names = TRUE)
  if (!length(candidates)) return(list())
  number <- as.integer(sub(paste0("^", prefix, "_"), "", basename(candidates)))
  candidates <- candidates[order(number)]
  lapply(candidates, read_text_model, label = prefix)
}

maximum_model_field <- function(fields) {
  if (!length(fields)) return(NULL)
  out <- fields[[1]]
  if (length(fields) > 1) {
    for (k in 2:length(fields)) out <- pmax(out, fields[[k]], na.rm = TRUE)
  }
  out[!is.finite(out)] <- NA_real_
  out
}

read_static_model <- function(name) {
  path <- file.path(results_dir, name)
  if (!file.exists(path)) return(NULL)
  read_text_model(path, name)
}

model_axes <- function(z) {
  list(x = (seq_len(nrow(z)) - 1) * grid$dx_m,
       y = (seq_len(ncol(z)) - 1) * grid$dy_m)
}

plot_model <- function(z, main, col, zlim = range(z, finite = TRUE), ...) {
  a <- model_axes(z)
  image(a$x, a$y, z, asp = 1, col = col, zlim = zlim,
        xlab = "Model x: source edge to coast (m)",
        ylab = "Model y (m)", main = main, ...)
  invisible(a)
}

model_to_rotated_raster <- function(z, label = "model field") {
  z <- as.matrix(z)
  expected <- c(grid$Mglob, grid$Nglob)
  if (!identical(dim(z), expected)) {
    stop(sprintf("%s has %.0f x %.0f values; expected Mglob x Nglob = %.0f x %.0f.",
                 label, nrow(z), ncol(z), expected[1], expected[2]))
  }

  # z is [model x, model y]. Convert it back to terra's native
  # [north-to-south row, west-to-east column] layout using the exact inverse
  # of the transforms in setup_port_fairy_funwave.R.
  gis_rows <- t(z)
  if (source_is_low_x) {
    gis_rows <- gis_rows[nrow(gis_rows):1, , drop = FALSE]
  } else {
    gis_rows <- gis_rows[, ncol(gis_rows):1, drop = FALSE]
  }

  out <- rast(depth_raster)
  values(out) <- as.vector(t(gis_rows))
  out
}

# Convert an Omerc raster back to model [x, y] order. This is the inverse of
# model_to_rotated_raster() and is also used to derive the maximum-eta wetting
# mask from the original DEM before any WGS84 projection.
rotated_raster_to_model_xy <- function(r_om) {
  gis_rows <- terra::as.matrix(r_om, wide = TRUE)
  if (source_is_low_x) {
    t(gis_rows[nrow(gis_rows):1, , drop = FALSE])
  } else {
    t(gis_rows[, ncol(gis_rows):1, drop = FALSE])
  }
}

symmetric_raster_limits <- function(r) {
  m <- global(abs(r), "max", na.rm = TRUE)[1, 1]
  if (!is.finite(m) || m == 0) m <- 1e-8
  c(-m, m)
}

write_model_output_rasters <- function(r_om, stem, method = "bilinear",
                                       already_public = FALSE) {
  # `make_public_output_rotated()` first masks native-grid values then
  # aggregates to the requested public resolution before any geographic
  # reprojection. This keeps output geometry consistent with the native 5 m
  # grid while retaining the maximum-eta inundation extent.
  if (!already_public) r_om <- make_public_output_rotated(r_om)
  writeRaster(r_om, file.path(case_dir, paste0(stem, "_rotated_omerc.tif")),
              overwrite = TRUE)
  r_ll <- project(r_om, "EPSG:4326", method = method)
  writeRaster(r_ll, file.path(case_dir, paste0(stem, "_wgs84.tif")),
              overwrite = TRUE)
  r_ll
}


symmetric_limits <- function(z) {
  m <- max(abs(z), na.rm = TRUE)
  if (!is.finite(m) || m == 0) m <- 1e-8
  c(-m, m)
}

positive_limits <- function(z) {
  m <- max(z, na.rm = TRUE)
  if (!is.finite(m) || m <= 0) m <- 1e-8
  c(0, m)
}

depth_raster <- rast(file.path(
  case_dir, paste0("depth_", grid_tag, "_positive_water_depth.tif")
))
elevation_raster <- rast(file.path(
  case_dir, paste0("elevation_", grid_tag, "_warped.tif")
))
depth_from_file <- read_text_model(file.path(case_dir, "depth.txt"), "depth.txt")
source_is_low_x <- identical(grid$source_edge, "minimum rotated x")

# Begin with the still-water DEM mask. Once all eta frames have been read this
# is replaced by the maximum-eta inundation mask below, still on the native
# rotated grid and before any WGS84 interpolation.
dem_water_model <- is.finite(depth_from_file) & depth_from_file > 0
dem_water_mask_rotated <- ifel(depth_raster > 0, 1, 0)
dem_water_mask_wgs84 <- project(dem_water_mask_rotated, "EPSG:4326", method = "near")
display_water_model <- dem_water_model
display_water_mask_rotated <- dem_water_mask_rotated
display_water_mask_wgs84 <- dem_water_mask_wgs84
display_water_mask_label <- "still-water DEM wet cells"

apply_display_water_mask_model <- function(z) {
  z <- as.matrix(z)
  z[!display_water_model] <- NA_real_
  z
}

apply_display_water_mask_rotated <- function(r_om) {
  r_om[is.na(display_water_mask_rotated) | display_water_mask_rotated <= 0] <- NA_real_
  r_om
}

# The internal wavemaker emits energy in both directions. The buoy-side side
# is intentionally hidden from coastal forecast products before projection and
# again afterwards, preventing interpolation from bleeding across the paddle.
shoreward_of_paddle_model <- outer(
  (seq_len(grid$Mglob[1]) - 1) * grid$dx_m[1], seq_len(grid$Nglob[1]),
  function(x, y) x >= grid$x_wk_m[1]
)
shoreward_of_paddle_rotated <- model_to_rotated_raster(
  ifelse(shoreward_of_paddle_model, 1, 0), "shoreward-of-wavemaker mask"
)
shoreward_of_paddle_wgs84 <- project(shoreward_of_paddle_rotated, "EPSG:4326", method = "near")

apply_output_visibility_mask_model <- function(z) {
  z <- apply_display_water_mask_model(z)
  z[!shoreward_of_paddle_model] <- NA_real_
  z
}

apply_output_visibility_mask_rotated <- function(r_om) {
  r_om <- apply_display_water_mask_rotated(r_om)
  r_om[is.na(shoreward_of_paddle_rotated) | shoreward_of_paddle_rotated <= 0] <- NA_real_
  r_om
}

make_public_output_rotated <- function(r_om) {
  r_om <- apply_output_visibility_mask_rotated(r_om)
  if (public_output_factor_x <= 1L && public_output_factor_y <= 1L) return(r_om)

  # A public cell is retained only where all contributing native cells are
  # inundated at any saved eta time and shoreward of the wavemaker. This keeps
  # WGS84 interpolation from leaking output across the coast or source side.
  visibility <- ifel(
    is.na(display_water_mask_rotated) | is.na(shoreward_of_paddle_rotated) |
      display_water_mask_rotated <= 0 | shoreward_of_paddle_rotated <= 0,
    0, 1
  )
  # Terra receives aggregation factors in raster row/column order: y then x.
  # The model factors are x then y, while terra receives raster row/column
  # factors as y then x.
  raster_aggregation_factors <- c(public_output_factor_y, public_output_factor_x)
  public_visibility <- aggregate(
    visibility, fact = raster_aggregation_factors, fun = min, na.rm = TRUE
  )
  r_om <- aggregate(r_om, fact = raster_aggregation_factors,
                     fun = mean, na.rm = TRUE)
  r_om[is.na(public_visibility) | public_visibility <= 0] <- NA_real_
  r_om
}

# Reconstruct depth.txt on the exact native Omerc raster, then derive maps
# from that raster. This is the same pathway used for FUNWAVE outputs below.
depth_from_file_rotated <- model_to_rotated_raster(depth_from_file, "depth.txt")
depth_file_minus_raster <- depth_from_file_rotated - depth_raster
max_depth_file_delta <- global(abs(depth_file_minus_raster), "max", na.rm = TRUE)[1, 1]

buoy_ll <- vect(matrix(c(forcing$buoy_lon, forcing$buoy_lat), ncol = 2),
                type = "points", crs = "EPSG:4326")
buoy_om <- project(buoy_ll, crs(depth_raster))

# Plot all map products in WGS84, but only after reconstructing them on the
# native rotated grid. The black outline is the rectangular FUNWAVE domain.
outline_from_raster <- function(r) {
  vect(rbind(
    c(xmin(r), ymin(r)), c(xmax(r), ymin(r)),
    c(xmax(r), ymax(r)), c(xmin(r), ymax(r)),
    c(xmin(r), ymin(r))
  ), type = "polygons", crs = crs(r))
}
model_outline_om <- outline_from_raster(depth_raster)
model_outline_ll_xy <- crds(project(model_outline_om, "EPSG:4326"))

# Model x is always written from the buoy-side source edge toward the coast.
# Draw the actual internal-wavemaker line on every map so the reconstructed
# fields can be checked against their physical source location.
source_x_om <- if (source_is_low_x) {
  xmin(depth_raster) + grid$x_wk_m
} else {
  xmax(depth_raster) - grid$x_wk_m
}
# Use the actual source-line endpoints recorded by setup. The present case
# spans the full model y extent; the fallback keeps older output folders
# renderable when they lack the endpoint metadata.
source_y_first_model_m <- if ("source_y_first_cell_m" %in% names(grid)) {
  grid$source_y_first_cell_m[1]
} else {
  0
}
source_y_last_model_m <- if ("source_y_last_cell_m" %in% names(grid)) {
  grid$source_y_last_cell_m[1]
} else {
  (grid$Nglob[1] - 1) * grid$dy_m[1]
}
source_y_om <- if (source_is_low_x) {
  c(ymin(depth_raster) + source_y_first_model_m,
    ymin(depth_raster) + source_y_last_model_m)
} else {
  c(ymax(depth_raster) - source_y_first_model_m,
    ymax(depth_raster) - source_y_last_model_m)
}
source_line_om <- vect(rbind(
  c(source_x_om, source_y_om[1]),
  c(source_x_om, source_y_om[2])
), type = "lines", crs = crs(depth_raster))
source_line_ll_xy <- crds(project(source_line_om, "EPSG:4326"))

plot_geographic <- function(r_ll, ..., show_legend = FALSE, range = NULL,
                            animation = FALSE) {
  # terra::plot uses range= to fix its colour scale; zlim= is for image().
  if (animation) {
    old_par <- graphics::par(no.readonly = TRUE)
    on.exit(graphics::par(old_par), add = TRUE)
    # Reserve a real in-frame margin for the colour bar; earlier GIFs placed
    # it at the device edge where the labels could be clipped.
    graphics::par(mar = c(4.1, 4.2, 3.6, 8.8), xpd = FALSE)
    plot_legend <- list(cex = 0.72, shrink = 0.82)
    if (is.null(range)) {
      plot(r_ll, ..., plg = plot_legend)
    } else {
      plot(r_ll, ..., range = range, plg = plot_legend)
    }
  } else if (is.null(range)) {
    plot(r_ll, ...)
  } else {
    plot(r_ll, ..., range = range)
  }
  lines(model_outline_ll_xy[, 1], model_outline_ll_xy[, 2],
        col = "black", lwd = 2)
  lines(source_line_ll_xy[, 1], source_line_ll_xy[, 2],
        col = "dodgerblue3", lwd = 2, lty = 2)
  points(crds(buoy_ll)[1, 1], crds(buoy_ll)[1, 2],
         pch = 16, col = "red", cex = 1.25)
  if (show_legend) {
    legend("bottomleft",
           legend = c("FUNWAVE grid", "internal wavemaker", "buoy"),
           col = c("black", "dodgerblue3", "red"), lwd = c(2, 2, NA),
           lty = c(1, 2, NA), pch = c(NA, NA, 16),
           bty = "n", cex = if (animation) 0.62 else 0.75,
           inset = if (animation) 0.012 else 0)
  }
}

local_map_bounds <- function(r_ll, padding = 0.04) {
  e <- ext(r_ll)
  x_pad <- (xmax(e) - xmin(e)) * padding
  y_pad <- (ymax(e) - ymin(e)) * padding
  c(west = xmin(e) - x_pad, south = ymin(e) - y_pad,
    east = xmax(e) + x_pad, north = ymax(e) + y_pad)
}

apply_water_masks_wgs84 <- function(r_ll) {
  water_mask <- display_water_mask_wgs84
  if (!compareGeom(water_mask, r_ll, stopOnError = FALSE)) {
    water_mask <- resample(water_mask, r_ll, method = "near")
  }
  r_ll[is.na(water_mask) | water_mask <= 0] <- NA_real_

  paddle_mask <- shoreward_of_paddle_wgs84
  if (!compareGeom(paddle_mask, r_ll, stopOnError = FALSE)) {
    paddle_mask <- resample(paddle_mask, r_ll, method = "near")
  }
  r_ll[is.na(paddle_mask) | paddle_mask <= 0] <- NA_real_

  r_ll
}

leaflet_model_map <- function(r_ll, palette, limits, legend_title,
                              bathymetry_r_ll = NULL,
                              bathymetry_palette = ygnbu_palette,
                              bathymetry_limits = NULL) {
  b <- local_map_bounds(r_ll, padding = 0.04)
  max_b <- local_map_bounds(r_ll, padding = 0.50)
  pal <- colorNumeric(palette = palette, domain = limits, na.color = "transparent")
  m <- leaflet(width = "100%", height = 480,
               options = leafletOptions(minZoom = 12, maxZoom = 18,
                                        maxBoundsViscosity = 1))
  # These public sources require no project key. Add imagery first, then
  # explicitly activate it after the layers control is built.
  m <- addProviderTiles(m, "Esri.WorldImagery", group = "Satellite imagery")
  m <- addProviderTiles(m, "OpenStreetMap.Mapnik", group = "OpenStreetMap")
  m <- setMaxBounds(m, max_b[["west"]], max_b[["south"]],
                    max_b[["east"]], max_b[["north"]])
  m <- fitBounds(m, b[["west"]], b[["south"]], b[["east"]], b[["north"]])
  has_bathymetry <- !is.null(bathymetry_r_ll)
  if (has_bathymetry) {
    if (is.null(bathymetry_limits)) {
      bathymetry_limits <- c(0, terra::global(bathymetry_r_ll, "max", na.rm = TRUE)[1, 1])
    }
    if (!all(is.finite(bathymetry_limits)) || bathymetry_limits[2] <= bathymetry_limits[1]) {
      bathymetry_limits <- c(0, 1)
    }
    bathy_pal <- colorNumeric(
      palette = bathymetry_palette, domain = bathymetry_limits,
      na.color = "transparent"
    )
    m <- addRasterImage(
      m, bathymetry_r_ll, colors = bathy_pal, opacity = 0.78, project = FALSE,
      group = "Bathymetry (water depth)"
    )
  }
  m <- addRasterImage(m, r_ll, colors = pal, opacity = 0.88, project = FALSE,
                      group = "Model output")
  m <- addPolylines(m, lng = model_outline_ll_xy[, 1], lat = model_outline_ll_xy[, 2],
                    color = "black", weight = 2, group = "FUNWAVE grid")
  m <- addPolylines(m, lng = source_line_ll_xy[, 1], lat = source_line_ll_xy[, 2],
                    color = "dodgerblue3", weight = 2, dashArray = "6 6",
                    group = "Internal wavemaker")
  m <- addCircleMarkers(m, lng = crds(buoy_ll)[1, 1], lat = crds(buoy_ll)[1, 2],
                        radius = 5, color = "red", fillOpacity = 1,
                        label = "Port Fairy wave buoy", group = "Buoy")
  if (exists("station_metadata", inherits = TRUE) && nrow(station_metadata)) {
    m <- addCircleMarkers(
      m, lng = station_metadata$lon, lat = station_metadata$lat,
      radius = 4, color = station_colours[seq_len(nrow(station_metadata))],
      fillOpacity = 1, label = station_metadata$station_id,
      group = "Transect stations"
    )
  }
  overlay_groups <- c(
    "Model output", "FUNWAVE grid", "Internal wavemaker", "Buoy", "Transect stations"
  )
  if (has_bathymetry) {
    overlay_groups <- c("Bathymetry (water depth)", overlay_groups)
  }
  m <- addLayersControl(
    m,
    baseGroups = c("OpenStreetMap", "Satellite imagery"),
    overlayGroups = overlay_groups,
    options = layersControlOptions(collapsed = FALSE)
  )
  # Keep imagery active: hiding OpenStreetMap alone can leave a Leaflet radio
  # control with neither base layer selected in some browsers.
  m <- leaflet::hideGroup(m, "OpenStreetMap")
  m <- leaflet::showGroup(m, "Satellite imagery")
  if (has_bathymetry) {
    m <- leaflet::hideGroup(m, "Bathymetry (water depth)")
    m <- addLegend(
      m, position = "bottomleft", pal = bathy_pal, values = bathymetry_limits,
      title = "Bathymetry\nwater depth (m)", opacity = 0.88,
      group = "Bathymetry (water depth)"
    )
  }
  m <- addLegend(m, position = "bottomright", pal = pal, values = limits,
                 title = legend_title, opacity = 0.88)
  # Compact controls; line breaks in legend titles are preserved so the map
  # controls do not consume a large share of the local-domain view.
  m <- htmlwidgets::onRender(m, "
    function(el) {
      el.querySelectorAll('.info.legend').forEach(function(node) {
        node.style.maxWidth = '165px';
        node.style.whiteSpace = 'pre-line';
        node.style.lineHeight = '1.15';
      });
      el.querySelectorAll('.leaflet-control-layers').forEach(function(node) {
        node.style.maxWidth = '165px';
      });
      el.querySelectorAll('.leaflet-control-layers label').forEach(function(node) {
        node.style.whiteSpace = 'normal';
        node.style.display = 'block';
        node.style.lineHeight = '1.15';
      });
    }
  ")
  m
}

frame_local_time <- function(elapsed_s) {
  format(buoy_time_utc + elapsed_s, "%d %b %Y %H:%M %Z", tz = port_fairy_tz)
}

add_wave_frame_annotation <- function(r_ll, elapsed_s) {
  e <- ext(r_ll)
  x_span <- xmax(e) - xmin(e)
  y_span <- ymax(e) - ymin(e)
  bearing_to <- forcing$boundary_direction_to_deg_true[1]
  direction_statistic <- forcing$boundary_direction_statistic[1]
  angle <- bearing_to * pi / 180

  # The map is in longitude/latitude. Scale the longitude component by the
  # local latitude before drawing the true-north travel-direction arrow.
  arrow_y <- 0.13 * y_span
  arrow_x <- arrow_y * sin(angle) / cos(mean(c(ymin(e), ymax(e))) * pi / 180)
  arrow_dy <- arrow_y * cos(angle)
  arrow_x0 <- if (arrow_x >= 0) xmin(e) + 0.08 * x_span else xmax(e) - 0.08 * x_span
  arrow_y0 <- if (arrow_dy >= 0) ymin(e) + 0.10 * y_span else ymax(e) - 0.10 * y_span
  arrows(arrow_x0, arrow_y0, arrow_x0 + arrow_x, arrow_y0 + arrow_dy,
         col = "red", lwd = 3, length = 0.10)

  label <- c(
    paste("Local:", frame_local_time(elapsed_s)),
    sprintf("Model +%02.0f:%04.1f", floor(elapsed_s / 60), elapsed_s %% 60),
    sprintf("Buoy Hs %.2f m | Tp %.1f s", forcing$hs_m[1], forcing$tp_s[1]),
    sprintf("%s waves from %.0f° (arrow travels to %.0f°)",
            tools::toTitleCase(direction_statistic),
            forcing$boundary_direction_from_deg_true[1], bearing_to)
  )
  legend("topright", legend = label, bty = "o", bg = rgb(1, 1, 1, 0.82),
         cex = 0.74, text.col = "black")
}

elevation_wgs84 <- project(elevation_raster, "EPSG:4326", method = "bilinear")
depth_file_wgs84 <- project(depth_from_file_rotated, "EPSG:4326", method = "bilinear")
depth_delta_wgs84 <- project(depth_file_minus_raster, "EPSG:4326", method = "bilinear")

eta <- read_field("eta")

eta_paths <- list.files(results_dir, pattern = "^eta_[0-9]{5}$",
                        full.names = TRUE)
# FUNWAVE uses sequential output-file numbers in eta_00000, eta_00001, ...;
# they are not elapsed seconds. Convert those numbers using PLOT_INTV so the
# animation labels and the final-one-sixth selection use model time.
eta_file_number <- as.integer(sub("^eta_", "", basename(eta_paths)))
eta_order <- order(eta_file_number)
eta_paths <- eta_paths[eta_order]
eta_file_number <- eta_file_number[eta_order]
eta_time <- eta_file_number * grid$plot_intv_s
if (length(eta_paths) &&
    (any(!is.finite(eta_time)) || any(eta_time < 0) ||
     max(eta_time) > grid$total_time_s[1] + grid$plot_intv_s[1] ||
     max(eta_time) < grid$total_time_s[1] - grid$plot_intv_s[1])) {
  stop("eta file numbers do not map to the configured simulation period.")
}

# The previous report used depth.txt/MASK as a permanent land mask, which hid
# valid beach cells whenever their still-water depth was <= 0. Derive the
# visibility mask from the warped AHD DEM plus the greatest saved positive eta
# instead. This happens in model x/y before aggregation or reprojection, so a
# public map cannot invent a flooded cell through geographic interpolation.
eta_max_model <- NULL
if (length(eta_paths)) {
  for (path in eta_paths) {
    z <- read_text_model(path, basename(path))
    z[!is.finite(z)] <- NA_real_
    eta_max_model <- if (is.null(eta_max_model)) {
      z
    } else {
      pmax(eta_max_model, z, na.rm = TRUE)
    }
  }
  eta_max_model[!is.finite(eta_max_model)] <- NA_real_
}
applied_still_water_ahd_m <- if (
  "portland_water_level_ahd_m_applied" %in% names(forcing) &&
  is.finite(forcing$portland_water_level_ahd_m_applied[1])
) forcing$portland_water_level_ahd_m_applied[1] else 0
elevation_model <- rotated_raster_to_model_xy(elevation_raster)
max_water_depth_model <- pmax(0, applied_still_water_ahd_m - elevation_model)
if (!is.null(eta_max_model)) {
  max_water_depth_model <- applied_still_water_ahd_m - elevation_model +
    pmax(eta_max_model, 0)
  display_water_model <- is.finite(max_water_depth_model) & max_water_depth_model > 0
  display_water_mask_rotated <- model_to_rotated_raster(
    ifelse(display_water_model, 1, 0), "maximum-eta inundation mask"
  )
  display_water_mask_wgs84 <- project(display_water_mask_rotated, "EPSG:4326", method = "near")
  display_water_mask_label <- "maximum saved eta inundation extent"
}

read_public_output_raster <- function(path, label = basename(path)) {
  z <- read_text_model(path, basename(path))
  make_public_output_rotated(model_to_rotated_raster(z, label))
}
maximum_public_raster <- function(fields) {
  if (!length(fields)) return(NULL)
  out <- fields[[1]]
  if (length(fields) > 1) {
    for (k in 2:length(fields)) {
      out <- terra::app(c(out, fields[[k]]), fun = max, na.rm = TRUE)
    }
  }
  out
}

# Retain report fields only at the requested public resolution. The raw eta
# matrices above are streamed solely to establish the maximum-eta inundation
# mask; report fields themselves remain compact public-resolution rasters.
eta_public_frames <- lapply(eta_paths, read_public_output_raster)
eta_max <- maximum_public_raster(eta_public_frames)
if (!is.null(eta_max)) {
  eta_max[!is.finite(eta_max)] <- NA_real_
}

# FUNWAVE's WaveHeight=T diagnostic is written by PREVIEW_MEAN as Hsig files.
# It uses the model time step internally, so it is the correct way to assess
# wave energy rather than comparing individual 15-second eta snapshots.
hsig_paths <- list.files(results_dir, pattern = "^Hsig_[0-9]{5}$",
                         full.names = TRUE)
hsig_number <- as.integer(sub("^Hsig_", "", basename(hsig_paths)))
hsig_paths <- hsig_paths[order(hsig_number)]
hsig_public_frames <- lapply(hsig_paths, read_public_output_raster)
hsig_max <- maximum_public_raster(hsig_public_frames)
if (!is.null(hsig_max)) {
  hsig_max[!is.finite(hsig_max)] <- NA_real_
}

# `eta_max` and `hsig_max` above are already rotated public-resolution rasters,
# not model-index matrices. The original raw field files remain untouched.
eta_frames <- NULL
hsig_frames <- NULL

# Read the five FUNWAVE station files.  Each file is time, eta, u, v and the
# position metadata was created with stations.txt by the case setup script.
station_metadata_path <- file.path(case_dir, "station_metadata.csv")
station_metadata <- if (file.exists(station_metadata_path)) {
  read.csv(station_metadata_path, check.names = FALSE)
} else {
  data.frame()
}
station_timeseries <- data.frame()
if (nrow(station_metadata)) {
  station_series <- lapply(seq_len(nrow(station_metadata)), function(k) {
    path <- file.path(results_dir, sprintf("sta_%04d", k))
    if (!file.exists(path)) return(NULL)
    z <- utils::read.table(path, header = FALSE)
    if (ncol(z) < 2) return(NULL)
    names(z)[seq_len(min(4, ncol(z)))] <- c("time_s", "eta_m", "u_m_s", "v_m_s")[seq_len(min(4, ncol(z)))]
    z$station_id <- station_metadata$station_id[k]
    z
  })
  station_series <- Filter(Negate(is.null), station_series)
  if (length(station_series)) station_timeseries <- do.call(rbind, station_series)
}
if (nrow(station_timeseries)) {
  station_plot <- plotly::plot_ly()
  for (k in seq_len(nrow(station_metadata))) {
    station_name <- station_metadata$station_id[k]
    z <- station_timeseries[station_timeseries$station_id == station_name, , drop = FALSE]
    if (nrow(z)) {
      station_plot <- plotly::add_lines(
        station_plot, data = z, x = ~time_s, y = ~eta_m, name = station_name,
        line = list(color = station_colours[k], width = 2),
        hovertemplate = paste0(station_name, "<br>t = %{x:.1f} s<br>eta = %{y:.3f} m<extra></extra>")
      )
    }
  }
  station_plot <- plotly::layout(
    station_plot, title = "FUNWAVE eta along buoy-to-nearshore transect",
    xaxis = list(title = "Model elapsed time (s)"),
    yaxis = list(title = "Free-surface elevation, eta (m)", range = eta_display_limits),
    legend = list(orientation = "h")
  )
  site_assets_dir <- file.path(report_site_dir, "assets")
  dir.create(site_assets_dir, recursive = TRUE, showWarnings = FALSE)
  htmlwidgets::saveWidget(
    station_plot, file.path(site_assets_dir, "port-fairy-transect-timeseries.html"),
    selfcontained = FALSE
  )
}

# Every run carries a small, explicit observed-condition context written by
# setup_port_fairy_funwave.R. Historical cases are centred on their requested
# time; the daily forecast shows the available seven days leading up to its
# selected forcing observation.
historical_wave_path <- file.path(case_dir, "historical_wave_observations.csv")
historical_portland_path <- file.path(case_dir, "historical_portland_water_levels.csv")
historical_window_metadata_path <- file.path(
  case_dir, "historical_observation_window_metadata.csv"
)
historical_observation_available <- all(file.exists(
  historical_wave_path, historical_portland_path, historical_window_metadata_path
))
historical_observation_page_url <- "assets/port-fairy-historical-observations.html"
historical_observation_page_path <- file.path(
  report_site_dir, historical_observation_page_url
)
historical_observation_page_status <- "Not a historical report."
historical_observation_summary <- data.frame(
  Diagnostic = "Historical observed-condition context", Value = "Not generated for the daily report.",
  stringsAsFactors = FALSE, check.names = FALSE
)

parse_utc_timestamp <- function(x) {
  as.POSIXct(sub(" UTC$", "", as.character(x)),
             format = "%Y-%m-%d %H:%M:%S", tz = "UTC")
}
add_requested_time_rule <- function(widget, requested_time, values) {
  finite_values <- values[is.finite(values)]
  if (!length(finite_values) || is.na(requested_time)) return(widget)
  limits <- range(finite_values)
  if (diff(limits) == 0) limits <- limits + c(-0.1, 0.1)
  plotly::add_segments(
    widget, x = requested_time, xend = requested_time,
    y = limits[1], yend = limits[2], inherit = FALSE, showlegend = FALSE,
    line = list(color = "#d7301f", width = 1.5, dash = "dot"),
    hoverinfo = "skip"
  )
}

add_portland_mean_annual_maximum_rule <- function(widget, times, level_ahd_m) {
  valid_times <- times[!is.na(times)]
  if (!is.finite(level_ahd_m) || length(valid_times) < 2) return(widget)
  plotly::add_segments(
    widget,
    x = min(valid_times), xend = max(valid_times),
    y = level_ahd_m, yend = level_ahd_m,
    inherit = FALSE, name = "Portland mean annual maximum",
    showlegend = FALSE,
    line = list(color = "#d95f0e", width = 1.7, dash = "dash"),
    hovertemplate = paste0(
      "Portland mean annual maximum: ",
      formatC(level_ahd_m, format = "f", digits = 2),
      " m AHD<extra></extra>"
    )
  )
}

if (historical_observation_available) {
  historical_wave <- tryCatch(
    read.csv(historical_wave_path, check.names = FALSE),
    error = function(e) {
      historical_observation_page_status <<- paste0("Could not read historical wave observations: ", conditionMessage(e))
      NULL
    }
  )
  historical_portland <- tryCatch(
    read.csv(historical_portland_path, check.names = FALSE),
    error = function(e) {
      historical_observation_page_status <<- paste0("Could not read historical Portland water levels: ", conditionMessage(e))
      NULL
    }
  )
  historical_window_metadata <- tryCatch(
    read.csv(historical_window_metadata_path, check.names = FALSE),
    error = function(e) {
      historical_observation_page_status <<- paste0("Could not read historical observation metadata: ", conditionMessage(e))
      NULL
    }
  )
  required_wave_columns <- c("time_utc", "hs_m", "tp_s", "direction_from_deg_true")
  required_tide_columns <- c("time_utc", "water_level_ahd_m")
  if (is.null(historical_wave) || is.null(historical_portland) ||
      is.null(historical_window_metadata) || nrow(historical_window_metadata) != 1 ||
      !all(required_wave_columns %in% names(historical_wave)) ||
      !all(required_tide_columns %in% names(historical_portland))) {
    historical_observation_available <- FALSE
    if (identical(historical_observation_page_status, "Not a historical report.")) {
      historical_observation_page_status <- "Historical observed-condition files have an unexpected structure."
    }
  } else {
    historical_wave$time_utc <- parse_utc_timestamp(historical_wave$time_utc)
    historical_portland$time_utc <- parse_utc_timestamp(historical_portland$time_utc)
    observation_context_mode <- if ("observation_mode" %in% names(historical_window_metadata)) {
      as.character(historical_window_metadata$observation_mode[1])
    } else {
      "historical"
    }
    observation_is_historical <- identical(observation_context_mode, "historical")
    observation_context_label <- if (observation_is_historical) "historical run" else "forecast"
    historical_wave$hover_local <- format(
      historical_wave$time_utc, tz = port_fairy_tz, format = "%d %b %Y %H:%M %Z"
    )
    historical_portland$hover_local <- format(
      historical_portland$time_utc, tz = port_fairy_tz, format = "%d %b %Y %H:%M %Z"
    )
    requested_historical_time <- parse_utc_timestamp(
      historical_window_metadata$requested_time_utc[1]
    )
    selected_historical_time <- parse_utc_timestamp(
      historical_window_metadata$selected_buoy_time_utc[1]
    )
    window_label <- paste0(
      format(parse_utc_timestamp(historical_window_metadata$window_start_utc[1]),
             tz = port_fairy_tz, format = "%d %b %Y %H:%M %Z"),
      " to ",
      format(parse_utc_timestamp(historical_window_metadata$window_end_utc[1]),
             tz = port_fairy_tz, format = "%d %b %Y %H:%M %Z")
    )
    portland_mean_annual_maximum_ahd_m <- if (
      "portland_mean_annual_maximum_ahd_m" %in% names(forcing)
    ) as.numeric(forcing$portland_mean_annual_maximum_ahd_m[1]) else NA_real_
    portland_mean_annual_maximum_years <- if (
      "portland_mean_annual_maximum_years" %in% names(forcing)
    ) as.integer(forcing$portland_mean_annual_maximum_years[1]) else NA_integer_
    local_time_axis <- list(
      title = "Port Fairy local time",
      type = "date",
      tickformat = "%d %b<br>%H:%M",
      hoverformat = "%d %b %Y %H:%M"
    )

    p_hs <- plotly::plot_ly(
      historical_wave, x = ~time_utc, y = ~hs_m, text = ~hover_local,
      type = "scatter", mode = "lines+markers",
      line = list(color = ygnbu_palette[9], width = 2),
      marker = list(color = ygnbu_palette[9], size = 4),
      hovertemplate = "%{text}<br>Hs: %{y:.2f} m<extra></extra>",
      showlegend = FALSE
    )
    p_hs <- add_requested_time_rule(p_hs, requested_historical_time,
                                    c(0, historical_wave$hs_m))
    p_hs <- plotly::add_markers(
      p_hs, data = data.frame(time_utc = selected_historical_time,
                              value = forcing$hs_m[1]),
      x = ~time_utc, y = ~value, inherit = FALSE, showlegend = FALSE,
      marker = list(color = "#d7301f", size = 9, symbol = "diamond"),
      hovertemplate = "FUNWAVE forcing record<br>Hs: %{y:.2f} m<extra></extra>"
    )
    p_hs <- plotly::layout(
      p_hs, xaxis = local_time_axis,
      yaxis = list(title = "Hs (m)", range = c(0, 6))
    )

    p_tp <- plotly::plot_ly(
      historical_wave, x = ~time_utc, y = ~tp_s, text = ~hover_local,
      type = "scatter", mode = "lines+markers",
      line = list(color = ygnbu_palette[7], width = 2),
      marker = list(color = ygnbu_palette[7], size = 4),
      hovertemplate = "%{text}<br>Tp: %{y:.2f} s<extra></extra>",
      showlegend = FALSE
    )
    p_tp <- add_requested_time_rule(p_tp, requested_historical_time, historical_wave$tp_s)
    p_tp <- plotly::add_markers(
      p_tp, data = data.frame(time_utc = selected_historical_time,
                              value = forcing$tp_s[1]),
      x = ~time_utc, y = ~value, inherit = FALSE, showlegend = FALSE,
      marker = list(color = "#d7301f", size = 9, symbol = "diamond"),
      hovertemplate = "FUNWAVE forcing record<br>Tp: %{y:.2f} s<extra></extra>"
    )
    p_tp <- plotly::layout(
      p_tp, xaxis = local_time_axis,
      yaxis = list(title = "Tp (s)", rangemode = "tozero")
    )

    p_direction <- plotly::plot_ly(
      historical_wave, x = ~time_utc, y = ~direction_from_deg_true,
      text = ~hover_local, type = "scatter", mode = "markers",
      marker = list(color = ygnbu_palette[5], size = 5),
      hovertemplate = "%{text}<br>Wave direction from: %{y:.0f}°<extra></extra>",
      showlegend = FALSE
    )
    p_direction <- add_requested_time_rule(
      p_direction, requested_historical_time, c(0, 360)
    )
    p_direction <- plotly::add_markers(
      p_direction, data = data.frame(time_utc = selected_historical_time,
                                     value = forcing$boundary_direction_from_deg_true[1]),
      x = ~time_utc, y = ~value, inherit = FALSE, showlegend = FALSE,
      marker = list(color = "#d7301f", size = 9, symbol = "diamond"),
      hovertemplate = "FUNWAVE forcing record<br>Direction from: %{y:.0f}°<extra></extra>"
    )
    p_direction <- plotly::layout(
      p_direction, xaxis = local_time_axis,
      yaxis = list(title = "Direction from (°)", range = c(0, 360), dtick = 90)
    )

    p_tide <- plotly::plot_ly(
      historical_portland, x = ~time_utc, y = ~water_level_ahd_m,
      text = ~hover_local, type = "scatter", mode = "lines",
      line = list(color = ygnbu_palette[3], width = 2),
      hovertemplate = "%{text}<br>Portland water level: %{y:.2f} m AHD<extra></extra>",
      showlegend = FALSE
    )
    p_tide <- add_requested_time_rule(
      p_tide, requested_historical_time, historical_portland$water_level_ahd_m
    )
    p_tide <- add_portland_mean_annual_maximum_rule(
      p_tide, historical_portland$time_utc, portland_mean_annual_maximum_ahd_m
    )
    applied_portland_time <- parse_utc_timestamp(forcing$portland_water_level_time_utc[1])
    if (is.finite(forcing$portland_water_level_ahd_m_applied[1]) &&
        !is.na(applied_portland_time)) {
      p_tide <- plotly::add_markers(
        p_tide,
        data = data.frame(
          time_utc = applied_portland_time,
          value = forcing$portland_water_level_ahd_m_applied[1]
        ),
        x = ~time_utc, y = ~value, inherit = FALSE, showlegend = FALSE,
        marker = list(color = "#d7301f", size = 9, symbol = "diamond"),
        hovertemplate = "FUNWAVE applied level<br>Portland: %{y:.2f} m AHD<extra></extra>"
      )
    }
    p_tide <- plotly::layout(
      p_tide, xaxis = local_time_axis,
      yaxis = list(title = "Portland level (m AHD)")
    )

    historical_observation_plot <- plotly::subplot(
      p_hs, p_tp, p_direction, p_tide,
      nrows = 4, shareX = FALSE, titleY = TRUE, margin = 0.045,
      heights = c(0.25, 0.25, 0.25, 0.25)
    )
    historical_observation_plot <- plotly::layout(
      historical_observation_plot,
      title = paste0("Observed Port Fairy waves and Portland water level: ", window_label),
      xaxis = local_time_axis,
      xaxis2 = local_time_axis,
      xaxis3 = local_time_axis,
      xaxis4 = local_time_axis,
      hovermode = "x unified",
      annotations = list(list(
        text = paste0(
          "Red dotted line: requested model time. Red diamond: observation applied to FUNWAVE.",
          if (is.finite(portland_mean_annual_maximum_ahd_m)) {
            paste0(
              " Orange dashed line: Portland mean annual maximum ",
              formatC(portland_mean_annual_maximum_ahd_m, format = "f", digits = 2),
              " m AHD (", portland_mean_annual_maximum_years, " sufficiently complete local years)."
            )
          } else {
            ""
          }
        ),
        x = 0, y = -0.08, xref = "paper", yref = "paper", showarrow = FALSE,
        xanchor = "left", font = list(size = 11)
      ))
    )
    historical_observation_plot <- plotly::config(
      historical_observation_plot, displaylogo = FALSE, responsive = TRUE
    )
    dir.create(dirname(historical_observation_page_path), recursive = TRUE, showWarnings = FALSE)
    historical_observation_page_status <- tryCatch({
      htmlwidgets::saveWidget(
        historical_observation_plot, historical_observation_page_path, selfcontained = FALSE
      )
      if (!file.exists(historical_observation_page_path) ||
          file.info(historical_observation_page_path)$size < 1000) {
        stop("The historical observed-condition page was not written.")
      }
      "Published interactive Hs, Tp, direction and Portland water-level plots."
    }, error = function(e) {
      paste0("Could not write historical observed-condition page: ", conditionMessage(e))
    })
    historical_observation_summary <- data.frame(
      Diagnostic = c("Requested historical time", "FUNWAVE buoy forcing record",
                     "Window", "QC 1/2 wave observations", "Portland observations",
                     "Water-level datum", "Portland mean annual maximum",
                     "Complete local years used", "Observed-condition page"),
      Value = c(
        format(requested_historical_time, tz = port_fairy_tz, usetz = TRUE),
        format(selected_historical_time, tz = port_fairy_tz, usetz = TRUE),
        window_label,
        historical_window_metadata$wave_records_qc_1_or_2[1],
        historical_window_metadata$portland_records[1],
        historical_window_metadata$portland_datum[1],
        ifelse(is.finite(portland_mean_annual_maximum_ahd_m),
               paste0(formatC(portland_mean_annual_maximum_ahd_m, format = "f", digits = 2), " m AHD"),
               "not available"),
        ifelse(is.finite(portland_mean_annual_maximum_years),
               portland_mean_annual_maximum_years, "not available"),
        historical_observation_page_status
      ),
      stringsAsFactors = FALSE, check.names = FALSE
    )
  }
}

# The retained particle pages use FUNWAVE's time-mean Umean/Vmean fields, not
# phase-resolved U/V snapshots. That avoids using wave-orbital velocities as
# currents: the latter reverse each wave period and make particles wobble.
site_assets_dir <- file.path(report_site_dir, "assets")
dir.create(site_assets_dir, recursive = TRUE, showWarnings = FALSE)
current_particle_page_path <- file.path(site_assets_dir, "port-fairy-flow-particles.html")
current_particle_page_url <- "assets/port-fairy-flow-particles.html"
current_particle_page_status <- "No paired Umean/Vmean output was produced."
u_paths <- list.files(results_dir, pattern = "^Umean_[0-9]{5}$",
                      ignore.case = TRUE, full.names = TRUE)
v_paths <- list.files(results_dir, pattern = "^Vmean_[0-9]{5}$",
                      ignore.case = TRUE, full.names = TRUE)
current_file_number <- function(paths) {
  as.integer(sub("^[[:alpha:]]+_", "", basename(paths)))
}
u_numbers <- current_file_number(u_paths)
v_numbers <- current_file_number(v_paths)
common_current_numbers <- sort(intersect(u_numbers, v_numbers))
current_numbers <- integer()
if (length(common_current_numbers)) {
  # A Umean/Vmean number labels the mean-output record, not elapsed seconds.
  # Use the newest paired record as the stable current field.
  current_numbers <- tail(common_current_numbers, 1)
}
current_label <- "No paired Umean/Vmean output was produced"
current_time_label <- "n/a"
current_peak_speed <- NA_real_
current_speed_floor <- NA_real_
current_seed_count <- 0L
current_streamline_count <- 0L
current_public_cell_count <- 0L
current_public_grid_shape <- "n/a"
current_selected_files <- "none"
current_candidate_count <- 0L
current_source_exclusion_m <- max(200, 20 * max(grid$dx_m[1], grid$dy_m[1]))
# Smooth the vector components, never the angle. A 10 m Gaussian standard
# deviation (about 24 m full-width at half-maximum) suppresses most
# crest/trough residuals while retaining finer along-beach circulation.
current_smoothing_sigma_m <- 10
current_smoothing_label <- paste0(
  "masked Gaussian component smoothing: sigma ",
  metres_label(current_smoothing_sigma_m), " m (FWHM ",
  metres_label(2.355 * current_smoothing_sigma_m), " m)"
)

# Umean/Vmean close to an internal wavemaker are dominated by the numerical
# source zone rather than by an interpretable nearshore circulation pattern.
# Keep the source visible as a reference line, but exclude a 200 m (minimum)
# shoreward buffer from all current-path seeding and particle display fields.
current_display_mask_model <- outer(
  (seq_len(grid$Mglob[1]) - 1) * grid$dx_m[1], seq_len(grid$Nglob[1]),
  function(x, y) x >= grid$x_wk_m[1] + current_source_exclusion_m
)
current_display_mask_rotated <- model_to_rotated_raster(
  ifelse(current_display_mask_model, 1, NA_real_), "current-display mask"
)
apply_current_display_mask_model <- function(z) {
  z <- apply_output_visibility_mask_model(z)
  z[!current_display_mask_model] <- NA_real_
  z
}
apply_current_display_mask_rotated <- function(r_om) {
  # Callers have already run make_public_output_rotated(), which applies the
  # maximum-eta and wavemaker masks at the field's own resolution. Do not
  # apply native masks again here, because terra rejects mismatched geometry.
  display_mask <- current_display_mask_rotated
  if (!compareGeom(display_mask, r_om, stopOnError = FALSE)) {
    display_mask <- resample(display_mask, r_om, method = "near")
  }
  r_om[is.na(display_mask) | display_mask <= 0] <- NA_real_
  r_om
}

smooth_mean_current_components <- function(u, v, sigma_m = current_smoothing_sigma_m) {
  # Work in native FUNWAVE x/y before any aggregation or geographic projection.
  # Invalid land/source cells carry zero numerator *and* zero weight, so this
  # normalized filter cannot import values across the existing water mask.
  valid <- is.finite(u) & is.finite(v)
  if (!any(valid) || !is.finite(sigma_m) || sigma_m <= 0) {
    u[!valid] <- NA_real_
    v[!valid] <- NA_real_
    return(list(u = u, v = v))
  }

  radius_x <- max(1L, ceiling(3 * sigma_m / grid$dx_m[1]))
  radius_y <- max(1L, ceiling(3 * sigma_m / grid$dy_m[1]))
  x_m <- seq.int(-radius_x, radius_x) * grid$dx_m[1]
  y_m <- seq.int(-radius_y, radius_y) * grid$dy_m[1]
  weights <- outer(x_m, y_m, function(x, y) {
    exp(-0.5 * ((x / sigma_m)^2 + (y / sigma_m)^2))
  })
  weights <- weights / sum(weights)

  matrix_raster <- function(z) {
    r <- rast(nrows = nrow(z), ncols = ncol(z),
              xmin = 0, xmax = nrow(z), ymin = 0, ymax = ncol(z))
    values(r) <- as.numeric(t(z))
    r
  }
  raster_matrix <- function(r, n_x, n_y) {
    matrix(values(r), nrow = n_x, ncol = n_y, byrow = TRUE)
  }
  focal_normalized <- function(z) {
    z[!valid] <- 0
    numerator <- focal(matrix_raster(z), w = weights, fun = "sum", fillvalue = 0)
    denominator <- focal(matrix_raster(ifelse(valid, 1, 0)), w = weights,
                         fun = "sum", fillvalue = 0)
    out <- raster_matrix(numerator, nrow(z), ncol(z)) /
      raster_matrix(denominator, nrow(z), ncol(z))
    out[!valid] <- NA_real_
    out
  }
  list(u = focal_normalized(u), v = focal_normalized(v))
}

model_xy_to_wgs84 <- function(xy_model) {
  xy_om <- if (source_is_low_x) {
    cbind(xmin(depth_raster) + xy_model[, 1], ymin(depth_raster) + xy_model[, 2])
  } else {
    cbind(xmax(depth_raster) - xy_model[, 1], ymax(depth_raster) - xy_model[, 2])
  }
  crds(project(vect(xy_om, type = "points", crs = crs(depth_raster)), "EPSG:4326"))
}

trace_current_streamline <- function(seed_xy, u_flow, v_flow, speed_floor,
                                     source_exclusion_m, step_m = 5,
                                     n_steps = 520) {
  # Bilinear midpoint integration avoids the jagged nearest-grid-cell paths
  # that occurred when a particle crossed a 5 m cell boundary.
  velocity_at <- function(xy) {
    g <- xy / c(grid$dx_m[1], grid$dy_m[1]) + 1
    i <- floor(g[1]); j <- floor(g[2])
    if (i < 1 || j < 1 || i >= nrow(u_flow) || j >= ncol(u_flow)) return(NULL)
    fx <- g[1] - i; fy <- g[2] - j
    u_values <- c(u_flow[i, j], u_flow[i + 1, j],
                  u_flow[i, j + 1], u_flow[i + 1, j + 1])
    v_values <- c(v_flow[i, j], v_flow[i + 1, j],
                  v_flow[i, j + 1], v_flow[i + 1, j + 1])
    if (any(!is.finite(u_values)) || any(!is.finite(v_values))) return(NULL)
    weights <- c((1 - fx) * (1 - fy), fx * (1 - fy),
                 (1 - fx) * fy, fx * fy)
    c(u = sum(weights * u_values), v = sum(weights * v_values))
  }
  inside <- function(xy) {
    xy[1] >= grid$x_wk_m[1] + source_exclusion_m &&
      xy[1] <= (nrow(u_flow) - 1) * grid$dx_m[1] &&
      xy[2] >= 0 && xy[2] <= (ncol(u_flow) - 1) * grid$dy_m[1]
  }
  advance <- function(direction) {
    xy <- seed_xy
    out <- matrix(xy, ncol = 2)
    for (step in seq_len(n_steps)) {
      v0 <- velocity_at(xy)
      speed0 <- if (is.null(v0)) NA_real_ else sqrt(sum(v0^2))
      if (!is.finite(speed0) || speed0 < speed_floor) break
      midpoint <- xy + direction * 0.5 * step_m * v0 / speed0
      vmid <- velocity_at(midpoint)
      speed_mid <- if (is.null(vmid)) NA_real_ else sqrt(sum(vmid^2))
      if (!is.finite(speed_mid) || speed_mid < speed_floor) break
      candidate <- xy + direction * step_m * vmid / speed_mid
      if (!inside(candidate)) break
      xy <- candidate
      out <- rbind(out, xy)
    }
    out
  }
  backward <- advance(-1)
  forward <- advance(1)
  rbind(backward[nrow(backward):1, , drop = FALSE], forward[-1, , drop = FALSE])
}

if (length(current_numbers)) {
  read_current_component <- function(paths, numbers) {
    path_numbers <- current_file_number(paths)
    fields <- lapply(numbers, function(number) {
      path <- paths[match(number, path_numbers)]
      apply_current_display_mask_model(read_text_model(path, basename(path)))
    })
    total <- matrix(0, nrow = nrow(fields[[1]]), ncol = ncol(fields[[1]]))
    count <- matrix(0, nrow = nrow(fields[[1]]), ncol = ncol(fields[[1]]))
    for (field in fields) {
      keep <- is.finite(field)
      total[keep] <- total[keep] + field[keep]
      count[keep] <- count[keep] + 1
    }
    total[count == 0] <- NA_real_
    total[count > 0] <- total[count > 0] / count[count > 0]
    total
  }

  current_selected_files <- paste(sprintf("%05d", current_numbers), collapse = ", ")
  u_flow <- read_current_component(u_paths, current_numbers)
  v_flow <- read_current_component(v_paths, current_numbers)
  current_smoothed <- smooth_mean_current_components(u_flow, v_flow)
  u_flow <- current_smoothed$u
  v_flow <- current_smoothed$v
  current_speed <- sqrt(u_flow^2 + v_flow^2)
  current_label <- paste0("Time-mean depth-averaged current: latest Umean/Vmean output; ",
                          current_smoothing_label)
  current_time_label <- paste0(metres_label(grid$mean_wave_interval_s[1]),
                               " s FUNWAVE averaging interval; ",
                               current_smoothing_label)
  valid_speed <- current_speed[is.finite(current_speed) & current_speed > 0]
  # The old 5 mm/s floor accepted only the rapid source-zone flow.  Use a
  # small data-derived threshold so valid weak circulation can seed paths.
  speed_floor <- if (length(valid_speed)) {
    max(0.0005, 0.10 * as.numeric(stats::quantile(valid_speed, 0.10, na.rm = TRUE)))
  } else {
    Inf
  }
  current_peak_speed <- if (length(valid_speed)) max(valid_speed) else NA_real_
  current_speed_floor <- speed_floor

  x_margin <- max(4 * grid$dx_m[1], 50)
  y_margin <- max(4 * grid$dy_m[1], 100)
  x_start <- grid$x_wk_m[1] + current_source_exclusion_m + x_margin
  x_end <- (nrow(u_flow) - 1) * grid$dx_m[1] - x_margin
  y_start <- y_margin
  y_end <- (ncol(u_flow) - 1) * grid$dy_m[1] - y_margin
  seeds <- if (x_start < x_end && y_start < y_end && is.finite(speed_floor)) {
    as.matrix(expand.grid(
      seq(x_start, x_end, length.out = 37),
      seq(y_start, y_end, length.out = 27)
    ))
  } else {
    matrix(numeric(), ncol = 2)
  }
  trace_seed_set <- function(seed_set, threshold) {
    candidates <- lapply(seq_len(nrow(seed_set)), function(k) {
      ij <- pmax(c(1, 1), pmin(c(nrow(u_flow), ncol(u_flow)),
                               round(seed_set[k, ] / c(grid$dx_m[1], grid$dy_m[1])) + 1L))
      if (!is.finite(current_speed[ij[1], ij[2]]) || current_speed[ij[1], ij[2]] < threshold) return(NULL)
      path <- trace_current_streamline(
        seed_set[k, ], u_flow, v_flow, threshold,
        source_exclusion_m = current_source_exclusion_m,
        step_m = max(2.5, min(7.5, min(grid$dx_m[1], grid$dy_m[1]))),
        n_steps = 520
      )
      if (nrow(path) < 20) return(NULL)
      list(seed = as.numeric(seed_set[k, ]), path = path)
    })
    Filter(Negate(is.null), candidates)
  }
  select_spread_streamlines <- function(candidates, max_paths = 288L,
                                        min_separation_m = 70) {
    if (!length(candidates)) return(list())
    path_lengths <- vapply(candidates, function(z) nrow(z$path), integer(1))
    remaining <- seq_along(candidates)
    first <- remaining[which.max(path_lengths[remaining])]
    chosen_indices <- first
    chosen_seeds <- matrix(candidates[[first]]$seed, nrow = 1)
    remaining <- setdiff(remaining, first)
    while (length(chosen_indices) < max_paths && length(remaining)) {
      nearest_distance <- vapply(remaining, function(candidate_index) {
        candidate <- candidates[[candidate_index]]
        min(sqrt(rowSums((chosen_seeds - matrix(
          candidate$seed, nrow = nrow(chosen_seeds), ncol = 2, byrow = TRUE
        ))^2)))
      }, numeric(1))
      eligible <- which(nearest_distance >= min_separation_m)
      if (!length(eligible)) break
      farthest <- max(nearest_distance[eligible])
      ties <- eligible[abs(nearest_distance[eligible] - farthest) < 1e-8]
      winner_position <- ties[which.max(path_lengths[remaining[ties]])]
      winner <- remaining[winner_position]
      chosen_indices <- c(chosen_indices, winner)
      chosen_seeds <- rbind(chosen_seeds, candidates[[winner]]$seed)
      remaining <- remaining[-winner_position]
    }
    candidates[chosen_indices]
  }
  current_seed_count <- nrow(seeds)
  streamline_candidates <- trace_seed_set(seeds, speed_floor)
  current_candidate_count <- length(streamline_candidates)
  selected_streamlines <- select_spread_streamlines(streamline_candidates)
  # Do not silently publish a blank animation because the initially selected
  # percentile threshold was too strict. This lower threshold remains based on
  # real FUNWAVE Umean/Vmean output; it merely allows slow, valid current paths to be
  # shown and records the fallback in the diagnostics.
  if (!length(selected_streamlines) && length(valid_speed)) {
    fallback_floor <- max(0.0001, 0.01 * as.numeric(stats::quantile(
      valid_speed, 0.05, na.rm = TRUE
    )))
    if (is.finite(fallback_floor) && fallback_floor < speed_floor) {
      speed_floor <- fallback_floor
      current_speed_floor <- speed_floor
      streamline_candidates <- trace_seed_set(seeds, speed_floor)
      current_candidate_count <- length(streamline_candidates)
      selected_streamlines <- select_spread_streamlines(streamline_candidates)
      if (length(selected_streamlines)) {
        current_label <- paste0(current_label, " (low-speed path fallback)")
      }
    }
  }
  streamlines_model <- lapply(selected_streamlines, function(z) z$path)
  streamlines_ll <- lapply(streamlines_model, model_xy_to_wgs84)
  streamlines_ll <- Filter(function(z) {
    is.matrix(z) && nrow(z) >= 4 && all(is.finite(z))
  }, streamlines_ll)
  current_streamline_count <- length(streamlines_ll)

  west <- min(model_outline_ll_xy[, 1]); east <- max(model_outline_ll_xy[, 1])
  south <- min(model_outline_ll_xy[, 2]); north <- max(model_outline_ll_xy[, 2])
  x_pad <- (east - west) * 0.04; y_pad <- (north - south) * 0.04
  current_bounds <- list(c(south - y_pad, west - x_pad), c(north + y_pad, east + x_pad))
  x_max_pad <- (east - west) * 0.50; y_max_pad <- (north - south) * 0.50
  current_max_bounds <- list(c(south - y_max_pad, west - x_max_pad),
                             c(north + y_max_pad, east + x_max_pad))
  outline_latlng <- lapply(seq_len(nrow(model_outline_ll_xy)), function(k) {
    unname(as.numeric(model_outline_ll_xy[k, c(2, 1)]))
  })
  source_latlng <- lapply(seq_len(nrow(source_line_ll_xy)), function(k) {
    unname(as.numeric(source_line_ll_xy[k, c(2, 1)]))
  })
  buoy_latlng <- unname(as.numeric(c(crds(buoy_ll)[1, 2], crds(buoy_ll)[1, 1])))
  debug_prefix <- paste(
    "Paired Umean/Vmean files:", length(common_current_numbers),
    "Selected file number(s):", current_selected_files,
    "Field:", current_label,
    "Output window:", current_time_label,
    "Peak speed (m/s):", ifelse(is.finite(current_peak_speed), format(round(current_peak_speed, 3), nsmall = 3), "n/a"),
    "Trace floor (m/s):", ifelse(is.finite(current_speed_floor), format(round(current_speed_floor, 3), nsmall = 3), "n/a"),
    "Wavemaker exclusion (m):", format(round(current_source_exclusion_m, 1), trim = TRUE),
    "Candidate paths:", current_candidate_count,
    sep = "\n"
  )

  # The particle paths are traced through time-mean Umean/Vmean on the native
  # rotated grid using bilinear, midpoint integration. They are then projected
  # to WGS84 as paths, retaining geographic direction without cell-to-cell
  # stair-stepping.
  particle_routes <- lapply(streamlines_ll, function(z) {
    lapply(seq_len(nrow(z)), function(k) unname(as.numeric(z[k, c(2, 1)])))
  })
  current_public_cell_count <- sum(vapply(particle_routes, length, integer(1)))
  current_public_grid_shape <- paste0(metres_label(grid$dx_m[1]), "m x ",
                                      metres_label(grid$dy_m[1]), "m native paths")
  option2_data <- list(
    bounds = current_bounds,
    max_bounds = current_max_bounds,
    outline = outline_latlng,
    source = source_latlng,
    buoy = buoy_latlng,
    routes = particle_routes,
    label = paste(current_label, paste0("Output window: ", current_time_label),
                  "Particle trails follow smoothly integrated Umean/Vmean paths on the native rotated grid.",
                  "Animation speed is visual only; direction is the simulated flow direction.",
                  sep = "\n"),
    debug = paste(debug_prefix, "Native current path grid:", current_public_grid_shape,
                  "Animated paths:", length(particle_routes),
                  "Path vertices:", current_public_cell_count,
                  sep = "\n"),
    routeCount = length(particle_routes)
  )
  option2_js <- paste0(
    "function(el, x, data) {",
    "var map=this;if(!map||typeof map.getPanes!=='function')map=this.map||this._map;",
    "if(!map||typeof map.getPanes!=='function'){console.error('Leaflet map was not available for flow particles');return;}",
    "document.documentElement.style.height='100%';document.body.style.margin='0';document.body.style.height='100%';el.style.width='100vw';el.style.height='100vh';",
    "var routes=data.routes||[],bounds=L.latLngBounds(data.bounds),maxBounds=L.latLngBounds(data.max_bounds||data.bounds),particles=[];",
    "function distanceM(a,b){var r=6371000,dlat=(b[0]-a[0])*Math.PI/180,dlon=(b[1]-a[1])*Math.PI/180,lat=(a[0]+b[0])*Math.PI/360;return r*Math.sqrt(dlat*dlat+Math.cos(lat)*Math.cos(lat)*dlon*dlon);}",
    "function prepareRoute(points){if(!points||points.length<4)return null;var cumulative=[0];for(var i=1;i<points.length;i++)cumulative.push(cumulative[i-1]+distanceM(points[i-1],points[i]));var length=cumulative[cumulative.length-1];return length>40?{points:points,cumulative:cumulative,length:length}:null;}",
    "function pointAt(route,metres){if(!route||metres<0||metres>route.length)return null;var i=0;while(i<route.cumulative.length-2&&route.cumulative[i+1]<metres)i++;var a=route.points[i],b=route.points[i+1],span=route.cumulative[i+1]-route.cumulative[i],f=span>0?(metres-route.cumulative[i])/span:0;return [a[0]+f*(b[0]-a[0]),a[1]+f*(b[1]-a[1])];}",
    "routes.map(prepareRoute).filter(function(route){return route!==null;}).forEach(function(route,i){for(var j=0;j<4;j++)particles.push({route:route,distance:(i*47+j*route.length/4)%route.length});});",
    "var FlowTrails=L.Layer.extend({initialize:function(){this._running=false;},onAdd:function(m){this._map=m;this._canvas=L.DomUtil.create('canvas','leaflet-zoom-animated',m.getPanes().overlayPane);this._canvas.style.pointerEvents='none';this._canvas.style.zIndex=650;this._resize();m.on('resize moveend zoomend',this._resize,this);this._running=true;this._last=performance.now();this._frame=window.requestAnimationFrame(this._animate.bind(this));},onRemove:function(m){this._running=false;window.cancelAnimationFrame(this._frame);m.off('resize moveend zoomend',this._resize,this);L.DomUtil.remove(this._canvas);this._map=null;},_resize:function(){if(!this._map||!this._canvas)return;var s=this._map.getSize(),d=window.devicePixelRatio||1;this._canvas.width=Math.max(1,Math.round(s.x*d));this._canvas.height=Math.max(1,Math.round(s.y*d));this._canvas.style.width=s.x+'px';this._canvas.style.height=s.y+'px';this._topLeft=this._map.containerPointToLayerPoint([0,0]);L.DomUtil.setPosition(this._canvas,this._topLeft);this._ctx=this._canvas.getContext('2d');this._ctx.setTransform(d,0,0,d,0,0);},_animate:function(t){if(!this._running)return;var dt=Math.min(.08,Math.max(.005,(t-this._last)/1000));this._last=t;var ctx=this._ctx,s=this._map.getSize();ctx.clearRect(0,0,s.x,s.y);ctx.lineCap='round';particles.forEach(function(p){p.distance+=dt*180;if(p.distance>p.route.length+220)p.distance=-Math.random()*220;var trail=[];for(var d=p.distance-220;d<=p.distance;d+=10){var ll=pointAt(p.route,d);if(ll)trail.push(ll);}if(trail.length<2)return;ctx.beginPath();trail.forEach(function(ll,k){var pt=map.latLngToLayerPoint(ll),px=pt.x-this._topLeft.x,py=pt.y-this._topLeft.y;if(k===0)ctx.moveTo(px,py);else ctx.lineTo(px,py);},this);ctx.strokeStyle='rgba(8,48,107,0.78)';ctx.lineWidth=1.45;ctx.stroke();var head=trail[trail.length-1],hp=map.latLngToLayerPoint(head);ctx.fillStyle='rgba(44,127,184,0.98)';ctx.beginPath();ctx.arc(hp.x-this._topLeft.x,hp.y-this._topLeft.y,2.3,0,Math.PI*2);ctx.fill();},this);this._frame=window.requestAnimationFrame(this._animate.bind(this));}});",
    "var flowTrails=new FlowTrails();flowTrails.addTo(map);",
    "L.control.layers({}, {'Flowing particle trails':flowTrails},{collapsed:false}).addTo(map);",
    "function panel(position,text){var c=L.control({position:position});c.onAdd=function(){var d=L.DomUtil.create('div');d.style.background='rgba(255,255,255,.92)';d.style.padding='6px 8px';d.style.font='12px/1.3 Arial';d.style.whiteSpace='pre-line';d.style.width='min(240px,calc(100vw - 44px))';d.style.maxWidth='240px';d.style.maxHeight='min(135px,calc(100vh - 150px))';d.style.overflow='auto';d.style.boxSizing='border-box';d.style.marginBottom='24px';d.textContent=text;return d;};c.addTo(map);}",
    "function addFullscreen(){var c=L.control({position:'topleft'});c.onAdd=function(){var b=L.DomUtil.create('button');b.type='button';b.title='Toggle full screen';b.textContent='⛶';b.style.cssText='width:30px;height:30px;background:#fff;border:0;font-size:19px;cursor:pointer';L.DomEvent.disableClickPropagation(b);L.DomEvent.on(b,'click',function(){if(!document.fullscreenElement&&el.requestFullscreen){el.requestFullscreen();}else if(document.exitFullscreen){document.exitFullscreen();}});return b;};c.addTo(map);document.addEventListener('fullscreenchange',function(){window.setTimeout(function(){map.invalidateSize();},80);});}addFullscreen();",
    "panel('bottomleft',data.label);panel('bottomright',data.debug);",
    "window.setTimeout(function(){map.invalidateSize();map.fitBounds(bounds,{padding:[4,4]});map.setMaxBounds(maxBounds);},250);",
    "}"
  )
  option2_widget <- leaflet::leaflet(
    width = "100vw", height = "100vh",
    options = leaflet::leafletOptions(minZoom = 12, maxZoom = 18, maxBoundsViscosity = 1)
  )
  option2_widget <- leaflet::addProviderTiles(option2_widget, "OpenStreetMap.Mapnik", group = "OpenStreetMap")
  option2_widget <- leaflet::addProviderTiles(option2_widget, "Esri.WorldImagery", group = "Satellite imagery")
  option2_widget <- leaflet::setMaxBounds(option2_widget,
                                           current_max_bounds[[1]][2], current_max_bounds[[1]][1],
                                           current_max_bounds[[2]][2], current_max_bounds[[2]][1])
  option2_widget <- leaflet::fitBounds(option2_widget,
                                       current_bounds[[1]][2], current_bounds[[1]][1],
                                       current_bounds[[2]][2], current_bounds[[2]][1])
  option2_widget <- leaflet::addPolylines(option2_widget,
                                           lng = model_outline_ll_xy[, 1], lat = model_outline_ll_xy[, 2],
                                           color = "black", weight = 2, group = "FUNWAVE grid")
  option2_widget <- leaflet::addPolylines(option2_widget,
                                           lng = source_line_ll_xy[, 1], lat = source_line_ll_xy[, 2],
                                           color = "dodgerblue3", weight = 2, dashArray = "6 6",
                                           group = "Internal wavemaker")
  option2_widget <- leaflet::addCircleMarkers(option2_widget,
                                               lng = crds(buoy_ll)[1, 1], lat = crds(buoy_ll)[1, 2],
                                               radius = 5, color = "red", fillOpacity = 1, group = "Buoy")
  for (route in streamlines_ll) {
    option2_widget <- leaflet::addPolylines(option2_widget,
                                             lng = route[, 1], lat = route[, 2],
                                             color = "#2c7fb8", weight = 1, opacity = 0.28,
                                             group = "Mean-flow paths")
  }
  option2_widget <- leaflet::addLayersControl(
    option2_widget,
    baseGroups = c("OpenStreetMap", "Satellite imagery"),
    overlayGroups = c("Mean-flow paths", "FUNWAVE grid", "Internal wavemaker", "Buoy"),
    options = leaflet::layersControlOptions(collapsed = FALSE)
  )
  option2_widget <- leaflet::hideGroup(option2_widget, "OpenStreetMap")
  option2_widget <- leaflet::showGroup(option2_widget, "Satellite imagery")
  option2_widget <- htmlwidgets::onRender(option2_widget, option2_js, data = option2_data)
  current_particle_page_status <- tryCatch({
    htmlwidgets::saveWidget(option2_widget, current_particle_page_path, selfcontained = TRUE)
    if (!file.exists(current_particle_page_path) || file.info(current_particle_page_path)$size < 5000) {
      stop("The particle page was not written or is unexpectedly small.")
    }
    paste0("Published with ", length(particle_routes), " animated current paths.")
  }, error = function(e) {
    paste0("Particle page write failed: ", conditionMessage(e))
  })
}

if (!file.exists(current_particle_page_path)) {
  fallback_text <- paste(
    "The flow-particle page could not be animated for this run.",
    paste0("Umean files found: ", length(u_paths)),
    paste0("Vmean files found: ", length(v_paths)),
    paste0("Paired fields: ", length(common_current_numbers)),
    paste0("Status: ", current_particle_page_status),
    sep = "\n"
  )
  fallback_saved <- tryCatch({
    htmltools::save_html(
      htmltools::tags$html(
        htmltools::tags$head(htmltools::tags$title("Port Fairy current-flow diagnostics")),
        htmltools::tags$body(
          style = "font-family: system-ui, sans-serif; max-width: 58rem; margin: 3rem auto; line-height: 1.45;",
          htmltools::tags$h1("Port Fairy current-flow diagnostics"),
          htmltools::tags$p("No animated paths were available for this model run."),
          htmltools::tags$pre(fallback_text)
        )
      ),
      file = current_particle_page_path
    )
    TRUE
  }, error = function(e) {
    current_particle_page_status <<- paste0(current_particle_page_status,
                                             " Fallback page write failed: ", conditionMessage(e))
    FALSE
  })
  if (fallback_saved) {
    current_particle_page_status <- paste0(current_particle_page_status,
                                           " A diagnostic page was published instead.")
  }
}

# -------------------------------------------------------------------------
# Additional Earth-style, time-mean current animation
#
# This is deliberately additional to the retained native-grid particle page
# above. It uses a display-field interpolation and a particle
# ageing/trail renderer rather than pre-traced routes, while still sourcing
# the same FUNWAVE Umean/Vmean outputs over T_INTV_mean.
#
# The browser animator below adapts the particle ageing, vector interpolation,
# destination-in trail fade, and per-frame evolve/draw approach in earth.js
# from https://github.com/cambecc/earth (MIT License).
# Copyright (c) 2014 Cameron Beccario.  The complete MIT notice is embedded
# in the generated page data so it accompanies the adapted implementation.
earth_flow_page_path <- file.path(site_assets_dir, "port-fairy-earth-nullschool-flow.html")
earth_flow_page_url <- "assets/port-fairy-earth-nullschool-flow.html"
earth_flow_page_status <- "No paired Umean/Vmean output was produced."
earth_u_paths <- list.files(results_dir, pattern = "^Umean_[0-9]{5}$",
                            ignore.case = TRUE, full.names = TRUE)
earth_v_paths <- list.files(results_dir, pattern = "^Vmean_[0-9]{5}$",
                            ignore.case = TRUE, full.names = TRUE)
earth_file_number <- function(paths) {
  as.integer(sub("^[[:alpha:]]+_", "", basename(paths)))
}
earth_u_numbers <- earth_file_number(earth_u_paths)
earth_v_numbers <- earth_file_number(earth_v_paths)
earth_common_numbers <- sort(intersect(earth_u_numbers, earth_v_numbers))
earth_selected_number <- NA_integer_
earth_selected_files <- "none"
earth_peak_speed <- NA_real_
earth_min_speed <- NA_real_
earth_display_grid <- "n/a"
earth_particle_count <- 0L
earth_speed_colour_cap_m_s <- 0.20
earth_speed_colours <- c(
  "#000078", "#0000BB", "#0000FF", "#0C3AFC", "#1975FA", "#25B0F7",
  "#32EBF5", "#73EEE6", "#B4F1D7", "#F5F5C8", "#F5F564", "#F5E032",
  "#F6CC00", "#F7A300", "#F97A00", "#FB5100", "#FF0000", "#DA0006",
  "#B5000D", "#910014", "#7B003D", "#660066", "#88008C", "#AB00B2",
  "#CD00D8", "#F000FF", "#CAA6DE", "#DBA7DD", "#EDA9D2"
)

earth_mit_notice <- paste(
  "The MIT License (MIT)",
  "Copyright (c) 2014 Cameron Beccario",
  "Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files (the Software), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished to do so, subject to the following conditions:",
  "The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.",
  "THE SOFTWARE IS PROVIDED AS IS, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.",
  sep = "\n\n"
)

matrix_as_rows <- function(z) {
  unname(lapply(seq_len(nrow(z)), function(k) as.numeric(z[k, ])))
}

earth_west <- min(model_outline_ll_xy[, 1]); earth_east <- max(model_outline_ll_xy[, 1])
earth_south <- min(model_outline_ll_xy[, 2]); earth_north <- max(model_outline_ll_xy[, 2])
earth_x_pad <- (earth_east - earth_west) * 0.04
earth_y_pad <- (earth_north - earth_south) * 0.04
earth_bounds <- list(
  c(earth_south - earth_y_pad, earth_west - earth_x_pad),
  c(earth_north + earth_y_pad, earth_east + earth_x_pad)
)
earth_max_x_pad <- (earth_east - earth_west) * 0.50
earth_max_y_pad <- (earth_north - earth_south) * 0.50
earth_max_bounds <- list(
  c(earth_south - earth_max_y_pad, earth_west - earth_max_x_pad),
  c(earth_north + earth_max_y_pad, earth_east + earth_max_x_pad)
)
earth_outline_latlng <- lapply(seq_len(nrow(model_outline_ll_xy)), function(k) {
  unname(as.numeric(model_outline_ll_xy[k, c(2, 1)]))
})
earth_source_latlng <- lapply(seq_len(nrow(source_line_ll_xy)), function(k) {
  unname(as.numeric(source_line_ll_xy[k, c(2, 1)]))
})
earth_buoy_latlng <- unname(as.numeric(c(crds(buoy_ll)[1, 2], crds(buoy_ll)[1, 1])))

if (length(earth_common_numbers)) {
  earth_selected_number <- tail(earth_common_numbers, 1)
  earth_selected_files <- paste0("Umean_", sprintf("%05d", earth_selected_number),
                                 ", Vmean_", sprintf("%05d", earth_selected_number))
  earth_u_path <- earth_u_paths[match(earth_selected_number, earth_u_numbers)]
  earth_v_path <- earth_v_paths[match(earth_selected_number, earth_v_numbers)]
  earth_u_native <- apply_current_display_mask_model(
    read_text_model(earth_u_path, basename(earth_u_path))
  )
  earth_v_native <- apply_current_display_mask_model(
    read_text_model(earth_v_path, basename(earth_v_path))
  )
  earth_smoothed <- smooth_mean_current_components(earth_u_native, earth_v_native)
  earth_u_native <- earth_smoothed$u
  earth_v_native <- earth_smoothed$v
  # This is a depth-integrated mean-current momentum-flux proxy, h |U| U.
  # It deliberately omits density and wave radiation stress, which cannot be
  # reconstructed from a single Umean/Vmean record. Its direction is therefore
  # the mean-current direction, while depth and speed reweight its magnitude.
  earth_depth_native <- max_water_depth_model
  earth_depth_native[!current_display_mask_model] <- NA_real_
  earth_native_speed <- sqrt(earth_u_native^2 + earth_v_native^2)
  earth_flux_u_native <- earth_depth_native * earth_native_speed * earth_u_native
  earth_flux_v_native <- earth_depth_native * earth_native_speed * earth_v_native

  # Use the same compact public x/y resolution as the map products. The
  # current field remains a model-derived display; station files remain exact
  # native grid-point outputs with no interpolation.
  earth_u_public_om <- apply_current_display_mask_rotated(
    make_public_output_rotated(
      model_to_rotated_raster(earth_u_native, basename(earth_u_path))
    )
  )
  earth_v_public_om <- apply_current_display_mask_rotated(
    make_public_output_rotated(
      model_to_rotated_raster(earth_v_native, basename(earth_v_path))
    )
  )
  earth_u_public <- rotated_raster_to_model_xy(earth_u_public_om)
  earth_v_public <- rotated_raster_to_model_xy(earth_v_public_om)
  earth_valid <- is.finite(earth_u_public) & is.finite(earth_v_public)
  earth_u_public[!earth_valid] <- NA_real_
  earth_v_public[!earth_valid] <- NA_real_
  earth_flux_u_public_om <- apply_current_display_mask_rotated(
    make_public_output_rotated(
      model_to_rotated_raster(earth_flux_u_native, "mean-current momentum-flux proxy x")
    )
  )
  earth_flux_v_public_om <- apply_current_display_mask_rotated(
    make_public_output_rotated(
      model_to_rotated_raster(earth_flux_v_native, "mean-current momentum-flux proxy y")
    )
  )
  earth_flux_u_public <- rotated_raster_to_model_xy(earth_flux_u_public_om)
  earth_flux_v_public <- rotated_raster_to_model_xy(earth_flux_v_public_om)
  earth_flux_valid <- is.finite(earth_flux_u_public) & is.finite(earth_flux_v_public)
  earth_flux_u_public[!earth_flux_valid] <- NA_real_
  earth_flux_v_public[!earth_flux_valid] <- NA_real_
  earth_speed <- sqrt(earth_u_public^2 + earth_v_public^2)
  earth_valid_speed <- earth_speed[is.finite(earth_speed) & earth_speed > 0]
  earth_peak_speed <- if (length(earth_valid_speed)) max(earth_valid_speed) else NA_real_
  earth_flux_speed <- sqrt(earth_flux_u_public^2 + earth_flux_v_public^2)
  earth_valid_flux_speed <- earth_flux_speed[is.finite(earth_flux_speed) & earth_flux_speed > 0]
  earth_flux_colour_cap <- if (length(earth_valid_flux_speed)) {
    max(1e-6, as.numeric(stats::quantile(earth_valid_flux_speed, 0.95, na.rm = TRUE)))
  } else {
    1e-6
  }
  earth_flux_min_speed <- if (length(earth_valid_flux_speed)) {
    max(1e-7, 0.01 * as.numeric(stats::quantile(earth_valid_flux_speed, 0.10, na.rm = TRUE)))
  } else {
    Inf
  }
  earth_min_speed <- if (length(earth_valid_speed)) {
    max(0.0005, 0.01 * as.numeric(stats::quantile(earth_valid_speed, 0.10, na.rm = TRUE)))
  } else {
    Inf
  }
  earth_display_dx_m <- grid$dx_m[1] * public_output_factor_x
  earth_display_dy_m <- grid$dy_m[1] * public_output_factor_y
  earth_display_grid <- paste0(metres_label(earth_display_dx_m), " m x ",
                               metres_label(earth_display_dy_m), " m")

  # A local Oblique Mercator grid is effectively affine across this small
  # domain. Passing its four geographic corners lets the browser render the
  # vector field in WGS84 while interpolation and particle advection remain
  # in FUNWAVE's x/y coordinate system.
  earth_native_corners <- rbind(
    c(0, 0),
    c((grid$Mglob[1] - 1) * grid$dx_m[1], 0),
    c((grid$Mglob[1] - 1) * grid$dx_m[1], (grid$Nglob[1] - 1) * grid$dy_m[1]),
    c(0, (grid$Nglob[1] - 1) * grid$dy_m[1])
  )
  earth_corner_ll <- model_xy_to_wgs84(earth_native_corners)
  earth_corner_latlng <- lapply(seq_len(nrow(earth_corner_ll)), function(k) {
    unname(as.numeric(earth_corner_ll[k, c(2, 1)]))
  })
  # This uses a canvas, so a dense, domain-wide set of trajectories remains
  # responsive.  The particles are seeded throughout valid display cells.
  earth_particle_count <- 5400L
  earth_data <- list(
    bounds = earth_bounds,
    max_bounds = earth_max_bounds,
    outline = earth_outline_latlng,
    source = earth_source_latlng,
    buoy = earth_buoy_latlng,
    corners = earth_corner_latlng,
    fields = list(
      current = list(
        u = matrix_as_rows(earth_u_public), v = matrix_as_rows(earth_v_public),
        min_speed = earth_min_speed, colour_cap = earth_speed_colour_cap_m_s,
        label = "Mean-current speed (m/s)"
      ),
      momentum_flux = list(
        u = matrix_as_rows(earth_flux_u_public), v = matrix_as_rows(earth_flux_v_public),
        min_speed = earth_flux_min_speed, colour_cap = earth_flux_colour_cap,
        label = "Depth-integrated mean-current momentum-flux proxy (m3/s2)"
      )
    ),
    nx = nrow(earth_u_public), ny = ncol(earth_u_public),
    dx_m = earth_display_dx_m, dy_m = earth_display_dy_m,
    min_speed_m_s = earth_min_speed,
    max_speed_m_s = ifelse(is.finite(earth_peak_speed), earth_peak_speed, 0),
    particle_count = earth_particle_count,
    speed_colour_cap_m_s = earth_speed_colour_cap_m_s,
    speed_colours = unname(earth_speed_colours),
    model_time_label = paste0("Latest smoothed FUNWAVE Umean/Vmean output; ",
                              metres_label(grid$mean_wave_interval_s[1]),
                              " s averaging interval; ", current_smoothing_label,
                              ". Choose one field in the layer control."),
    credit = "Particle animator adapted from Cameron Beccario's earth (MIT).",
    mit_notice = earth_mit_notice
  )
  earth_js <- paste0(
    "function(el, x, data) {",
    "/* Adapted from earth.js in cambecc/earth. Copyright (c) 2014 Cameron Beccario; MIT notice supplied in data.mit_notice. */",
    "var map=this;if(!map||typeof map.getPanes!=='function')map=this.map||this._map;",
    "if(!map||typeof map.getPanes!=='function'){console.error('Leaflet map was not available for Earth-style particles');return;}",
    "document.documentElement.style.height='100%';document.body.style.margin='0';document.body.style.height='100%';el.style.width='100vw';el.style.height='100vh';",
    "var bounds=L.latLngBounds(data.bounds),maxBounds=L.latLngBounds(data.max_bounds||data.bounds),fields=data.fields||{},activeField='current',u=(fields.current||{}).u||[],v=(fields.current||{}).v||[],nx=data.nx,ny=data.ny;",
    "function finite(z){return typeof z==='number'&&isFinite(z);}",
    "function bilerp(a,b,c,d,fx,fy){return a*(1-fx)*(1-fy)+b*fx*(1-fy)+c*(1-fx)*fy+d*fx*fy;}",
    "function vectorAt(px,py){if(!(px>=0&&py>=0&&px<nx-1&&py<ny-1))return null;var i=Math.floor(px),j=Math.floor(py),fx=px-i,fy=py-j;var u00=u[i][j],u10=u[i+1][j],u01=u[i][j+1],u11=u[i+1][j+1],v00=v[i][j],v10=v[i+1][j],v01=v[i][j+1],v11=v[i+1][j+1];if(!finite(u00)||!finite(u10)||!finite(u01)||!finite(u11)||!finite(v00)||!finite(v10)||!finite(v01)||!finite(v11))return null;var ux=bilerp(u00,u10,u01,u11,fx,fy),vy=bilerp(v00,v10,v01,v11,fx,fy);return {u:ux,v:vy,m:Math.sqrt(ux*ux+vy*vy)};}",
    "function modelToLatLng(px,py){var fx=Math.max(0,Math.min(1,px/(nx-1))),fy=Math.max(0,Math.min(1,py/(ny-1))),c=data.corners;return [bilerp(c[0][0],c[1][0],c[3][0],c[2][0],fx,fy),bilerp(c[0][1],c[1][1],c[3][1],c[2][1],fx,fy)];}",
    "var osm=L.tileLayer('https://{s}.tile.openstreetmap.org/{z}/{x}/{y}.png',{attribution:'&copy; OpenStreetMap contributors'});",
    "var satellite=L.tileLayer('https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}',{attribution:'Tiles &copy; Esri'}).addTo(map);if(map.hasLayer(osm))map.removeLayer(osm);if(!map.hasLayer(satellite))satellite.addTo(map);",
    "var grid=L.polyline(data.outline,{color:'black',weight:2}),paddle=L.polyline(data.source,{color:'dodgerblue',weight:2,dashArray:'6 6'}),buoy=L.circleMarker(data.buoy,{radius:5,color:'red',fillOpacity:1});",
    "grid.addTo(map);paddle.addTo(map);buoy.addTo(map);map.fitBounds(bounds,{padding:[4,4]});map.setMaxBounds(maxBounds);",
    "function panel(position,text){var c=L.control({position:position});c.onAdd=function(){var d=L.DomUtil.create('div');d.style.background='rgba(255,255,255,.92)';d.style.padding='6px 8px';d.style.font='12px/1.3 Arial';d.style.whiteSpace='pre-line';d.style.width='min(250px,calc(100vw - 44px))';d.style.maxWidth='250px';d.style.maxHeight='min(110px,calc(100vh - 150px))';d.style.overflow='auto';d.style.boxSizing='border-box';d.style.marginBottom='24px';d.textContent=text;return d;};c.addTo(map);return c;}",
    "panel('bottomleft',data.model_time_label+'\\nDisplay grid: '+data.dx_m+' m x '+data.dy_m+' m\\nThe momentum option is h |U| U, a mean-current proxy rather than wave radiation stress.');map.attributionControl.addAttribution(data.credit);",
    "var MAX_PARTICLE_AGE=160,DISPLAY_TIME_SCALE=24000,particleCount=Math.max(450,Number(data.particle_count)||5400);",
    "var EarthFlowLayer=L.Layer.extend({initialize:function(){this._running=false;this._particles=[];},onAdd:function(m){this._map=m;this._canvas=L.DomUtil.create('canvas','leaflet-zoom-animated',m.getPanes().overlayPane);this._canvas.style.pointerEvents='none';this._canvas.style.zIndex=650;this._resize();m.on('resize moveend zoomend',this._resize,this);for(var k=0;k<particleCount;k++)this._particles.push(this._randomize({age:Math.random()*MAX_PARTICLE_AGE},k));this._running=true;this._last=performance.now();this._frame=window.requestAnimationFrame(this._animate.bind(this));},onRemove:function(m){this._running=false;window.cancelAnimationFrame(this._frame);m.off('resize moveend zoomend',this._resize,this);L.DomUtil.remove(this._canvas);this._map=null;},_resize:function(){if(!this._map||!this._canvas)return;var s=this._map.getSize(),d=window.devicePixelRatio||1;this._canvas.width=Math.max(1,Math.round(s.x*d));this._canvas.height=Math.max(1,Math.round(s.y*d));this._canvas.style.width=s.x+'px';this._canvas.style.height=s.y+'px';this._topLeft=this._map.containerPointToLayerPoint([0,0]);L.DomUtil.setPosition(this._canvas,this._topLeft);this._ctx=this._canvas.getContext('2d');this._ctx.setTransform(d,0,0,d,0,0);},_randomize:function(p,slot){var initial=Number.isInteger(slot),side=Math.ceil(Math.sqrt(particleCount));for(var n=0;n<1600;n++){if(initial){var q=(slot+n*997)%particleCount,col=q%side,row=Math.floor(q/side);p.x=Math.max(0,Math.min(nx-1,((col+.5+.25*(Math.random()-.5))/side)*(nx-1)));p.y=Math.max(0,Math.min(ny-1,((row+.5+.25*(Math.random()-.5))/side)*(ny-1)));}else{p.x=Math.random()*(nx-1);p.y=Math.random()*(ny-1);}var w=vectorAt(p.x,p.y);if(w&&w.m>=data.min_speed_m_s){p.age=Math.random()*MAX_PARTICLE_AGE;return p;}}p.age=MAX_PARTICLE_AGE+1;return p;},_colour:function(m){var q=Math.max(0,Math.min(1,m/Math.max(data.max_speed_m_s,1e-6)));if(q<.25)return '#ffffcc';if(q<.5)return '#a1dab4';if(q<.75)return '#41b6c4';return '#225ea8';},_animate:function(now){if(!this._running)return;var dt=Math.min(.05,Math.max(.001,(now-this._last)/1000));this._last=now;var ctx=this._ctx,s=this._map.getSize(),self=this;var previous=ctx.globalCompositeOperation;ctx.globalCompositeOperation='destination-in';ctx.fillStyle='rgba(0,0,0,0.96)';ctx.fillRect(0,0,s.x,s.y);ctx.globalCompositeOperation=previous;ctx.lineWidth=1.2;ctx.lineCap='round';this._particles.forEach(function(p){if(p.age>MAX_PARTICLE_AGE)self._randomize(p);var w=vectorAt(p.x,p.y);if(!w||w.m<data.min_speed_m_s){p.age=MAX_PARTICLE_AGE+1;return;}var stepM=Math.max(1.5,Math.min(18,w.m*dt*DISPLAY_TIME_SCALE)),mx=p.x+0.5*stepM*w.u/w.m/data.dx_m,my=p.y+0.5*stepM*w.v/w.m/data.dy_m,mid=vectorAt(mx,my)||w,xt=p.x+stepM*mid.u/mid.m/data.dx_m,yt=p.y+stepM*mid.v/mid.m/data.dy_m,next=vectorAt(xt,yt);if(!next||next.m<data.min_speed_m_s){p.age=MAX_PARTICLE_AGE+1;return;}var a=modelToLatLng(p.x,p.y),b=modelToLatLng(xt,yt),pa=map.latLngToLayerPoint(a),pb=map.latLngToLayerPoint(b);ctx.beginPath();ctx.strokeStyle=self._colour(w.m);ctx.moveTo(pa.x-self._topLeft.x,pa.y-self._topLeft.y);ctx.lineTo(pb.x-self._topLeft.x,pb.y-self._topLeft.y);ctx.stroke();p.x=xt;p.y=yt;p.age+=dt*60;});this._frame=window.requestAnimationFrame(this._animate.bind(this));}});",
    "EarthFlowLayer.prototype._colour=function(m){var colours=data.speed_colours||['#EDA9D2','#000078'],cap=Math.max(1e-6,Number(data.speed_colour_cap_m_s)||0.20),q=Math.max(0,Math.min(1,m/cap)),index=Math.min(colours.length-1,Math.round(q*(colours.length-1)));return colours[index];};",
    "var earthFlow=new EarthFlowLayer();earthFlow.addTo(map);",
    "function addFullscreen(){var c=L.control({position:'topleft'});c.onAdd=function(){var b=L.DomUtil.create('button');b.type='button';b.title='Toggle full screen';b.textContent='⛶';b.style.cssText='width:30px;height:30px;background:#fff;border:0;font-size:19px;cursor:pointer';L.DomEvent.disableClickPropagation(b);L.DomEvent.on(b,'click',function(){if(!document.fullscreenElement&&el.requestFullscreen){el.requestFullscreen();}else if(document.exitFullscreen){document.exitFullscreen();}});return b;};c.addTo(map);document.addEventListener('fullscreenchange',function(){window.setTimeout(function(){map.invalidateSize();},80);});}addFullscreen();",
    "var speedLegend=L.control({position:'bottomright'});speedLegend.onAdd=function(){var d=L.DomUtil.create('div');d.style.background='rgba(255,255,255,.92)';d.style.padding='6px 8px';d.style.font='12px Arial';d.style.lineHeight='1.25';d.style.marginBottom='24px';d.style.width='min(220px,calc(100vw - 44px))';d.style.maxWidth='220px';d.style.maxHeight='min(128px,calc(100vh - 150px))';d.style.overflow='auto';d.style.whiteSpace='normal';d.style.overflowWrap='anywhere';d.style.boxSizing='border-box';var colours=data.speed_colours||[];var swatches='';for(var i=0;i<colours.length;i++){swatches+='<i style=\"display:inline-block;width:7px;height:14px;background:'+colours[i]+'\"></i>';}d.innerHTML='<b class=\"pf-field-legend\">Mean-current speed (m/s)</b><br><span style=\"float:left\">0</span><span class=\"pf-field-cap\" style=\"float:right\">0.20+ capped</span><br><span style=\"display:flex;clear:both\">'+swatches+'</span>';return d;};speedLegend.addTo(map);",
    "function chooseField(name){var f=fields[name];if(!f)return;activeField=name;u=f.u||[];v=f.v||[];data.min_speed_m_s=Number(f.min_speed)||0;data.speed_colour_cap_m_s=Math.max(1e-6,Number(f.colour_cap)||1e-6);var legend=speedLegend.getContainer();if(legend){var title=legend.querySelector('.pf-field-legend'),cap=legend.querySelector('.pf-field-cap');if(title)title.textContent=f.label||name;if(cap)cap.textContent=data.speed_colour_cap_m_s.toFixed(name==='current'?2:3)+'+ capped';}earthFlow._particles=[];for(var k=0;k<particleCount;k++)earthFlow._particles.push(earthFlow._randomize({age:Math.random()*MAX_PARTICLE_AGE},k));}",
    "var fieldControl=L.control({position:'topright'});fieldControl.onAdd=function(){var d=L.DomUtil.create('div');d.style.background='rgba(255,255,255,.92)';d.style.padding='6px 8px';d.style.font='12px Arial';d.style.width='min(220px,calc(100vw - 44px))';d.style.maxWidth='220px';d.style.boxSizing='border-box';d.style.whiteSpace='normal';d.innerHTML='<b>Animation field</b><br><select><option value=\"current\">Mean current</option><option value=\"momentum_flux\">Momentum-flux proxy</option></select>';var s=d.querySelector('select');s.style.maxWidth='100%';L.DomEvent.disableClickPropagation(d);L.DomEvent.on(s,'change',function(){chooseField(s.value);});return d;};fieldControl.addTo(map);",
    "L.control.layers({'Satellite imagery':satellite,'OpenStreetMap':osm},{'Earth-style particles':earthFlow,'FUNWAVE grid':grid,'Internal wavemaker':paddle,'Buoy':buoy},{collapsed:false}).addTo(map);",
    "window.setTimeout(function(){map.invalidateSize();map.fitBounds(bounds,{padding:[4,4]});map.setMaxBounds(maxBounds);},250);",
    "}"
  )
  earth_widget <- leaflet::leaflet(
    width = "100vw", height = "100vh",
    options = leaflet::leafletOptions(minZoom = 12, maxZoom = 18, maxBoundsViscosity = 1)
  )
  earth_widget <- htmlwidgets::onRender(earth_widget, earth_js, data = earth_data)
  earth_flow_page_status <- tryCatch({
    htmlwidgets::saveWidget(earth_widget, earth_flow_page_path, selfcontained = TRUE)
    if (!file.exists(earth_flow_page_path) || file.info(earth_flow_page_path)$size < 5000) {
      stop("The Earth-style mean-current page was not written or is unexpectedly small.")
    }
    paste0("Published from Umean/Vmean file ", sprintf("%05d", earth_selected_number),
           "; peak mean speed ", format(round(earth_peak_speed, 3), nsmall = 3), " m/s.")
  }, error = function(e) {
    paste0("Earth-style flow page write failed: ", conditionMessage(e))
  })
}

if (!file.exists(earth_flow_page_path)) {
  earth_fallback <- paste(
    "The Earth-style mean-current particle page could not be animated for this run.",
    paste0("Umean files found: ", length(earth_u_paths)),
    paste0("Vmean files found: ", length(earth_v_paths)),
    paste0("Paired mean-current fields: ", length(earth_common_numbers)),
    paste0("Status: ", earth_flow_page_status),
    sep = "\n"
  )
  earth_fallback_saved <- tryCatch({
    htmltools::save_html(
      htmltools::tags$html(
        htmltools::tags$head(htmltools::tags$title("Port Fairy Earth-style mean-current diagnostics")),
        htmltools::tags$body(
          style = "font-family: system-ui, sans-serif; max-width: 58rem; margin: 3rem auto; line-height: 1.45;",
          htmltools::tags$h1("Port Fairy Earth-style mean-current diagnostics"),
          htmltools::tags$p("No animated mean-current field was available for this model run."),
          htmltools::tags$pre(earth_fallback)
        )
      ),
      file = earth_flow_page_path
    )
    TRUE
  }, error = function(e) {
    earth_flow_page_status <<- paste0(earth_flow_page_status,
                                      " Fallback page write failed: ", conditionMessage(e))
    FALSE
  })
  if (earth_fallback_saved) {
    earth_flow_page_status <- paste0(earth_flow_page_status,
                                     " A diagnostic page was published instead.")
  }
}

earth_flow_debug <- data.frame(
  Diagnostic = c("Umean field files found", "Vmean field files found",
                 "Paired Umean/Vmean fields", "Selected file(s)",
                 "Mean-current display grid", "Peak mean speed (m/s)",
                 "Earth-style mean-current page"),
  Value = c(length(earth_u_paths), length(earth_v_paths), length(earth_common_numbers),
            earth_selected_files, earth_display_grid,
            ifelse(is.finite(earth_peak_speed), format(round(earth_peak_speed, 3), nsmall = 3), "n/a"),
            earth_flow_page_status),
  check.names = FALSE, stringsAsFactors = FALSE
)

# A single fixed range makes every eta frame and map comparable across days.
eta_animation_limits <- eta_display_limits

# Both animations are conventional six-second looping GIFs. Their frame
# intervals adapt to the number of saved outputs.
gif_duration_s <- 6
# eta_00000 is the zero-time initial condition, not an evolved model state.
# It is retained in eta_max but deliberately excluded from the full GIF.
full_animation_indices <- which(eta_time > 0)
full_gif_interval <- gif_duration_s / max(1, length(full_animation_indices))
final_sixth_indices <- which(eta_time >= final_sixth_start_s)
final_sixth_gif_interval <- gif_duration_s / max(1, length(final_sixth_indices))
full_animation_saved_start_label <- if (length(full_animation_indices)) {
  elapsed_clock(min(eta_time[full_animation_indices]))
} else {
  "n/a"
}
full_animation_saved_end_label <- if (length(full_animation_indices)) {
  elapsed_clock(max(eta_time[full_animation_indices]))
} else {
  "n/a"
}
final_sixth_saved_start_label <- if (length(final_sixth_indices)) {
  elapsed_clock(min(eta_time[final_sixth_indices]))
} else {
  "n/a"
}
final_sixth_saved_end_label <- if (length(final_sixth_indices)) {
  elapsed_clock(max(eta_time[final_sixth_indices]))
} else {
  "n/a"
}

# Write real GIF files before the HTML is assembled. This is more reliable than
# Knitr's animation hook, which can leave an empty figure placeholder when the
# report is moved to the Pages output directory.
site_assets_dir <- file.path(report_site_dir, "assets")
dir.create(site_assets_dir, recursive = TRUE, showWarnings = FALSE)
full_gif_url <- "assets/port-fairy-full.gif"
final_sixth_gif_url <- "assets/port-fairy-final-sixth.gif"

draw_eta_frame <- function(k) {
  r_om <- eta_public_frames[[k]]
  r_ll <- project(r_om, "EPSG:4326", method = "bilinear")
  r_ll <- apply_water_masks_wgs84(r_ll)
  plot_geographic(r_ll,
                  main = sprintf("Free-surface elevation: t = %.1f s (WGS84)", eta_time[k]),
                  col = eta_animation_palette,
                  range = eta_animation_limits, show_legend = TRUE,
                  animation = TRUE)
  add_wave_frame_annotation(r_ll, eta_time[k])
}

write_eta_gif <- function(indices, file_name, interval) {
  if (!length(indices)) return(NULL)
  gif_path <- file.path(site_assets_dir, file_name)
  gifski::save_gif(
    expr = for (k in indices) draw_eta_frame(k),
    gif_file = gif_path, width = 960, height = 720, res = 110,
    delay = interval, loop = TRUE, progress = FALSE
  )
  if (!file.exists(gif_path) || file.info(gif_path)$size < 1000) {
    stop("GIF was not written: ", gif_path)
  }
  gif_path
}

full_gif_path <- write_eta_gif(full_animation_indices, "port-fairy-full.gif",
                               full_gif_interval)
final_sixth_gif_path <- write_eta_gif(final_sixth_indices,
                                      "port-fairy-final-sixth.gif",
                                      final_sixth_gif_interval)
```

```{r report-table-style, results='asis', echo=FALSE}
cat('<style>
.pf-report-table-details { margin: 0.8rem 0 1.5rem; border: 1px solid #dfe6e9; border-radius: 0.35rem; background: #fff; }
.pf-report-table-details > summary { cursor: pointer; padding: 0.55rem 0.7rem; color: #2c3e50; font-weight: 600; background: #f6f8fa; }
.pf-report-table-details[open] > summary { border-bottom: 1px solid #dfe6e9; }
.pf-report-table { width: 100%; margin: 0.8rem 0 1.5rem; border-collapse: collapse; font-size: 0.93rem; }
.pf-report-table th { background: #2c3e50; color: #fff; text-align: left; padding: 0.55rem 0.7rem; }
.pf-report-table td { padding: 0.45rem 0.7rem; border-bottom: 1px solid #dfe6e9; vertical-align: top; }
.pf-report-table td:first-child { width: 34%; color: #2c3e50; font-weight: 600; }
.pf-report-table tbody tr:nth-child(even) { background: #f6f8fa; }
</style>')
```

## Interpretation first

This is a `r paste0(metres_label(grid$dx_m[1]), " m")`, `r simulation_label`
grid and forcing diagnostic—not a validated harbour forecast. Each raw
FUNWAVE text field is first rebuilt on the exact rotated Oblique Mercator grid,
then masked seaward of the internal wavemaker. Beach visibility is derived on
that native grid from the AHD DEM and the greatest saved positive eta, so cells
that wet above MSL are retained before any WGS84 projection. The
historical workflow publishes report maps, animations and GeoTIFFs at
`r paste0(public_grid_tag, " cells")` in that rotated CRS, then reprojects them
to WGS84 for geographic display. Current paths remain traced on native model
cells before being projected to WGS84. Interactive maps start on the local
domain but can zoom out to roughly twice its span; current-animation pages
use the full browser viewport and offer a full-screen button.

Live report: <https://julianog.github.io/run_surf_wave_forecast/>. Source:
<https://github.com/JulianOG/run_surf_wave_forecast>.

```{r forcing-status}
direction_status <- if (forcing$inward_source_component > 0) {
  "The wave vector has an inward component from the selected source edge."
} else {
  "Warning: the wave vector points out of the selected source edge. Do not interpret this run as a coastal-wave prediction."
}
report_table(transpose_summary(data.frame(
  source_edge = grid$source_edge,
  boundary_direction_statistic = forcing$boundary_direction_statistic,
  requested_model_x_bearing_deg_true = forcing$requested_model_x_bearing_deg_true,
  realised_model_x_bearing_deg_true = forcing$model_x_bearing_deg_true,
  funwave_theta_peak_deg = forcing$funwave_theta_peak_deg,
  inward_source_component = forcing$inward_source_component,
  map_water_visibility = display_water_mask_label,
  interpretation = direction_status
), digits = 2))
```

## Latest buoy forcing

The irregular `WK_IRR` source is a TMA/JONSWAP-style parameterisation made from
the newest usable IMOS Spotter observation. It uses the buoy **mean** direction
and matching mean directional spread when present (`SSWMD` and `WMDS`), falling
back to the peak values only when the mean values are unavailable. The buoy
provides integral Hs and period parameters—not a phase-resolved spectrum.

```{r forcing}
report_table(transpose_summary(forcing, digits = 2))
```

```{r historical-observed-conditions-heading, results='asis', echo=FALSE}
if (historical_observation_available) {
  cat(paste0("## Observed conditions around this ", observation_context_label, "\n\n"))
  if (observation_is_historical) {
    cat(
      "This interactive context shows QC 1/2 Port Fairy buoy Hs, Tp and wave direction, ",
      "together with hourly Portland water level converted from assumed LAT to AHD, ",
      "for seven days either side of the requested historical time. The red diamond ",
      "marks the measurement used to force FUNWAVE; it may precede the requested time ",
      "because historical forcing never uses a future observation.\n\n",
      sep = ""
    )
  } else {
    cat(
      "This interactive context shows QC 1/2 Port Fairy buoy Hs, Tp and wave direction, ",
      "together with hourly Portland water level converted from assumed LAT to AHD, ",
      "for the seven days ending at the latest available forecast forcing record. The red ",
      "diamond marks the measurement applied to FUNWAVE.\n\n",
      sep = ""
    )
  }
}
```

```{r historical-observed-conditions, echo=FALSE}
if (historical_observation_available) {
  if (file.exists(historical_observation_page_path)) {
    htmltools::tags$iframe(
      src = historical_observation_page_url, width = "100%", height = "940",
      style = "border: 1px solid #dfe6e9; display: block;"
    )
  } else {
    htmltools::tags$p(historical_observation_page_status)
  }
}
```

Each observed-variable panel carries its own Port Fairy local-time axis, so the
timing of waves and water levels can be compared without relying on a shared
bottom axis. The orange dashed line in the Portland panel is the mean of annual
hourly maxima from sufficiently complete local years in the selected UHSLC file.

```{r historical-observed-conditions-table, echo=FALSE}
if (historical_observation_available) report_table(historical_observation_summary)
```

## Latest free-surface elevation

This map is the final saved instantaneous free-surface displacement, `eta`, and
is not a wave-height map. Its layer control includes the model's positive-water-
depth bathymetry, which can be viewed alone by turning off **Model output**.

```{r eta, echo=FALSE}
if (is.null(eta)) {
  plot.new(); text(0.5, 0.5, "No eta output was produced.")
} else {
  eta_plot <- apply_output_visibility_mask_model(eta)
  eta_rotated <- model_to_rotated_raster(eta_plot, "eta")
  eta_wgs84 <- write_model_output_rasters(eta_rotated, "eta_latest")
  eta_wgs84 <- apply_water_masks_wgs84(eta_wgs84)
  writeRaster(eta_wgs84, file.path(case_dir, "eta_latest_wgs84.tif"), overwrite = TRUE)
  bathymetry_wgs84 <- project(depth_raster, "EPSG:4326", method = "bilinear")
  bathymetry_limit <- terra::global(depth_raster, "max", na.rm = TRUE)[1, 1]
  leaflet_model_map(eta_wgs84, eta_animation_palette,
                    eta_display_limits, "Free-surface\nelevation (m)",
                    bathymetry_r_ll = bathymetry_wgs84,
                    bathymetry_palette = ygnbu_palette,
                    bathymetry_limits = c(0, bathymetry_limit))
}
```

## Free-surface animations

Both GIFs use a shared colour scale and are written as standalone assets from
this Rmd, after each raw output has been reconstructed on the rotated FUNWAVE
grid and projected to WGS84. The zero-time initial condition is
intentionally omitted from the full animation. Each GIF loops for six seconds:
one represents the complete `r simulation_label` run (saved frames
`r full_animation_saved_start_label`--`r full_animation_saved_end_label`), and
the other shows the final one-sixth (configured `r final_sixth_start_label`--`r simulation_end_label`;
saved frames `r final_sixth_saved_start_label`--`r final_sixth_saved_end_label`).

The fixed −5 to 5 m YlGnBu colour range makes colours comparable among frames
and between the two animations. The legend and source/buoy key are drawn inside
the enlarged animation frame rather than in its clipped outer margin.

### Full `r simulation_label` simulation: saved `r full_animation_saved_start_label`--`r full_animation_saved_end_label`

This animation shows the development from the first evolved model output to the
last saved model time. It is a visual compression of the simulation timeline,
so the six-second playback rate is not physical wave celerity.

```{r eta-animation-full}
if (is.null(full_gif_path)) {
  plot.new(); text(0.5, 0.5, "No evolved eta output was produced.")
} else {
  htmltools::tags$img(src = full_gif_url, alt = "Full Port Fairy free-surface animation",
                       style = "max-width: 100%; height: auto; display: block;")
}
```

### Final one-sixth: saved `r final_sixth_saved_start_label`--`r final_sixth_saved_end_label`

This animation isolates the final sixth of the run, after the initial forcing
adjustment has passed. It uses the same spatial extent and colour scale as the
full simulation so features can be compared directly.

```{r eta-animation-final-sixth}
if (is.null(final_sixth_gif_path)) {
  plot.new(); text(0.5, 0.5, "No final-one-sixth eta output was produced.")
} else {
  htmltools::tags$img(src = final_sixth_gif_url,
                       alt = "Final one-sixth Port Fairy free-surface animation",
                       style = "max-width: 100%; height: auto; display: block;")
}
```

## Maximum free-surface elevation

This is the maximum positive instantaneous `eta` at each cell. It is useful for
seeing crest reach, but it is **not** a wave-height statistic.

```{r hmax-eta, echo=FALSE}
if (is.null(eta_max)) {
  plot.new(); text(0.5, 0.5, "No eta output was produced.")
} else {
  eta_max_wgs84 <- write_model_output_rasters(eta_max, "eta_max", already_public = TRUE)
  eta_max_wgs84 <- apply_water_masks_wgs84(eta_max_wgs84)
  writeRaster(eta_max_wgs84, file.path(case_dir, "eta_max_wgs84.tif"), overwrite = TRUE)
  leaflet_model_map(eta_max_wgs84, ygnbu_palette,
                    eta_max_display_limits, "Maximum free-surface\nelevation (m)")
}
```

## Peak simulated significant wave height

`WaveHeight = T` is evaluated by FUNWAVE's mean-wave diagnostic after the
configured spin-up. The map below is the largest `Hsig` value at each cell from
the saved mean-wave outputs, rather than a maximum of individual `eta` frames.
Its 0–6 m colour scale is fixed so daily and historical peaks remain comparable.

```{r hsig-max, echo=FALSE}
if (is.null(hsig_max)) {
  plot.new(); text(0.5, 0.5, "No Hsig output was produced. Check STEADY_TIME and T_INTV_mean.")
} else {
  hsig_max_wgs84 <- write_model_output_rasters(hsig_max, "hsig_max", already_public = TRUE)
  hsig_max_wgs84 <- apply_water_masks_wgs84(hsig_max_wgs84)
  writeRaster(hsig_max_wgs84, file.path(case_dir, "hsig_max_wgs84.tif"), overwrite = TRUE)
  leaflet_model_map(hsig_max_wgs84, ygnbu_palette,
                    hsig_display_limits, "Peak simulated\nsignificant wave height (m)")
}
```

## Data acknowledgement and citation

Buoy data were sourced from the **Integrated Marine Observing System (IMOS)**,
which is enabled by the National Collaborative Research Infrastructure Strategy
(NCRIS). The wave data were collected and quality controlled by the University
of Western Australia. This dataset is an output from the Catching Oz Waves
project, which received investment from the Australian Research Data Commons
(ARDC): <https://doi.org/10.47486/DP748>. ARDC is funded by NCRIS.

Suggested data citation: *Deakin University (year of data downloaded), Wave
buoys Observations -- Australia -- near real-time, downloaded from the IMOS URL
on the date of download.*

For historical reports, the same Port Fairy records are read from the public
AODN delayed-mode collection when available. The exact NetCDF filename used
for the model forcing and the source filename for each displayed observation
are retained in the report provenance/output files.

## Provenance

```{r provenance}
report_table(transpose_summary(grid, digits = 3))
```

## Buoy-to-nearshore eta time series

These are native FUNWAVE station outputs from five selected integer grid cells:
the buoy end is black-purple and the nearshore end is pale yellow on the Magma
scale. They are written every second from the model state—there is no spatial
or temporal interpolation, and the horizontal axis is model elapsed time.

```{r station-timeseries, echo=FALSE}
if (!nrow(station_timeseries)) {
  plot.new(); text(0.5, 0.5, "No FUNWAVE station output was produced.")
} else {
  htmltools::tags$iframe(
    src = "assets/port-fairy-transect-timeseries.html", width = "100%", height = "560",
    style = "border: 1px solid #dfe6e9; display: block;"
  )
}
```

## Animated current particles

This existing particle page now uses the latest FUNWAVE time-mean `Umean`/
`Vmean` field rather than phase-resolved `U`/`V` wave-orbital velocities. Its
paths are traced on native rotated grid cells with nearest-cell velocities,
then projected to WGS84. This removes the wave-period direction reversals that
made the particles wobble; animation speed remains visual only.

```{r current-particles, echo=FALSE}
htmltools::tags$p(
  htmltools::tags$a(
    "Open the full animated current-flow particle page",
    href = current_particle_page_url, target = "_blank",
    style = "font-size: 1.1rem; font-weight: 600;"
  )
)
```

## Earth-style current particles

This is an additional particle renderer inspired by the open-source
[earth.nullschool.net animator](https://github.com/cambecc/earth). It uses the
same latest `Umean`/`Vmean` field as the repaired page above, but advects and
fades particles directly in the browser to give continuous trails. The
browser's bilinear field interpolation is display-only; it does not alter
FUNWAVE outputs or the exact grid-cell transect series.

```{r earth-style-current-particles, echo=FALSE}
htmltools::tags$p(
  htmltools::tags$a(
    "Open the Earth-style mean-current particle page",
    href = earth_flow_page_url, target = "_blank",
    style = "font-size: 1.1rem; font-weight: 600;"
  )
)
report_table(earth_flow_debug)
```
