---
title: "Port Fairy near-real-time FUNWAVE test"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 11, fig.height = 6)

suppressPackageStartupMessages({
  library(terra)
  library(leaflet)
})

case_dir <- normalizePath("../port_fairy/output", mustWork = TRUE)
# The normal forecast writes support files under site/assets. Historical runs
# set PORT_FAIRY_SITE_DIR to an isolated report folder so their maps/assets do
# not overwrite the live report's equivalents.
report_site_dir <- normalizePath(
  file.path("..", Sys.getenv("PORT_FAIRY_SITE_DIR", unset = "site")),
  mustWork = FALSE
)
results_dir <- file.path(case_dir, "results")
forcing <- read.csv(file.path(case_dir, "latest_buoy_forcing.csv"),
                    check.names = FALSE)
grid <- read.csv(file.path(case_dir, "grid_metadata.csv"), check.names = FALSE)
if (nrow(forcing) != 1 || nrow(grid) != 1) {
  stop("Expected one row in forcing and grid metadata files.")
}
if (!is.finite(grid$dx_m[1]) || !is.finite(grid$dy_m[1]) ||
    !is.finite(grid$total_time_s[1]) || !is.finite(grid$plot_intv_s[1]) ||
    grid$dx_m[1] <= 0 || grid$dy_m[1] <= 0 ||
    grid$total_time_s[1] <= 0 || grid$plot_intv_s[1] <= 0) {
  stop("Grid metadata must contain positive dx_m, dy_m, total_time_s and plot_intv_s values.")
}
metres_label <- function(x) sub("\\.?0+$", "", format(x, trim = TRUE, scientific = FALSE))
public_output_dx_m <- if ("public_output_dx_m" %in% names(grid) &&
                          is.finite(grid$public_output_dx_m[1])) {
  grid$public_output_dx_m[1]
} else {
  max(10, grid$dx_m[1])
}
public_output_dy_m <- if ("public_output_dy_m" %in% names(grid) &&
                          is.finite(grid$public_output_dy_m[1])) {
  grid$public_output_dy_m[1]
} else {
  public_output_dx_m
}
public_output_factor_x <- public_output_dx_m / grid$dx_m[1]
public_output_factor_y <- public_output_dy_m / grid$dy_m[1]
if (public_output_factor_x < 1 || public_output_factor_y < 1 ||
    abs(public_output_factor_x - round(public_output_factor_x)) > 1e-8 ||
    abs(public_output_factor_y - round(public_output_factor_y)) > 1e-8) {
  stop("The requested public-output x/y resolutions must be integer multiples of the model x/y resolutions.")
}
public_output_factor_x <- as.integer(round(public_output_factor_x))
public_output_factor_y <- as.integer(round(public_output_factor_y))
grid_tag <- if ("grid_file_tag" %in% names(grid) &&
                nzchar(as.character(grid$grid_file_tag[1]))) {
  as.character(grid$grid_file_tag[1])
} else if (abs(grid$dx_m[1] - grid$dy_m[1]) < 1e-8) {
  paste0(metres_label(grid$dx_m[1]), "m")
} else {
  paste0(metres_label(grid$dx_m[1]), "m_x_", metres_label(grid$dy_m[1]), "m_y")
}
public_grid_tag <- paste0(metres_label(public_output_dx_m), "m x ",
                          metres_label(public_output_dy_m), "m")
simulation_minutes <- grid$total_time_s[1] / 60
final_sixth_start_s <- (5 / 6) * grid$total_time_s[1]
elapsed_clock <- function(elapsed_s) {
  sprintf("%02.0f:%02.0f", floor(elapsed_s / 60), round(elapsed_s %% 60))
}
simulation_label <- paste0(sprintf("%.0f", simulation_minutes), "-minute")
final_sixth_start_label <- elapsed_clock(final_sixth_start_s)
simulation_end_label <- elapsed_clock(grid$total_time_s[1])
eta_display_limits <- c(-5, 5)
eta_max_display_limits <- c(0, 5)
hsig_display_limits <- c(0, 4)
ygnbu_palette <- RColorBrewer::brewer.pal(9, "YlGnBu")
# Station 1 is the buoy/offshore end; use deep blue offshore and yellow at
# the nearshore end of the transect.
station_colours <- ygnbu_palette[c(9, 8, 6, 4, 2)]
eta_animation_palette <- grDevices::colorRampPalette(ygnbu_palette)(100)

pretty_field_name <- function(x) {
  x <- gsub("_", " ", x)
  x <- tools::toTitleCase(x)
  x <- gsub("\\bFunwave\\b", "FUNWAVE", x)
  x <- gsub("\\bHs\\b", "Hs", x)
  x <- gsub("\\bTp\\b", "Tp", x)
  x <- gsub("\\bMglob\\b", "Mglob", x)
  x <- gsub("\\bNglob\\b", "Nglob", x)
  x <- gsub("\\bDx\\b", "DX", x)
  x <- gsub("\\bDy\\b", "DY", x)
  x <- gsub("\\bWk\\b", "WK", x)
  x
}

transpose_summary <- function(x, digits = 2) {
  x <- as.data.frame(x, check.names = FALSE, stringsAsFactors = FALSE)
  if (nrow(x) != 1) stop("transpose_summary() expects a one-row data frame.")
  data.frame(
    Item = pretty_field_name(names(x)),
    Value = vapply(x[1, ], function(value) {
      if (is.numeric(value)) format(value, digits = digits, trim = TRUE, scientific = FALSE) else as.character(value)
    }, character(1)),
    check.names = FALSE
  )
}

report_table <- function(x, digits = 2) {
  knitr::kable(x, format = "html", digits = digits, escape = TRUE,
               table.attr = 'class="pf-report-table"')
}

# The buoys report UTC. Show the animation using Port Fairy's local civil time
# while retaining the model's elapsed time from that observation.
port_fairy_tz <- "Australia/Melbourne"
buoy_time_utc <- as.POSIXct(
  sub(" UTC$", "", forcing$time_utc[1]),
  format = "%Y-%m-%d %H:%M:%S", tz = "UTC"
)
if (is.na(buoy_time_utc)) stop("Could not parse the buoy observation time.")

# FUNWAVE writes text grids as Nglob rows (model y) by Mglob columns (model x).
# R's image() expects z[x, y], so transpose every model output exactly once.
as_model_xy <- function(z, label = "model field") {
  z <- as.matrix(z)
  if (all(dim(z) == c(grid$Nglob, grid$Mglob))) return(t(z))
  if (all(dim(z) == c(grid$Mglob, grid$Nglob))) return(z)
  stop(sprintf("%s has %.0f x %.0f values; expected %.0f x %.0f (y x x).",
               label, nrow(z), ncol(z), grid$Nglob, grid$Mglob))
}

read_text_model <- function(path, label = basename(path)) {
  as_model_xy(utils::read.table(path, header = FALSE), label)
}

read_field <- function(prefix, directory = results_dir) {
  candidates <- list.files(directory, pattern = paste0("^", prefix, "_[0-9]{5}$"),
                           full.names = TRUE)
  if (!length(candidates)) return(NULL)
  number <- as.integer(sub(paste0("^", prefix, "_"), "", basename(candidates)))
  read_text_model(candidates[which.max(number)], prefix)
}


read_all_fields <- function(prefix, directory = results_dir) {
  candidates <- list.files(directory, pattern = paste0("^", prefix, "_[0-9]{5}$"),
                           full.names = TRUE)
  if (!length(candidates)) return(list())
  number <- as.integer(sub(paste0("^", prefix, "_"), "", basename(candidates)))
  candidates <- candidates[order(number)]
  lapply(candidates, read_text_model, label = prefix)
}

maximum_model_field <- function(fields) {
  if (!length(fields)) return(NULL)
  out <- fields[[1]]
  if (length(fields) > 1) {
    for (k in 2:length(fields)) out <- pmax(out, fields[[k]], na.rm = TRUE)
  }
  out[!is.finite(out)] <- NA_real_
  out
}

read_static_model <- function(name) {
  path <- file.path(results_dir, name)
  if (!file.exists(path)) return(NULL)
  read_text_model(path, name)
}

model_axes <- function(z) {
  list(x = (seq_len(nrow(z)) - 1) * grid$dx_m,
       y = (seq_len(ncol(z)) - 1) * grid$dy_m)
}

plot_model <- function(z, main, col, zlim = range(z, finite = TRUE), ...) {
  a <- model_axes(z)
  image(a$x, a$y, z, asp = 1, col = col, zlim = zlim,
        xlab = "Model x: source edge to coast (m)",
        ylab = "Model y (m)", main = main, ...)
  invisible(a)
}

model_to_rotated_raster <- function(z, label = "model field") {
  z <- as.matrix(z)
  expected <- c(grid$Mglob, grid$Nglob)
  if (!identical(dim(z), expected)) {
    stop(sprintf("%s has %.0f x %.0f values; expected Mglob x Nglob = %.0f x %.0f.",
                 label, nrow(z), ncol(z), expected[1], expected[2]))
  }

  # z is [model x, model y]. Convert it back to terra's native
  # [north-to-south row, west-to-east column] layout using the exact inverse
  # of the transforms in setup_port_fairy_funwave.R.
  gis_rows <- t(z)
  if (source_is_low_x) {
    gis_rows <- gis_rows[nrow(gis_rows):1, , drop = FALSE]
  } else {
    gis_rows <- gis_rows[, ncol(gis_rows):1, drop = FALSE]
  }

  out <- rast(depth_raster)
  values(out) <- as.vector(t(gis_rows))
  out
}

symmetric_raster_limits <- function(r) {
  m <- global(abs(r), "max", na.rm = TRUE)[1, 1]
  if (!is.finite(m) || m == 0) m <- 1e-8
  c(-m, m)
}

write_model_output_rasters <- function(r_om, stem, method = "bilinear",
                                       already_public = FALSE) {
  # `make_public_output_rotated()` first masks native-grid values then
  # aggregates to the requested public resolution before any geographic
  # reprojection. This keeps a 2.5 m historical simulation numerical while
  # publishing compact 10 m maps and GeoTIFFs.
  if (!already_public) r_om <- make_public_output_rotated(r_om)
  writeRaster(r_om, file.path(case_dir, paste0(stem, "_rotated_omerc.tif")),
              overwrite = TRUE)
  r_ll <- project(r_om, "EPSG:4326", method = method)
  writeRaster(r_ll, file.path(case_dir, paste0(stem, "_wgs84.tif")),
              overwrite = TRUE)
  r_ll
}


symmetric_limits <- function(z) {
  m <- max(abs(z), na.rm = TRUE)
  if (!is.finite(m) || m == 0) m <- 1e-8
  c(-m, m)
}

positive_limits <- function(z) {
  m <- max(z, na.rm = TRUE)
  if (!is.finite(m) || m <= 0) m <- 1e-8
  c(0, m)
}

depth_raster <- rast(file.path(
  case_dir, paste0("depth_", grid_tag, "_positive_water_depth.tif")
))
elevation_raster <- rast(file.path(
  case_dir, paste0("elevation_", grid_tag, "_warped.tif")
))
depth_from_file <- read_text_model(file.path(case_dir, "depth.txt"), "depth.txt")
source_is_low_x <- identical(grid$source_edge, "minimum rotated x")

# This mask comes from the warped DEM supplied to FUNWAVE, not from the
# model's dynamic wet/dry field. It deliberately limits displayed water levels
# to DEM cells with positive water depth.
dem_water_model <- is.finite(depth_from_file) & depth_from_file > 0
dem_water_mask_rotated <- ifel(depth_raster > 0, 1, 0)
dem_water_mask_wgs84 <- project(dem_water_mask_rotated, "EPSG:4326", method = "near")

apply_dem_water_mask_model <- function(z) {
  z <- as.matrix(z)
  z[!dem_water_model] <- NA_real_
  z
}

apply_dem_water_mask_rotated <- function(r_om) {
  r_om[is.na(dem_water_mask_rotated) | dem_water_mask_rotated <= 0] <- NA_real_
  r_om
}

# The internal wavemaker emits energy in both directions. The buoy-side side
# is intentionally hidden from coastal forecast products before projection and
# again afterwards, preventing interpolation from bleeding across the paddle.
shoreward_of_paddle_model <- outer(
  (seq_len(grid$Mglob[1]) - 1) * grid$dx_m[1], seq_len(grid$Nglob[1]),
  function(x, y) x >= grid$x_wk_m[1]
)
shoreward_of_paddle_rotated <- model_to_rotated_raster(
  ifelse(shoreward_of_paddle_model, 1, 0), "shoreward-of-wavemaker mask"
)
shoreward_of_paddle_wgs84 <- project(shoreward_of_paddle_rotated, "EPSG:4326", method = "near")

apply_output_visibility_mask_model <- function(z) {
  z <- apply_dem_water_mask_model(z)
  z[!shoreward_of_paddle_model] <- NA_real_
  z
}

apply_output_visibility_mask_rotated <- function(r_om) {
  r_om <- apply_dem_water_mask_rotated(r_om)
  r_om[is.na(shoreward_of_paddle_rotated) | shoreward_of_paddle_rotated <= 0] <- NA_real_
  r_om
}

make_public_output_rotated <- function(r_om) {
  r_om <- apply_output_visibility_mask_rotated(r_om)
  if (public_output_factor_x <= 1L && public_output_factor_y <= 1L) return(r_om)

  # A public cell is retained only where all contributing native cells are
  # water and shoreward of the wavemaker. This prevents a 10 m aggregation
  # from leaking displayed values across the coast or source-side mask.
  visibility <- ifel(
    is.na(dem_water_mask_rotated) | is.na(shoreward_of_paddle_rotated) |
      dem_water_mask_rotated <= 0 | shoreward_of_paddle_rotated <= 0,
    0, 1
  )
  public_visibility <- aggregate(
    visibility, fact = c(public_output_factor_x, public_output_factor_y),
    fun = min, na.rm = TRUE
  )
  r_om <- aggregate(r_om, fact = c(public_output_factor_x, public_output_factor_y),
                     fun = mean, na.rm = TRUE)
  r_om[is.na(public_visibility) | public_visibility <= 0] <- NA_real_
  r_om
}

# Reconstruct depth.txt on the exact native Omerc raster, then derive maps
# from that raster. This is the same pathway used for FUNWAVE outputs below.
depth_from_file_rotated <- model_to_rotated_raster(depth_from_file, "depth.txt")
depth_file_minus_raster <- depth_from_file_rotated - depth_raster
max_depth_file_delta <- global(abs(depth_file_minus_raster), "max", na.rm = TRUE)[1, 1]

buoy_ll <- vect(matrix(c(forcing$buoy_lon, forcing$buoy_lat), ncol = 2),
                type = "points", crs = "EPSG:4326")
buoy_om <- project(buoy_ll, crs(depth_raster))

# Plot all map products in WGS84, but only after reconstructing them on the
# native rotated grid. The black outline is the rectangular FUNWAVE domain.
outline_from_raster <- function(r) {
  vect(rbind(
    c(xmin(r), ymin(r)), c(xmax(r), ymin(r)),
    c(xmax(r), ymax(r)), c(xmin(r), ymax(r)),
    c(xmin(r), ymin(r))
  ), type = "polygons", crs = crs(r))
}
model_outline_om <- outline_from_raster(depth_raster)
model_outline_ll_xy <- crds(project(model_outline_om, "EPSG:4326"))

# Model x is always written from the buoy-side source edge toward the coast.
# Draw the actual internal-wavemaker line on every map so the reconstructed
# fields can be checked against their physical source location.
source_x_om <- if (source_is_low_x) {
  xmin(depth_raster) + grid$x_wk_m
} else {
  xmax(depth_raster) - grid$x_wk_m
}
source_line_om <- vect(rbind(
  c(source_x_om, ymin(depth_raster)),
  c(source_x_om, ymax(depth_raster))
), type = "lines", crs = crs(depth_raster))
source_line_ll_xy <- crds(project(source_line_om, "EPSG:4326"))

plot_geographic <- function(r_ll, ..., show_legend = FALSE, range = NULL) {
  # terra::plot uses range= to fix its colour scale; zlim= is for image().
  if (is.null(range)) plot(r_ll, ...) else plot(r_ll, ..., range = range)
  lines(model_outline_ll_xy[, 1], model_outline_ll_xy[, 2],
        col = "black", lwd = 2)
  lines(source_line_ll_xy[, 1], source_line_ll_xy[, 2],
        col = "dodgerblue3", lwd = 2, lty = 2)
  points(crds(buoy_ll)[1, 1], crds(buoy_ll)[1, 2],
         pch = 16, col = "red", cex = 1.25)
  if (show_legend) {
    legend("bottomleft",
           legend = c("FUNWAVE grid", "internal wavemaker", "buoy"),
           col = c("black", "dodgerblue3", "red"), lwd = c(2, 2, NA),
           lty = c(1, 2, NA), pch = c(NA, NA, 16),
           bty = "n", cex = 0.75)
  }
}

local_map_bounds <- function(r_ll, padding = 0.04) {
  e <- ext(r_ll)
  x_pad <- (xmax(e) - xmin(e)) * padding
  y_pad <- (ymax(e) - ymin(e)) * padding
  c(west = xmin(e) - x_pad, south = ymin(e) - y_pad,
    east = xmax(e) + x_pad, north = ymax(e) + y_pad)
}

apply_water_masks_wgs84 <- function(r_ll) {
  dem_mask <- dem_water_mask_wgs84
  if (!compareGeom(dem_mask, r_ll, stopOnError = FALSE)) {
    dem_mask <- resample(dem_mask, r_ll, method = "near")
  }
  r_ll[is.na(dem_mask) | dem_mask <= 0] <- NA_real_

  paddle_mask <- shoreward_of_paddle_wgs84
  if (!compareGeom(paddle_mask, r_ll, stopOnError = FALSE)) {
    paddle_mask <- resample(paddle_mask, r_ll, method = "near")
  }
  r_ll[is.na(paddle_mask) | paddle_mask <= 0] <- NA_real_

  if (is.null(mask_wgs84)) return(r_ll)
  wet_mask <- mask_wgs84
  if (!compareGeom(wet_mask, r_ll, stopOnError = FALSE)) {
    wet_mask <- resample(wet_mask, r_ll, method = "near")
  }
  r_ll[is.na(wet_mask) | wet_mask <= 0] <- NA_real_
  r_ll
}

leaflet_model_map <- function(r_ll, palette, limits, legend_title) {
  b <- local_map_bounds(r_ll)
  pal <- colorNumeric(palette = palette, domain = limits, na.color = "transparent")
  m <- leaflet(width = "100%", height = 480,
               options = leafletOptions(minZoom = 14, maxZoom = 18,
                                        maxBoundsViscosity = 1))
  # These public sources require no project key.
  m <- addProviderTiles(m, "OpenStreetMap.Mapnik", group = "OpenStreetMap")
  m <- addProviderTiles(m, "Esri.WorldImagery", group = "Satellite imagery")
  m <- setMaxBounds(m, b[["west"]], b[["south"]], b[["east"]], b[["north"]])
  m <- fitBounds(m, b[["west"]], b[["south"]], b[["east"]], b[["north"]])
  m <- addRasterImage(m, r_ll, colors = pal, opacity = 0.88, project = FALSE,
                      group = "Model output")
  m <- addPolylines(m, lng = model_outline_ll_xy[, 1], lat = model_outline_ll_xy[, 2],
                    color = "black", weight = 2, group = "FUNWAVE grid")
  m <- addPolylines(m, lng = source_line_ll_xy[, 1], lat = source_line_ll_xy[, 2],
                    color = "dodgerblue3", weight = 2, dashArray = "6 6",
                    group = "Internal wavemaker")
  m <- addCircleMarkers(m, lng = crds(buoy_ll)[1, 1], lat = crds(buoy_ll)[1, 2],
                        radius = 5, color = "red", fillOpacity = 1,
                        label = "Port Fairy wave buoy", group = "Buoy")
  if (exists("station_metadata", inherits = TRUE) && nrow(station_metadata)) {
    m <- addCircleMarkers(
      m, lng = station_metadata$lon, lat = station_metadata$lat,
      radius = 4, color = station_colours[seq_len(nrow(station_metadata))],
      fillOpacity = 1, label = station_metadata$station_id,
      group = "Transect stations"
    )
  }
  m <- addLayersControl(
    m,
    baseGroups = c("OpenStreetMap", "Satellite imagery"),
    overlayGroups = c("Model output", "FUNWAVE grid", "Internal wavemaker", "Buoy", "Transect stations"),
    options = layersControlOptions(collapsed = FALSE)
  )
  m <- addLegend(m, position = "bottomright", pal = pal, values = limits,
                 title = legend_title, opacity = 0.88)
  m
}

frame_local_time <- function(elapsed_s) {
  format(buoy_time_utc + elapsed_s, "%d %b %Y %H:%M %Z", tz = port_fairy_tz)
}

add_wave_frame_annotation <- function(r_ll, elapsed_s) {
  e <- ext(r_ll)
  x_span <- xmax(e) - xmin(e)
  y_span <- ymax(e) - ymin(e)
  bearing_to <- forcing$boundary_direction_to_deg_true[1]
  direction_statistic <- forcing$boundary_direction_statistic[1]
  angle <- bearing_to * pi / 180

  # The map is in longitude/latitude. Scale the longitude component by the
  # local latitude before drawing the true-north travel-direction arrow.
  arrow_y <- 0.13 * y_span
  arrow_x <- arrow_y * sin(angle) / cos(mean(c(ymin(e), ymax(e))) * pi / 180)
  arrow_dy <- arrow_y * cos(angle)
  arrow_x0 <- if (arrow_x >= 0) xmin(e) + 0.08 * x_span else xmax(e) - 0.08 * x_span
  arrow_y0 <- if (arrow_dy >= 0) ymin(e) + 0.10 * y_span else ymax(e) - 0.10 * y_span
  arrows(arrow_x0, arrow_y0, arrow_x0 + arrow_x, arrow_y0 + arrow_dy,
         col = "red", lwd = 3, length = 0.10)

  label <- c(
    paste("Local:", frame_local_time(elapsed_s)),
    sprintf("Model +%02.0f:%04.1f", floor(elapsed_s / 60), elapsed_s %% 60),
    sprintf("Buoy Hs %.2f m | Tp %.1f s", forcing$hs_m[1], forcing$tp_s[1]),
    sprintf("%s waves from %.0f° (arrow travels to %.0f°)",
            tools::toTitleCase(direction_statistic),
            forcing$boundary_direction_from_deg_true[1], bearing_to)
  )
  legend("topright", legend = label, bty = "o", bg = rgb(1, 1, 1, 0.82),
         cex = 0.74, text.col = "black")
}

elevation_wgs84 <- project(elevation_raster, "EPSG:4326", method = "bilinear")
depth_file_wgs84 <- project(depth_from_file_rotated, "EPSG:4326", method = "bilinear")
depth_delta_wgs84 <- project(depth_file_minus_raster, "EPSG:4326", method = "bilinear")

eta <- read_field("eta")
mask <- read_field("mask")
mask_wgs84 <- NULL
if (!is.null(mask)) {
  mask_rotated <- model_to_rotated_raster(mask, "mask")
  mask_wgs84 <- project(mask_rotated, "EPSG:4326", method = "near")
}

eta_paths <- list.files(results_dir, pattern = "^eta_[0-9]{5}$",
                        full.names = TRUE)
# FUNWAVE uses sequential output-file numbers in eta_00000, eta_00001, ...;
# they are not elapsed seconds. Convert those numbers using PLOT_INTV so the
# animation labels and the final-one-sixth selection use model time.
eta_file_number <- as.integer(sub("^eta_", "", basename(eta_paths)))
eta_order <- order(eta_file_number)
eta_paths <- eta_paths[eta_order]
eta_file_number <- eta_file_number[eta_order]
eta_time <- eta_file_number * grid$plot_intv_s
if (length(eta_paths) &&
    (any(!is.finite(eta_time)) || any(eta_time < 0) ||
     max(eta_time) > grid$total_time_s[1] + grid$plot_intv_s[1] ||
     max(eta_time) < grid$total_time_s[1] - grid$plot_intv_s[1])) {
  stop("eta file numbers do not map to the configured simulation period.")
}
read_public_output_raster <- function(path, label = basename(path)) {
  z <- read_text_model(path, basename(path))
  if (!is.null(mask)) z[mask <= 0] <- NA_real_
  make_public_output_rotated(model_to_rotated_raster(z, label))
}
maximum_public_raster <- function(fields) {
  if (!length(fields)) return(NULL)
  out <- fields[[1]]
  if (length(fields) > 1) {
    for (k in 2:length(fields)) {
      out <- terra::app(c(out, fields[[k]]), fun = max, na.rm = TRUE)
    }
  }
  out
}

# Retain report fields only at the requested public resolution. This is
# essential for 2.5 m historical cases: FUNWAVE still writes native 2.5 m
# files, but the report never holds all of those large matrices in memory.
eta_public_frames <- lapply(eta_paths, read_public_output_raster)
eta_max <- maximum_public_raster(eta_public_frames)
if (!is.null(eta_max)) {
  eta_max[!is.finite(eta_max)] <- NA_real_
}

# FUNWAVE's WaveHeight=T diagnostic is written by PREVIEW_MEAN as Hsig files.
# It uses the model time step internally, so it is the correct way to assess
# wave energy rather than comparing individual 15-second eta snapshots.
hsig_paths <- list.files(results_dir, pattern = "^Hsig_[0-9]{5}$",
                         full.names = TRUE)
hsig_number <- as.integer(sub("^Hsig_", "", basename(hsig_paths)))
hsig_paths <- hsig_paths[order(hsig_number)]
hsig_public_frames <- lapply(hsig_paths, read_public_output_raster)
hsig_max <- maximum_public_raster(hsig_public_frames)
if (!is.null(hsig_max)) {
  hsig_max[!is.finite(hsig_max)] <- NA_real_
}

# `eta_max` and `hsig_max` above are already rotated public-resolution rasters,
# not model-index matrices. The original raw field files remain untouched.
eta_frames <- NULL
hsig_frames <- NULL

# Read the five FUNWAVE station files.  Each file is time, eta, u, v and the
# position metadata was created with stations.txt by the case setup script.
station_metadata_path <- file.path(case_dir, "station_metadata.csv")
station_metadata <- if (file.exists(station_metadata_path)) {
  read.csv(station_metadata_path, check.names = FALSE)
} else {
  data.frame()
}
station_timeseries <- data.frame()
if (nrow(station_metadata)) {
  station_series <- lapply(seq_len(nrow(station_metadata)), function(k) {
    path <- file.path(results_dir, sprintf("sta_%04d", k))
    if (!file.exists(path)) return(NULL)
    z <- utils::read.table(path, header = FALSE)
    if (ncol(z) < 2) return(NULL)
    names(z)[seq_len(min(4, ncol(z)))] <- c("time_s", "eta_m", "u_m_s", "v_m_s")[seq_len(min(4, ncol(z)))]
    z$station_id <- station_metadata$station_id[k]
    z
  })
  station_series <- Filter(Negate(is.null), station_series)
  if (length(station_series)) station_timeseries <- do.call(rbind, station_series)
}
if (nrow(station_timeseries)) {
  station_plot <- plotly::plot_ly()
  for (k in seq_len(nrow(station_metadata))) {
    station_name <- station_metadata$station_id[k]
    z <- station_timeseries[station_timeseries$station_id == station_name, , drop = FALSE]
    if (nrow(z)) {
      station_plot <- plotly::add_lines(
        station_plot, data = z, x = ~time_s, y = ~eta_m, name = station_name,
        line = list(color = station_colours[k], width = 2),
        hovertemplate = paste0(station_name, "<br>t = %{x:.1f} s<br>eta = %{y:.3f} m<extra></extra>")
      )
    }
  }
  station_plot <- plotly::layout(
    station_plot, title = "FUNWAVE eta along buoy-to-nearshore transect",
    xaxis = list(title = "Model elapsed time (s)"),
    yaxis = list(title = "Free-surface elevation, eta (m)", range = eta_display_limits),
    legend = list(orientation = "h")
  )
  site_assets_dir <- file.path(report_site_dir, "assets")
  dir.create(site_assets_dir, recursive = TRUE, showWarnings = FALSE)
  htmlwidgets::saveWidget(
    station_plot, file.path(site_assets_dir, "port-fairy-transect-timeseries.html"),
    selfcontained = FALSE
  )
}

# The existing streamline diagnostic and particle-trail page use FUNWAVE's
# time-mean Umean/Vmean fields, not phase-resolved U/V snapshots.  That avoids
# using wave-orbital velocities as currents: the latter reverse each wave
# period, which makes otherwise smooth particles visibly wobble.  Option 2 is
# still a native-grid particle-trail map; option 1 remains an inspectable
# streamline diagnostic.
site_assets_dir <- file.path(report_site_dir, "assets")
dir.create(site_assets_dir, recursive = TRUE, showWarnings = FALSE)
current_particle_page_path <- file.path(site_assets_dir, "port-fairy-flow-particles.html")
current_particle_page_url <- "assets/port-fairy-flow-particles.html"
current_particle_page_status <- "No paired Umean/Vmean output was produced."
current_streamline_debug_url <- NULL
current_streamline_page_status <- "No paired Umean/Vmean output was produced."
u_paths <- list.files(results_dir, pattern = "^Umean_[0-9]{5}$",
                      ignore.case = TRUE, full.names = TRUE)
v_paths <- list.files(results_dir, pattern = "^Vmean_[0-9]{5}$",
                      ignore.case = TRUE, full.names = TRUE)
current_file_number <- function(paths) {
  as.integer(sub("^[[:alpha:]]+_", "", basename(paths)))
}
u_numbers <- current_file_number(u_paths)
v_numbers <- current_file_number(v_paths)
common_current_numbers <- sort(intersect(u_numbers, v_numbers))
current_numbers <- integer()
if (length(common_current_numbers)) {
  # A Umean/Vmean number labels the mean-output record, not elapsed seconds.
  # Use the newest paired record as the stable current field.
  current_numbers <- tail(common_current_numbers, 1)
}
current_label <- "No paired Umean/Vmean output was produced"
current_time_label <- "n/a"
current_peak_speed <- NA_real_
current_speed_floor <- NA_real_
current_seed_count <- 0L
current_streamline_count <- 0L
current_public_cell_count <- 0L
current_public_grid_shape <- "n/a"
current_selected_files <- "none"

model_xy_to_wgs84 <- function(xy_model) {
  xy_om <- if (source_is_low_x) {
    cbind(xmin(depth_raster) + xy_model[, 1], ymin(depth_raster) + xy_model[, 2])
  } else {
    cbind(xmax(depth_raster) - xy_model[, 1], ymax(depth_raster) - xy_model[, 2])
  }
  crds(project(vect(xy_om, type = "points", crs = crs(depth_raster)), "EPSG:4326"))
}

trace_current_streamline <- function(seed_xy, u_flow, v_flow, speed_floor,
                                     step_m = 25, n_steps = 100) {
  advance <- function(direction) {
    xy <- seed_xy
    out <- matrix(xy, ncol = 2)
    for (step in seq_len(n_steps)) {
      ij <- pmax(c(1, 1), pmin(c(nrow(u_flow), ncol(u_flow)),
                               round(xy / c(grid$dx_m[1], grid$dy_m[1])) + 1L))
      ux <- u_flow[ij[1], ij[2]]
      vy <- v_flow[ij[1], ij[2]]
      speed <- sqrt(ux^2 + vy^2)
      if (!is.finite(speed) || speed < speed_floor) break
      xy <- xy + direction * step_m * c(ux, vy) / speed
      if (xy[1] < grid$x_wk_m[1] || xy[1] > (nrow(u_flow) - 1) * grid$dx_m[1] ||
          xy[2] < 0 || xy[2] > (ncol(u_flow) - 1) * grid$dy_m[1]) break
      out <- rbind(out, xy)
    }
    out
  }
  backward <- advance(-1)
  forward <- advance(1)
  rbind(backward[nrow(backward):1, , drop = FALSE], forward[-1, , drop = FALSE])
}

if (length(current_numbers)) {
  read_current_component <- function(paths, numbers) {
    path_numbers <- current_file_number(paths)
    fields <- lapply(numbers, function(number) {
      path <- paths[match(number, path_numbers)]
      apply_output_visibility_mask_model(read_text_model(path, basename(path)))
    })
    total <- matrix(0, nrow = nrow(fields[[1]]), ncol = ncol(fields[[1]]))
    count <- matrix(0, nrow = nrow(fields[[1]]), ncol = ncol(fields[[1]]))
    for (field in fields) {
      keep <- is.finite(field)
      total[keep] <- total[keep] + field[keep]
      count[keep] <- count[keep] + 1
    }
    total[count == 0] <- NA_real_
    total[count > 0] <- total[count > 0] / count[count > 0]
    total
  }

  current_selected_files <- paste(sprintf("%05d", current_numbers), collapse = ", ")
  u_flow <- read_current_component(u_paths, current_numbers)
  v_flow <- read_current_component(v_paths, current_numbers)
  current_speed <- sqrt(u_flow^2 + v_flow^2)
  current_label <- "Time-mean depth-averaged current: latest Umean/Vmean output"
  current_time_label <- paste0(metres_label(grid$mean_wave_interval_s[1]),
                               " s FUNWAVE averaging interval")
  valid_speed <- current_speed[is.finite(current_speed) & current_speed > 0]
  speed_floor <- if (length(valid_speed)) {
    max(0.005, 0.02 * as.numeric(stats::quantile(valid_speed, 0.15, na.rm = TRUE)))
  } else {
    Inf
  }
  current_peak_speed <- if (length(valid_speed)) max(valid_speed) else NA_real_
  current_speed_floor <- speed_floor

  x_margin <- max(2 * grid$dx_m[1], 25)
  y_margin <- max(2 * grid$dy_m[1], 25)
  x_start <- grid$x_wk_m[1] + x_margin
  x_end <- (nrow(u_flow) - 1) * grid$dx_m[1] - x_margin
  y_start <- y_margin
  y_end <- (ncol(u_flow) - 1) * grid$dy_m[1] - y_margin
  seeds <- if (x_start < x_end && y_start < y_end && is.finite(speed_floor)) {
    as.matrix(expand.grid(
      seq(x_start, x_end, length.out = 15),
      seq(y_start, y_end, length.out = 11)
    ))
  } else {
    matrix(numeric(), ncol = 2)
  }
  trace_seed_set <- function(seed_set, threshold) {
    paths <- lapply(seq_len(nrow(seed_set)), function(k) {
      ij <- pmax(c(1, 1), pmin(c(nrow(u_flow), ncol(u_flow)),
                               round(seed_set[k, ] / c(grid$dx_m[1], grid$dy_m[1])) + 1L))
      if (!is.finite(current_speed[ij[1], ij[2]]) || current_speed[ij[1], ij[2]] < threshold) return(NULL)
      trace_current_streamline(
        seed_set[k, ], u_flow, v_flow, threshold,
        step_m = max(5, min(15, 2 * min(grid$dx_m[1], grid$dy_m[1]))),
        n_steps = 180
      )
    })
    Filter(function(z) !is.null(z) && nrow(z) >= 4, paths)
  }
  current_seed_count <- nrow(seeds)
  streamlines_model <- trace_seed_set(seeds, speed_floor)
  # Do not silently publish a blank animation because the initially selected
  # percentile threshold was too strict. This lower threshold remains based on
      # real FUNWAVE Umean/Vmean output; it merely allows slow, valid current paths to be
  # shown and records the fallback in the diagnostics.
  if (!length(streamlines_model) && length(valid_speed)) {
    fallback_floor <- max(0.0005, 0.005 * as.numeric(stats::quantile(
      valid_speed, 0.05, na.rm = TRUE
    )))
    if (is.finite(fallback_floor) && fallback_floor < speed_floor) {
      speed_floor <- fallback_floor
      current_speed_floor <- speed_floor
      streamlines_model <- trace_seed_set(seeds, speed_floor)
      if (length(streamlines_model)) {
        current_label <- paste0(current_label, " (low-speed path fallback)")
      }
    }
  }
  streamlines_ll <- lapply(streamlines_model, model_xy_to_wgs84)
  streamlines_ll <- Filter(function(z) {
    is.matrix(z) && nrow(z) >= 4 && all(is.finite(z))
  }, streamlines_ll)
  current_streamline_count <- length(streamlines_ll)

  west <- min(model_outline_ll_xy[, 1]); east <- max(model_outline_ll_xy[, 1])
  south <- min(model_outline_ll_xy[, 2]); north <- max(model_outline_ll_xy[, 2])
  x_pad <- (east - west) * 0.04; y_pad <- (north - south) * 0.04
  current_bounds <- list(c(south - y_pad, west - x_pad), c(north + y_pad, east + x_pad))
  outline_latlng <- lapply(seq_len(nrow(model_outline_ll_xy)), function(k) {
    unname(as.numeric(model_outline_ll_xy[k, c(2, 1)]))
  })
  source_latlng <- lapply(seq_len(nrow(source_line_ll_xy)), function(k) {
    unname(as.numeric(source_line_ll_xy[k, c(2, 1)]))
  })
  buoy_latlng <- unname(as.numeric(c(crds(buoy_ll)[1, 2], crds(buoy_ll)[1, 1])))
  debug_prefix <- paste(
    "Paired Umean/Vmean files:", length(common_current_numbers),
    "Selected file number(s):", current_selected_files,
    "Field:", current_label,
    "Output window:", current_time_label,
    "Peak speed (m/s):", ifelse(is.finite(current_peak_speed), format(round(current_peak_speed, 3), nsmall = 3), "n/a"),
    "Trace floor (m/s):", ifelse(is.finite(current_speed_floor), format(round(current_speed_floor, 3), nsmall = 3), "n/a"),
    sep = "\n"
  )

  # Option 1: retain a small, independently inspectable streamline map. The
  # debug panel exposes every failure point instead of silently drawing no
  # paths. It is intentionally kept separate from the grid-particle map.
  option1_data <- list(
    bounds = current_bounds,
    lines = lapply(streamlines_ll, function(z) {
      lapply(seq_len(nrow(z)), function(k) unname(as.numeric(z[k, c(2, 1)])))
    }),
    outline = outline_latlng,
    source = source_latlng,
    buoy = buoy_latlng,
    label = paste(current_label, paste0("Output window: ", current_time_label), sep = "\n"),
    debug = paste(debug_prefix, "Option 1 seeds:", current_seed_count,
                  "Option 1 completed paths:", current_streamline_count,
                  sep = "\n")
  )
  option1_js <- paste0(
    "function(el, x, data) {",
    "var map=this;if(!map||typeof map.getPanes!=='function')map=this.map||this._map;",
    "if(!map||typeof map.getPanes!=='function'){console.error('Leaflet map was not available for option 1 streamlines');return;}",
    "var bounds=L.latLngBounds(data.bounds),lines=data.lines||[];",
    "var osm=L.tileLayer('https://{s}.tile.openstreetmap.org/{z}/{x}/{y}.png',{attribution:'&copy; OpenStreetMap contributors'}).addTo(map);",
    "var satellite=L.tileLayer('https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}',{attribution:'Tiles &copy; Esri'});",
    "var paths=L.layerGroup().addTo(map),particles=L.layerGroup().addTo(map);",
    "var grid=L.polyline(data.outline,{color:'black',weight:2}),paddle=L.polyline(data.source,{color:'dodgerblue',weight:2,dashArray:'6 6'}),buoy=L.circleMarker(data.buoy,{radius:5,color:'red',fillOpacity:1});",
    "grid.addTo(map);paddle.addTo(map);buoy.addTo(map);map.fitBounds(bounds,{padding:[4,4]});map.setMaxBounds(bounds);",
    "L.control.layers({'OpenStreetMap':osm,'Satellite imagery':satellite},{'Option 1 flow paths':paths,'Option 1 moving dots':particles,'FUNWAVE grid':grid,'Internal wavemaker':paddle,'Buoy':buoy},{collapsed:false}).addTo(map);",
    "function panel(position,text){var c=L.control({position:position});c.onAdd=function(){var d=L.DomUtil.create('div');d.style.background='rgba(255,255,255,.92)';d.style.padding='6px';d.style.font='12px Arial';d.style.whiteSpace='pre-line';d.textContent=text;return d;};c.addTo(map);}",
    "panel('bottomleft',data.label);panel('topright',data.debug);",
    "var dots=[];lines.forEach(function(line){if(line&&line.length>1){L.polyline(line,{color:'#2c7fb8',weight:1.2,opacity:.72}).addTo(paths);dots.push({line:line,marker:L.circleMarker(line[0],{radius:3,color:'#08306b',weight:0,fillOpacity:.95}).addTo(particles)});}});",
    "if(dots.length){function animate(t){dots.forEach(function(item,i){var q=((t/6500)+(i/dots.length))%1*(item.line.length-1),j=Math.floor(q),a=item.line[j],b=item.line[Math.min(j+1,item.line.length-1)],f=q-j;item.marker.setLatLng([a[0]+f*(b[0]-a[0]),a[1]+f*(b[1]-a[1])]);});window.requestAnimationFrame(animate);}window.requestAnimationFrame(animate);}",
    "window.setTimeout(function(){map.invalidateSize();map.fitBounds(bounds,{padding:[4,4]});},250);",
    "}"
  )
  option1_widget <- leaflet::leaflet(
    width = "100%", height = 560,
    options = leaflet::leafletOptions(minZoom = 14, maxZoom = 18, maxBoundsViscosity = 1)
  )
  option1_widget <- htmlwidgets::onRender(option1_widget, option1_js, data = option1_data)
  option1_page_path <- file.path(site_assets_dir, "port-fairy-current-streamline-diagnostic.html")
  current_streamline_page_status <- tryCatch({
    # Keep the diagnostic page independent of the report's own HTML assets.
    # If it cannot be written, option 2 still renders and the failure is shown
    # in the report table rather than aborting the complete report.
    htmlwidgets::saveWidget(option1_widget, option1_page_path, selfcontained = TRUE)
    if (!file.exists(option1_page_path) || file.info(option1_page_path)$size < 5000) {
      stop("The option 1 streamline page was not written or is unexpectedly small.")
    }
    paste0("Published with ", current_streamline_count, " traced current paths.")
  }, error = function(e) {
    paste0("Option 1 streamline page write failed: ", conditionMessage(e))
  })
  if (file.exists(option1_page_path) && file.info(option1_page_path)$size >= 5000) {
    current_streamline_debug_url <- "assets/port-fairy-current-streamline-diagnostic.html"
  }

  # The particle paths are traced through time-mean Umean/Vmean on the native
  # rotated grid, using nearest native cells (no spatial interpolation). They
  # are then projected to WGS84 as paths, so they retain the correct geographic
  # direction even though FUNWAVE components are model-axis components.
  particle_routes <- lapply(streamlines_ll, function(z) {
    lapply(seq_len(nrow(z)), function(k) unname(as.numeric(z[k, c(2, 1)])))
  })
  current_public_cell_count <- sum(vapply(particle_routes, length, integer(1)))
  current_public_grid_shape <- paste0(metres_label(grid$dx_m[1]), "m x ",
                                      metres_label(grid$dy_m[1]), "m native paths")
  option2_data <- list(
    bounds = current_bounds,
    outline = outline_latlng,
    source = source_latlng,
    buoy = buoy_latlng,
    routes = particle_routes,
    label = paste(current_label, paste0("Output window: ", current_time_label),
                  "Particle trails follow Umean/Vmean paths traced on the native rotated grid.",
                  "Animation speed is visual only; direction is the simulated flow direction.",
                  sep = "\n"),
    debug = paste(debug_prefix, "Native current path grid:", current_public_grid_shape,
                  "Animated paths:", length(particle_routes),
                  "Path vertices:", current_public_cell_count,
                  sep = "\n"),
    routeCount = length(particle_routes)
  )
  option2_js <- paste0(
    "function(el, x, data) {",
    "var map=this;if(!map||typeof map.getPanes!=='function')map=this.map||this._map;",
    "if(!map||typeof map.getPanes!=='function'){console.error('Leaflet map was not available for flow particles');return;}",
    "var routes=data.routes||[],bounds=L.latLngBounds(data.bounds),particles=[];",
    "function pointAt(route,phase){var n=route.length-1;if(n<1)return null;var t=((phase%1)+1)%1*n,i=Math.floor(t),f=t-i,a=route[i],b=route[Math.min(i+1,n)];return [a[0]+f*(b[0]-a[0]),a[1]+f*(b[1]-a[1])];}",
    "routes.forEach(function(route,i){if(route&&route.length>3){for(var j=0;j<3;j++)particles.push({route:route,offset:(i*0.137+j/3)%1});}});",
    "var FlowTrails=L.Layer.extend({initialize:function(){this._running=false;},onAdd:function(m){this._map=m;this._canvas=L.DomUtil.create('canvas','leaflet-zoom-animated',m.getPanes().overlayPane);this._canvas.style.pointerEvents='none';this._canvas.style.zIndex=650;this._resize();m.on('resize moveend zoomend',this._resize,this);this._running=true;this._frame=window.requestAnimationFrame(this._animate.bind(this));},onRemove:function(m){this._running=false;window.cancelAnimationFrame(this._frame);m.off('resize moveend zoomend',this._resize,this);L.DomUtil.remove(this._canvas);this._map=null;},_resize:function(){if(!this._map||!this._canvas)return;var s=this._map.getSize(),d=window.devicePixelRatio||1;this._canvas.width=Math.max(1,Math.round(s.x*d));this._canvas.height=Math.max(1,Math.round(s.y*d));this._canvas.style.width=s.x+'px';this._canvas.style.height=s.y+'px';this._topLeft=this._map.containerPointToLayerPoint([0,0]);L.DomUtil.setPosition(this._canvas,this._topLeft);this._ctx=this._canvas.getContext('2d');this._ctx.setTransform(d,0,0,d,0,0);},_animate:function(t){if(!this._running)return;var ctx=this._ctx,s=this._map.getSize(),phase=t/18000;ctx.clearRect(0,0,s.x,s.y);ctx.lineCap='round';particles.forEach(function(p){var trail=[];for(var q=13;q>=0;q--){var ll=pointAt(p.route,phase+p.offset-q*0.010);if(ll)trail.push(ll);}if(trail.length<2)return;ctx.beginPath();trail.forEach(function(ll,k){var pt=map.latLngToLayerPoint(ll),x=pt.x-this._topLeft.x,y=pt.y-this._topLeft.y;if(k===0)ctx.moveTo(x,y);else ctx.lineTo(x,y);},this);ctx.strokeStyle='rgba(8,48,107,0.72)';ctx.lineWidth=1.3;ctx.stroke();var head=trail[trail.length-1],hp=map.latLngToLayerPoint(head);ctx.fillStyle='rgba(44,127,184,0.96)';ctx.beginPath();ctx.arc(hp.x-this._topLeft.x,hp.y-this._topLeft.y,2.1,0,Math.PI*2);ctx.fill();},this);this._frame=window.requestAnimationFrame(this._animate.bind(this));}});",
    "var flowTrails=new FlowTrails();flowTrails.addTo(map);",
    "L.control.layers({}, {'Flowing particle trails':flowTrails},{collapsed:false}).addTo(map);",
    "function panel(position,text){var c=L.control({position:position});c.onAdd=function(){var d=L.DomUtil.create('div');d.style.background='rgba(255,255,255,.92)';d.style.padding='6px';d.style.font='12px Arial';d.style.whiteSpace='pre-line';d.textContent=text;return d;};c.addTo(map);}",
    "panel('bottomleft',data.label);panel('topright',data.debug);",
    "window.setTimeout(function(){map.invalidateSize();map.fitBounds(bounds,{padding:[4,4]});map.setMaxBounds(bounds);},250);",
    "}"
  )
  option2_widget <- leaflet::leaflet(
    width = "100%", height = 620,
    options = leaflet::leafletOptions(minZoom = 14, maxZoom = 18, maxBoundsViscosity = 1)
  )
  option2_widget <- leaflet::addProviderTiles(option2_widget, "OpenStreetMap.Mapnik", group = "OpenStreetMap")
  option2_widget <- leaflet::addProviderTiles(option2_widget, "Esri.WorldImagery", group = "Satellite imagery")
  option2_widget <- leaflet::setMaxBounds(option2_widget,
                                           current_bounds[[1]][2], current_bounds[[1]][1],
                                           current_bounds[[2]][2], current_bounds[[2]][1])
  option2_widget <- leaflet::fitBounds(option2_widget,
                                       current_bounds[[1]][2], current_bounds[[1]][1],
                                       current_bounds[[2]][2], current_bounds[[2]][1])
  option2_widget <- leaflet::addPolylines(option2_widget,
                                           lng = model_outline_ll_xy[, 1], lat = model_outline_ll_xy[, 2],
                                           color = "black", weight = 2, group = "FUNWAVE grid")
  option2_widget <- leaflet::addPolylines(option2_widget,
                                           lng = source_line_ll_xy[, 1], lat = source_line_ll_xy[, 2],
                                           color = "dodgerblue3", weight = 2, dashArray = "6 6",
                                           group = "Internal wavemaker")
  option2_widget <- leaflet::addCircleMarkers(option2_widget,
                                               lng = crds(buoy_ll)[1, 1], lat = crds(buoy_ll)[1, 2],
                                               radius = 5, color = "red", fillOpacity = 1, group = "Buoy")
  for (route in streamlines_ll) {
    option2_widget <- leaflet::addPolylines(option2_widget,
                                             lng = route[, 1], lat = route[, 2],
                                             color = "#2c7fb8", weight = 1, opacity = 0.28,
                                             group = "Mean-flow paths")
  }
  option2_widget <- leaflet::addLayersControl(
    option2_widget,
    baseGroups = c("OpenStreetMap", "Satellite imagery"),
    overlayGroups = c("Mean-flow paths", "FUNWAVE grid", "Internal wavemaker", "Buoy"),
    options = leaflet::layersControlOptions(collapsed = FALSE)
  )
  option2_widget <- htmlwidgets::onRender(option2_widget, option2_js, data = option2_data)
  current_particle_page_status <- tryCatch({
    htmlwidgets::saveWidget(option2_widget, current_particle_page_path, selfcontained = TRUE)
    if (!file.exists(current_particle_page_path) || file.info(current_particle_page_path)$size < 5000) {
      stop("The particle page was not written or is unexpectedly small.")
    }
    paste0("Published with ", length(particle_routes), " animated current paths.")
  }, error = function(e) {
    paste0("Particle page write failed: ", conditionMessage(e))
  })
}

if (!file.exists(current_particle_page_path)) {
  fallback_text <- paste(
    "The flow-particle page could not be animated for this run.",
    paste0("Umean files found: ", length(u_paths)),
    paste0("Vmean files found: ", length(v_paths)),
    paste0("Paired fields: ", length(common_current_numbers)),
    paste0("Status: ", current_particle_page_status),
    sep = "\n"
  )
  fallback_saved <- tryCatch({
    htmltools::save_html(
      htmltools::tags$html(
        htmltools::tags$head(htmltools::tags$title("Port Fairy current-flow diagnostics")),
        htmltools::tags$body(
          style = "font-family: system-ui, sans-serif; max-width: 58rem; margin: 3rem auto; line-height: 1.45;",
          htmltools::tags$h1("Port Fairy current-flow diagnostics"),
          htmltools::tags$p("No animated paths were available for this model run."),
          htmltools::tags$pre(fallback_text)
        )
      ),
      file = current_particle_page_path
    )
    TRUE
  }, error = function(e) {
    current_particle_page_status <<- paste0(current_particle_page_status,
                                             " Fallback page write failed: ", conditionMessage(e))
    FALSE
  })
  if (fallback_saved) {
    current_particle_page_status <- paste0(current_particle_page_status,
                                           " A diagnostic page was published instead.")
  }
}

current_debug <- data.frame(
  Diagnostic = c("Umean field files found", "Vmean field files found", "Paired Umean/Vmean field files",
                 "Selected file number(s)", "Current field used", "Mean-current output window",
                 "Peak vector speed (m/s)", "Trace / particle floor (m/s)",
                 "Option 1 seeds", "Option 1 completed paths",
                 "Particle-path native grid",
                 "Option 1 diagnostic page", "Option 2 path vertices", "Particle-trail page"),
  Value = c(length(u_paths), length(v_paths), length(common_current_numbers),
            current_selected_files, current_label, current_time_label,
            ifelse(is.finite(current_peak_speed), format(round(current_peak_speed, 3), nsmall = 3), "n/a"),
            ifelse(is.finite(current_speed_floor), format(round(current_speed_floor, 3), nsmall = 3), "n/a"),
            current_seed_count, current_streamline_count,
            current_public_grid_shape, current_streamline_page_status,
            current_public_cell_count, current_particle_page_status),
  check.names = FALSE, stringsAsFactors = FALSE
)

# -------------------------------------------------------------------------
# Additional Earth-style, time-mean current animation
#
# This is deliberately additional to the repaired Umean/Vmean streamline and
# path pages above.  It uses a display-field interpolation and a particle
# ageing/trail renderer rather than pre-traced routes, while still sourcing
# the same FUNWAVE Umean/Vmean outputs over T_INTV_mean.
#
# The browser animator below adapts the particle ageing, vector interpolation,
# destination-in trail fade, and per-frame evolve/draw approach in earth.js
# from https://github.com/cambecc/earth (MIT License).
# Copyright (c) 2014 Cameron Beccario.  The complete MIT notice is embedded
# in the generated page data so it accompanies the adapted implementation.
earth_flow_page_path <- file.path(site_assets_dir, "port-fairy-earth-nullschool-flow.html")
earth_flow_page_url <- "assets/port-fairy-earth-nullschool-flow.html"
earth_flow_page_status <- "No paired Umean/Vmean output was produced."
earth_u_paths <- list.files(results_dir, pattern = "^Umean_[0-9]{5}$",
                            ignore.case = TRUE, full.names = TRUE)
earth_v_paths <- list.files(results_dir, pattern = "^Vmean_[0-9]{5}$",
                            ignore.case = TRUE, full.names = TRUE)
earth_file_number <- function(paths) {
  as.integer(sub("^[[:alpha:]]+_", "", basename(paths)))
}
earth_u_numbers <- earth_file_number(earth_u_paths)
earth_v_numbers <- earth_file_number(earth_v_paths)
earth_common_numbers <- sort(intersect(earth_u_numbers, earth_v_numbers))
earth_selected_number <- NA_integer_
earth_selected_files <- "none"
earth_peak_speed <- NA_real_
earth_min_speed <- NA_real_
earth_display_grid <- "n/a"
earth_particle_count <- 0L

earth_mit_notice <- paste(
  "The MIT License (MIT)",
  "Copyright (c) 2014 Cameron Beccario",
  "Permission is hereby granted, free of charge, to any person obtaining a copy of this software and associated documentation files (the Software), to deal in the Software without restriction, including without limitation the rights to use, copy, modify, merge, publish, distribute, sublicense, and/or sell copies of the Software, and to permit persons to whom the Software is furnished to do so, subject to the following conditions:",
  "The above copyright notice and this permission notice shall be included in all copies or substantial portions of the Software.",
  "THE SOFTWARE IS PROVIDED AS IS, WITHOUT WARRANTY OF ANY KIND, EXPRESS OR IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY, FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM, OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN THE SOFTWARE.",
  sep = "\n\n"
)

# Convert the public Omerc raster back to model [x, y] order. This is the
# inverse of model_to_rotated_raster() and preserves the model velocity axes.
rotated_raster_to_model_xy <- function(r_om) {
  gis_rows <- terra::as.matrix(r_om, wide = TRUE)
  if (source_is_low_x) {
    t(gis_rows[nrow(gis_rows):1, , drop = FALSE])
  } else {
    t(gis_rows[, ncol(gis_rows):1, drop = FALSE])
  }
}
matrix_as_rows <- function(z) {
  unname(lapply(seq_len(nrow(z)), function(k) as.numeric(z[k, ])))
}

earth_west <- min(model_outline_ll_xy[, 1]); earth_east <- max(model_outline_ll_xy[, 1])
earth_south <- min(model_outline_ll_xy[, 2]); earth_north <- max(model_outline_ll_xy[, 2])
earth_x_pad <- (earth_east - earth_west) * 0.04
earth_y_pad <- (earth_north - earth_south) * 0.04
earth_bounds <- list(
  c(earth_south - earth_y_pad, earth_west - earth_x_pad),
  c(earth_north + earth_y_pad, earth_east + earth_x_pad)
)
earth_outline_latlng <- lapply(seq_len(nrow(model_outline_ll_xy)), function(k) {
  unname(as.numeric(model_outline_ll_xy[k, c(2, 1)]))
})
earth_source_latlng <- lapply(seq_len(nrow(source_line_ll_xy)), function(k) {
  unname(as.numeric(source_line_ll_xy[k, c(2, 1)]))
})
earth_buoy_latlng <- unname(as.numeric(c(crds(buoy_ll)[1, 2], crds(buoy_ll)[1, 1])))

if (length(earth_common_numbers)) {
  earth_selected_number <- tail(earth_common_numbers, 1)
  earth_selected_files <- paste0("Umean_", sprintf("%05d", earth_selected_number),
                                 ", Vmean_", sprintf("%05d", earth_selected_number))
  earth_u_path <- earth_u_paths[match(earth_selected_number, earth_u_numbers)]
  earth_v_path <- earth_v_paths[match(earth_selected_number, earth_v_numbers)]
  earth_u_native <- apply_output_visibility_mask_model(
    read_text_model(earth_u_path, basename(earth_u_path))
  )
  earth_v_native <- apply_output_visibility_mask_model(
    read_text_model(earth_v_path, basename(earth_v_path))
  )

  # Use the same compact public x/y resolution as the map products. The
  # current field remains a model-derived display; station files remain exact
  # native grid-point outputs with no interpolation.
  earth_u_public_om <- make_public_output_rotated(
    model_to_rotated_raster(earth_u_native, basename(earth_u_path))
  )
  earth_v_public_om <- make_public_output_rotated(
    model_to_rotated_raster(earth_v_native, basename(earth_v_path))
  )
  earth_u_public <- rotated_raster_to_model_xy(earth_u_public_om)
  earth_v_public <- rotated_raster_to_model_xy(earth_v_public_om)
  earth_valid <- is.finite(earth_u_public) & is.finite(earth_v_public)
  earth_u_public[!earth_valid] <- NA_real_
  earth_v_public[!earth_valid] <- NA_real_
  earth_speed <- sqrt(earth_u_public^2 + earth_v_public^2)
  earth_valid_speed <- earth_speed[is.finite(earth_speed) & earth_speed > 0]
  earth_peak_speed <- if (length(earth_valid_speed)) max(earth_valid_speed) else NA_real_
  earth_min_speed <- if (length(earth_valid_speed)) {
    max(0.0005, 0.01 * as.numeric(stats::quantile(earth_valid_speed, 0.10, na.rm = TRUE)))
  } else {
    Inf
  }
  earth_display_dx_m <- grid$dx_m[1] * public_output_factor_x
  earth_display_dy_m <- grid$dy_m[1] * public_output_factor_y
  earth_display_grid <- paste0(metres_label(earth_display_dx_m), " m x ",
                               metres_label(earth_display_dy_m), " m")

  # A local Oblique Mercator grid is effectively affine across this small
  # domain. Passing its four geographic corners lets the browser render the
  # vector field in WGS84 while interpolation and particle advection remain
  # in FUNWAVE's x/y coordinate system.
  earth_native_corners <- rbind(
    c(0, 0),
    c((grid$Mglob[1] - 1) * grid$dx_m[1], 0),
    c((grid$Mglob[1] - 1) * grid$dx_m[1], (grid$Nglob[1] - 1) * grid$dy_m[1]),
    c(0, (grid$Nglob[1] - 1) * grid$dy_m[1])
  )
  earth_corner_ll <- model_xy_to_wgs84(earth_native_corners)
  earth_corner_latlng <- lapply(seq_len(nrow(earth_corner_ll)), function(k) {
    unname(as.numeric(earth_corner_ll[k, c(2, 1)]))
  })
  earth_particle_count <- max(350L, min(1800L, as.integer(round(2.0 * 900))))
  earth_data <- list(
    bounds = earth_bounds,
    outline = earth_outline_latlng,
    source = earth_source_latlng,
    buoy = earth_buoy_latlng,
    corners = earth_corner_latlng,
    u = matrix_as_rows(earth_u_public),
    v = matrix_as_rows(earth_v_public),
    nx = nrow(earth_u_public), ny = ncol(earth_u_public),
    dx_m = earth_display_dx_m, dy_m = earth_display_dy_m,
    min_speed_m_s = earth_min_speed,
    max_speed_m_s = ifelse(is.finite(earth_peak_speed), earth_peak_speed, 0),
    particle_count = earth_particle_count,
    model_time_label = paste0("Latest FUNWAVE Umean/Vmean output; ",
                              metres_label(grid$mean_wave_interval_s[1]),
                              " s averaging interval"),
    credit = "Particle animator adapted from Cameron Beccario's earth (MIT).",
    mit_notice = earth_mit_notice
  )
  earth_js <- paste0(
    "function(el, x, data) {",
    "/* Adapted from earth.js in cambecc/earth. Copyright (c) 2014 Cameron Beccario; MIT notice supplied in data.mit_notice. */",
    "var map=this;if(!map||typeof map.getPanes!=='function')map=this.map||this._map;",
    "if(!map||typeof map.getPanes!=='function'){console.error('Leaflet map was not available for Earth-style particles');return;}",
    "var bounds=L.latLngBounds(data.bounds),u=data.u||[],v=data.v||[],nx=data.nx,ny=data.ny;",
    "function finite(z){return typeof z==='number'&&isFinite(z);}",
    "function bilerp(a,b,c,d,fx,fy){return a*(1-fx)*(1-fy)+b*fx*(1-fy)+c*(1-fx)*fy+d*fx*fy;}",
    "function vectorAt(px,py){if(!(px>=0&&py>=0&&px<nx-1&&py<ny-1))return null;var i=Math.floor(px),j=Math.floor(py),fx=px-i,fy=py-j;var u00=u[i][j],u10=u[i+1][j],u01=u[i][j+1],u11=u[i+1][j+1],v00=v[i][j],v10=v[i+1][j],v01=v[i][j+1],v11=v[i+1][j+1];if(!finite(u00)||!finite(u10)||!finite(u01)||!finite(u11)||!finite(v00)||!finite(v10)||!finite(v01)||!finite(v11))return null;var ux=bilerp(u00,u10,u01,u11,fx,fy),vy=bilerp(v00,v10,v01,v11,fx,fy);return {u:ux,v:vy,m:Math.sqrt(ux*ux+vy*vy)};}",
    "function modelToLatLng(px,py){var fx=Math.max(0,Math.min(1,px/(nx-1))),fy=Math.max(0,Math.min(1,py/(ny-1))),c=data.corners;return [bilerp(c[0][0],c[1][0],c[3][0],c[2][0],fx,fy),bilerp(c[0][1],c[1][1],c[3][1],c[2][1],fx,fy)];}",
    "var osm=L.tileLayer('https://{s}.tile.openstreetmap.org/{z}/{x}/{y}.png',{attribution:'&copy; OpenStreetMap contributors'}).addTo(map);",
    "var satellite=L.tileLayer('https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}',{attribution:'Tiles &copy; Esri'});",
    "var grid=L.polyline(data.outline,{color:'black',weight:2}),paddle=L.polyline(data.source,{color:'dodgerblue',weight:2,dashArray:'6 6'}),buoy=L.circleMarker(data.buoy,{radius:5,color:'red',fillOpacity:1});",
    "grid.addTo(map);paddle.addTo(map);buoy.addTo(map);map.fitBounds(bounds,{padding:[4,4]});map.setMaxBounds(bounds);",
    "function panel(position,text){var c=L.control({position:position});c.onAdd=function(){var d=L.DomUtil.create('div');d.style.background='rgba(255,255,255,.92)';d.style.padding='6px';d.style.font='12px Arial';d.style.whiteSpace='pre-line';d.textContent=text;return d;};c.addTo(map);return c;}",
    "panel('bottomleft',data.model_time_label+'\\nDisplay field: '+data.dx_m+' m x '+data.dy_m+' m\\nParticle speed is visual only.');panel('bottomright',data.credit);",
    "var MAX_PARTICLE_AGE=100,DISPLAY_TIME_SCALE=160,particleCount=Math.max(350,Math.min(1800,Math.round(map.getSize().x*2)));",
    "var EarthFlowLayer=L.Layer.extend({initialize:function(){this._running=false;this._particles=[];},onAdd:function(m){this._map=m;this._canvas=L.DomUtil.create('canvas','leaflet-zoom-animated',m.getPanes().overlayPane);this._canvas.style.pointerEvents='none';this._canvas.style.zIndex=650;this._resize();m.on('resize moveend zoomend',this._resize,this);for(var k=0;k<particleCount;k++)this._particles.push(this._randomize({age:Math.random()*MAX_PARTICLE_AGE}));this._running=true;this._last=performance.now();this._frame=window.requestAnimationFrame(this._animate.bind(this));},onRemove:function(m){this._running=false;window.cancelAnimationFrame(this._frame);m.off('resize moveend zoomend',this._resize,this);L.DomUtil.remove(this._canvas);this._map=null;},_resize:function(){if(!this._map||!this._canvas)return;var s=this._map.getSize(),d=window.devicePixelRatio||1;this._canvas.width=Math.max(1,Math.round(s.x*d));this._canvas.height=Math.max(1,Math.round(s.y*d));this._canvas.style.width=s.x+'px';this._canvas.style.height=s.y+'px';this._topLeft=this._map.containerPointToLayerPoint([0,0]);L.DomUtil.setPosition(this._canvas,this._topLeft);this._ctx=this._canvas.getContext('2d');this._ctx.setTransform(d,0,0,d,0,0);},_randomize:function(p){for(var n=0;n<1200;n++){p.x=Math.random()*(nx-1);p.y=Math.random()*(ny-1);var w=vectorAt(p.x,p.y);if(w&&w.m>=data.min_speed_m_s){p.age=Math.random()*MAX_PARTICLE_AGE;return p;}}p.age=MAX_PARTICLE_AGE+1;return p;},_colour:function(m){var q=Math.max(0,Math.min(1,m/Math.max(data.max_speed_m_s,1e-6)));if(q<.25)return '#ffffcc';if(q<.5)return '#a1dab4';if(q<.75)return '#41b6c4';return '#225ea8';},_animate:function(now){if(!this._running)return;var dt=Math.min(.05,Math.max(.001,(now-this._last)/1000));this._last=now;var ctx=this._ctx,s=this._map.getSize(),self=this;var previous=ctx.globalCompositeOperation;ctx.globalCompositeOperation='destination-in';ctx.fillStyle='rgba(0,0,0,0.92)';ctx.fillRect(0,0,s.x,s.y);ctx.globalCompositeOperation=previous;ctx.lineWidth=1.15;ctx.lineCap='round';this._particles.forEach(function(p){if(p.age>MAX_PARTICLE_AGE)self._randomize(p);var w=vectorAt(p.x,p.y);if(!w||w.m<data.min_speed_m_s){p.age=MAX_PARTICLE_AGE+1;return;}var xt=p.x+w.u*dt*DISPLAY_TIME_SCALE/data.dx_m,yt=p.y+w.v*dt*DISPLAY_TIME_SCALE/data.dy_m,next=vectorAt(xt,yt);if(!next||next.m<data.min_speed_m_s){p.age=MAX_PARTICLE_AGE+1;return;}var a=modelToLatLng(p.x,p.y),b=modelToLatLng(xt,yt),pa=map.latLngToLayerPoint(a),pb=map.latLngToLayerPoint(b);ctx.beginPath();ctx.strokeStyle=self._colour(w.m);ctx.moveTo(pa.x-self._topLeft.x,pa.y-self._topLeft.y);ctx.lineTo(pb.x-self._topLeft.x,pb.y-self._topLeft.y);ctx.stroke();p.x=xt;p.y=yt;p.age+=1;});this._frame=window.requestAnimationFrame(this._animate.bind(this));}});",
    "var earthFlow=new EarthFlowLayer();earthFlow.addTo(map);",
    "L.control.layers({'OpenStreetMap':osm,'Satellite imagery':satellite},{'Earth-style mean-current particles':earthFlow,'FUNWAVE grid':grid,'Internal wavemaker':paddle,'Buoy':buoy},{collapsed:false}).addTo(map);",
    "window.setTimeout(function(){map.invalidateSize();map.fitBounds(bounds,{padding:[4,4]});map.setMaxBounds(bounds);},250);",
    "}"
  )
  earth_widget <- leaflet::leaflet(
    width = "100%", height = 620,
    options = leaflet::leafletOptions(minZoom = 14, maxZoom = 18, maxBoundsViscosity = 1)
  )
  earth_widget <- htmlwidgets::onRender(earth_widget, earth_js, data = earth_data)
  earth_flow_page_status <- tryCatch({
    htmlwidgets::saveWidget(earth_widget, earth_flow_page_path, selfcontained = TRUE)
    if (!file.exists(earth_flow_page_path) || file.info(earth_flow_page_path)$size < 5000) {
      stop("The Earth-style mean-current page was not written or is unexpectedly small.")
    }
    paste0("Published from Umean/Vmean file ", sprintf("%05d", earth_selected_number),
           "; peak mean speed ", format(round(earth_peak_speed, 3), nsmall = 3), " m/s.")
  }, error = function(e) {
    paste0("Earth-style flow page write failed: ", conditionMessage(e))
  })
}

if (!file.exists(earth_flow_page_path)) {
  earth_fallback <- paste(
    "The Earth-style mean-current particle page could not be animated for this run.",
    paste0("Umean files found: ", length(earth_u_paths)),
    paste0("Vmean files found: ", length(earth_v_paths)),
    paste0("Paired mean-current fields: ", length(earth_common_numbers)),
    paste0("Status: ", earth_flow_page_status),
    sep = "\n"
  )
  earth_fallback_saved <- tryCatch({
    htmltools::save_html(
      htmltools::tags$html(
        htmltools::tags$head(htmltools::tags$title("Port Fairy Earth-style mean-current diagnostics")),
        htmltools::tags$body(
          style = "font-family: system-ui, sans-serif; max-width: 58rem; margin: 3rem auto; line-height: 1.45;",
          htmltools::tags$h1("Port Fairy Earth-style mean-current diagnostics"),
          htmltools::tags$p("No animated mean-current field was available for this model run."),
          htmltools::tags$pre(earth_fallback)
        )
      ),
      file = earth_flow_page_path
    )
    TRUE
  }, error = function(e) {
    earth_flow_page_status <<- paste0(earth_flow_page_status,
                                      " Fallback page write failed: ", conditionMessage(e))
    FALSE
  })
  if (earth_fallback_saved) {
    earth_flow_page_status <- paste0(earth_flow_page_status,
                                     " A diagnostic page was published instead.")
  }
}

earth_flow_debug <- data.frame(
  Diagnostic = c("Umean field files found", "Vmean field files found",
                 "Paired Umean/Vmean fields", "Selected file(s)",
                 "Mean-current display grid", "Peak mean speed (m/s)",
                 "Earth-style mean-current page"),
  Value = c(length(earth_u_paths), length(earth_v_paths), length(earth_common_numbers),
            earth_selected_files, earth_display_grid,
            ifelse(is.finite(earth_peak_speed), format(round(earth_peak_speed, 3), nsmall = 3), "n/a"),
            earth_flow_page_status),
  check.names = FALSE, stringsAsFactors = FALSE
)

# A single fixed range makes every eta frame and map comparable across days.
eta_animation_limits <- eta_display_limits

# Both animations are conventional six-second looping GIFs. Their frame
# intervals adapt to the number of saved outputs.
gif_duration_s <- 6
# eta_00000 is the zero-time initial condition, not an evolved model state.
# It is retained in eta_max but deliberately excluded from the full GIF.
full_animation_indices <- which(eta_time > 0)
full_gif_interval <- gif_duration_s / max(1, length(full_animation_indices))
final_sixth_indices <- which(eta_time >= final_sixth_start_s)
final_sixth_gif_interval <- gif_duration_s / max(1, length(final_sixth_indices))
full_animation_saved_start_label <- if (length(full_animation_indices)) {
  elapsed_clock(min(eta_time[full_animation_indices]))
} else {
  "n/a"
}
full_animation_saved_end_label <- if (length(full_animation_indices)) {
  elapsed_clock(max(eta_time[full_animation_indices]))
} else {
  "n/a"
}
final_sixth_saved_start_label <- if (length(final_sixth_indices)) {
  elapsed_clock(min(eta_time[final_sixth_indices]))
} else {
  "n/a"
}
final_sixth_saved_end_label <- if (length(final_sixth_indices)) {
  elapsed_clock(max(eta_time[final_sixth_indices]))
} else {
  "n/a"
}

# Write real GIF files before the HTML is assembled. This is more reliable than
# Knitr's animation hook, which can leave an empty figure placeholder when the
# report is moved to the Pages output directory.
site_assets_dir <- file.path(report_site_dir, "assets")
dir.create(site_assets_dir, recursive = TRUE, showWarnings = FALSE)
full_gif_url <- "assets/port-fairy-full.gif"
final_sixth_gif_url <- "assets/port-fairy-final-sixth.gif"

draw_eta_frame <- function(k) {
  r_om <- eta_public_frames[[k]]
  r_ll <- project(r_om, "EPSG:4326", method = "bilinear")
  r_ll <- apply_water_masks_wgs84(r_ll)
  plot_geographic(r_ll,
                  main = sprintf("Free-surface elevation: t = %.1f s (WGS84)", eta_time[k]),
                  col = eta_animation_palette,
                  range = eta_animation_limits, show_legend = TRUE)
  add_wave_frame_annotation(r_ll, eta_time[k])
}

write_eta_gif <- function(indices, file_name, interval) {
  if (!length(indices)) return(NULL)
  gif_path <- file.path(site_assets_dir, file_name)
  gifski::save_gif(
    expr = for (k in indices) draw_eta_frame(k),
    gif_file = gif_path, width = 800, height = 600, res = 100,
    delay = interval, loop = TRUE, progress = FALSE
  )
  if (!file.exists(gif_path) || file.info(gif_path)$size < 1000) {
    stop("GIF was not written: ", gif_path)
  }
  gif_path
}

full_gif_path <- write_eta_gif(full_animation_indices, "port-fairy-full.gif",
                               full_gif_interval)
final_sixth_gif_path <- write_eta_gif(final_sixth_indices,
                                      "port-fairy-final-sixth.gif",
                                      final_sixth_gif_interval)
```

```{r report-table-style, results='asis', echo=FALSE}
cat('<style>
.pf-report-table { width: 100%; margin: 0.8rem 0 1.5rem; border-collapse: collapse; font-size: 0.93rem; }
.pf-report-table th { background: #2c3e50; color: #fff; text-align: left; padding: 0.55rem 0.7rem; }
.pf-report-table td { padding: 0.45rem 0.7rem; border-bottom: 1px solid #dfe6e9; vertical-align: top; }
.pf-report-table td:first-child { width: 34%; color: #2c3e50; font-weight: 600; }
.pf-report-table tbody tr:nth-child(even) { background: #f6f8fa; }
</style>')
```

## Interpretation first

This is a `r paste0(metres_label(grid$dx_m[1]), " m")`, `r simulation_label`
grid and forcing diagnostic—not a validated harbour forecast. Each raw
FUNWAVE text field is first rebuilt on the exact rotated Oblique Mercator grid,
DEM-masked to water cells, and masked seaward of the internal wavemaker. The
historical workflow aggregates report maps, animations and GeoTIFFs to
`r paste0(public_grid_tag, " cells")` in that rotated CRS, then reprojects them
to WGS84 for geographic display. Current paths remain traced on native model
cells before being projected to WGS84. The interactive maps are limited to the
local model extent and use a compact display height.

Live report: <https://julianog.github.io/run_surf_wave_forecast/>. Source:
<https://github.com/JulianOG/run_surf_wave_forecast>.

```{r forcing-status}
direction_status <- if (forcing$inward_source_component > 0) {
  "The wave vector has an inward component from the selected source edge."
} else {
  "Warning: the wave vector points out of the selected source edge. Do not interpret this run as a coastal-wave prediction."
}
report_table(transpose_summary(data.frame(
  source_edge = grid$source_edge,
  boundary_direction_statistic = forcing$boundary_direction_statistic,
  requested_model_x_bearing_deg_true = forcing$requested_model_x_bearing_deg_true,
  realised_model_x_bearing_deg_true = forcing$model_x_bearing_deg_true,
  funwave_theta_peak_deg = forcing$funwave_theta_peak_deg,
  inward_source_component = forcing$inward_source_component,
  interpretation = direction_status
), digits = 2))
```

## Latest buoy forcing

The irregular `WK_IRR` source is a TMA/JONSWAP-style parameterisation made from
the newest usable IMOS Spotter observation. It uses the buoy **mean** direction
and matching mean directional spread when present (`SSWMD` and `WMDS`), falling
back to the peak values only when the mean values are unavailable. The buoy
provides integral Hs and period parameters—not a phase-resolved spectrum.

```{r forcing}
report_table(transpose_summary(forcing, digits = 2))
```

## Latest free-surface elevation

```{r eta, echo=FALSE}
if (is.null(eta)) {
  plot.new(); text(0.5, 0.5, "No eta output was produced.")
} else {
  eta_plot <- apply_output_visibility_mask_model(eta)
  if (!is.null(mask)) eta_plot[mask <= 0] <- NA_real_
  eta_rotated <- model_to_rotated_raster(eta_plot, "eta")
  eta_wgs84 <- write_model_output_rasters(eta_rotated, "eta_latest")
  eta_wgs84 <- apply_water_masks_wgs84(eta_wgs84)
  writeRaster(eta_wgs84, file.path(case_dir, "eta_latest_wgs84.tif"), overwrite = TRUE)
  leaflet_model_map(eta_wgs84, eta_animation_palette,
                    eta_display_limits, "Free-surface elevation (m)")
}
```

## Free-surface animations

Both GIFs use a shared colour scale and are written as standalone assets from
this Rmd, after each raw output has been reconstructed on the rotated FUNWAVE
grid and projected to WGS84. The zero-time initial condition is
intentionally omitted from the full animation. Each GIF loops for six seconds:
one represents the complete `r simulation_label` run (saved frames
`r full_animation_saved_start_label`--`r full_animation_saved_end_label`), and
the other shows the final one-sixth (configured `r final_sixth_start_label`--`r simulation_end_label`;
saved frames `r final_sixth_saved_start_label`--`r final_sixth_saved_end_label`).

### Full `r simulation_label` simulation: saved `r full_animation_saved_start_label`--`r full_animation_saved_end_label`

```{r eta-animation-full}
if (is.null(full_gif_path)) {
  plot.new(); text(0.5, 0.5, "No evolved eta output was produced.")
} else {
  htmltools::tags$img(src = full_gif_url, alt = "Full Port Fairy free-surface animation",
                       style = "max-width: 100%; height: auto; display: block;")
}
```

### Final one-sixth: saved `r final_sixth_saved_start_label`--`r final_sixth_saved_end_label`

```{r eta-animation-final-sixth}
if (is.null(final_sixth_gif_path)) {
  plot.new(); text(0.5, 0.5, "No final-one-sixth eta output was produced.")
} else {
  htmltools::tags$img(src = final_sixth_gif_url,
                       alt = "Final one-sixth Port Fairy free-surface animation",
                       style = "max-width: 100%; height: auto; display: block;")
}
```

## Maximum free-surface elevation

This is the maximum positive instantaneous `eta` at each cell. It is useful for
seeing crest reach, but it is **not** a wave-height statistic.

```{r hmax-eta, echo=FALSE}
if (is.null(eta_max)) {
  plot.new(); text(0.5, 0.5, "No eta output was produced.")
} else {
  eta_max_wgs84 <- write_model_output_rasters(eta_max, "eta_max", already_public = TRUE)
  eta_max_wgs84 <- apply_water_masks_wgs84(eta_max_wgs84)
  writeRaster(eta_max_wgs84, file.path(case_dir, "eta_max_wgs84.tif"), overwrite = TRUE)
  leaflet_model_map(eta_max_wgs84, ygnbu_palette,
                    eta_max_display_limits, "Maximum free-surface elevation (m)")
}
```

## Peak simulated significant wave height

`WaveHeight = T` is evaluated by FUNWAVE's mean-wave diagnostic after the
configured spin-up. The map below is the largest `Hsig` value at each cell from
the saved mean-wave outputs, rather than a maximum of individual `eta` frames.

```{r hsig-max, echo=FALSE}
if (is.null(hsig_max)) {
  plot.new(); text(0.5, 0.5, "No Hsig output was produced. Check STEADY_TIME and T_INTV_mean.")
} else {
  hsig_max_wgs84 <- write_model_output_rasters(hsig_max, "hsig_max", already_public = TRUE)
  hsig_max_wgs84 <- apply_water_masks_wgs84(hsig_max_wgs84)
  writeRaster(hsig_max_wgs84, file.path(case_dir, "hsig_max_wgs84.tif"), overwrite = TRUE)
  leaflet_model_map(hsig_max_wgs84, ygnbu_palette,
                    hsig_display_limits, "Peak simulated significant wave height (m)")
}
```

## Data acknowledgement and citation

Buoy data were sourced from the **Integrated Marine Observing System (IMOS)**,
which is enabled by the National Collaborative Research Infrastructure Strategy
(NCRIS). The wave data were collected and quality controlled by the University
of Western Australia. This dataset is an output from the Catching Oz Waves
project, which received investment from the Australian Research Data Commons
(ARDC): <https://doi.org/10.47486/DP748>. ARDC is funded by NCRIS.

Suggested data citation: *Deakin University (year of data downloaded), Wave
buoys Observations -- Australia -- near real-time, downloaded from the IMOS URL
on the date of download.*

## Provenance

```{r provenance}
report_table(transpose_summary(grid, digits = 3))
```

## Buoy-to-nearshore eta time series

These are native FUNWAVE station outputs from five selected integer grid cells:
the buoy end is dark blue and the nearshore end is yellow. They are written
every second from the model state—there is no spatial or temporal interpolation.

```{r station-timeseries, echo=FALSE}
if (!nrow(station_timeseries)) {
  plot.new(); text(0.5, 0.5, "No FUNWAVE station output was produced.")
} else {
  htmltools::tags$iframe(
    src = "assets/port-fairy-transect-timeseries.html", width = "100%", height = "560",
    style = "border: 1px solid #dfe6e9; display: block;"
  )
}
```

## Animated current particles

This existing particle page now uses the latest FUNWAVE time-mean `Umean`/
`Vmean` field rather than phase-resolved `U`/`V` wave-orbital velocities. Its
paths are traced on native rotated grid cells with nearest-cell velocities,
then projected to WGS84. This removes the wave-period direction reversals that
made the particles wobble; animation speed remains visual only.

```{r current-particles, echo=FALSE}
htmltools::tags$p(
  htmltools::tags$a(
    "Open the full animated current-flow particle page",
    href = current_particle_page_url, target = "_blank",
    style = "font-size: 1.1rem; font-weight: 600;"
  )
)
```

## Earth-style current particles

This is an additional particle renderer inspired by the open-source
[earth.nullschool.net animator](https://github.com/cambecc/earth). It uses the
same latest `Umean`/`Vmean` field as the repaired page above, but advects and
fades particles directly in the browser to give continuous trails. The
browser's bilinear field interpolation is display-only; it does not alter
FUNWAVE outputs or the exact grid-cell transect series.

```{r earth-style-current-particles, echo=FALSE}
htmltools::tags$p(
  htmltools::tags$a(
    "Open the Earth-style mean-current particle page",
    href = earth_flow_page_url, target = "_blank",
    style = "font-size: 1.1rem; font-weight: 600;"
  )
)
report_table(earth_flow_debug)
```

## Streamline diagnostic

Option 1 is retained as a separate, inspectable streamline diagnostic. The
table exposes the `Umean`/`Vmean` files, selected mean-output record, velocity
threshold, seed count, and completed-path count so an empty result can be
diagnosed rather than silently displayed as a blank map.

```{r current-streamline-diagnostic, echo=FALSE}
report_table(current_debug)
if (!is.null(current_streamline_debug_url)) {
  htmltools::tags$p(
    htmltools::tags$a(
      "Open the option 1 streamline diagnostic page",
      href = current_streamline_debug_url, target = "_blank"
    )
  )
}
```
