---
title: "Port Fairy near-real-time FUNWAVE test"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 11, fig.height = 6)

suppressPackageStartupMessages({
  library(terra)
  library(leaflet)
})

case_dir <- normalizePath("../port_fairy/output", mustWork = TRUE)
results_dir <- file.path(case_dir, "results")
forcing <- read.csv(file.path(case_dir, "latest_buoy_forcing.csv"),
                    check.names = FALSE)
grid <- read.csv(file.path(case_dir, "grid_metadata.csv"), check.names = FALSE)
if (nrow(forcing) != 1 || nrow(grid) != 1) {
  stop("Expected one row in forcing and grid metadata files.")
}
if (!is.finite(grid$dx_m[1]) || !is.finite(grid$total_time_s[1]) ||
    !is.finite(grid$plot_intv_s[1]) || grid$dx_m[1] <= 0 ||
    grid$total_time_s[1] <= 0 || grid$plot_intv_s[1] <= 0) {
  stop("Grid metadata must contain positive dx_m, total_time_s and plot_intv_s values.")
}
grid_tag <- paste0(sprintf("%.0f", grid$dx_m[1]), "m")
simulation_minutes <- grid$total_time_s[1] / 60
final_sixth_start_s <- (5 / 6) * grid$total_time_s[1]
elapsed_clock <- function(elapsed_s) {
  sprintf("%02.0f:%02.0f", floor(elapsed_s / 60), round(elapsed_s %% 60))
}
simulation_label <- paste0(sprintf("%.0f", simulation_minutes), "-minute")
final_sixth_start_label <- elapsed_clock(final_sixth_start_s)
simulation_end_label <- elapsed_clock(grid$total_time_s[1])
eta_display_limits <- c(-5, 5)
eta_max_display_limits <- c(0, 5)
hsig_display_limits <- c(0, 4)
ygnbu_palette <- RColorBrewer::brewer.pal(9, "YlGnBu")
station_colours <- ygnbu_palette[c(2, 4, 6, 8, 9)]

pretty_field_name <- function(x) {
  x <- gsub("_", " ", x)
  x <- tools::toTitleCase(x)
  x <- gsub("\\bFunwave\\b", "FUNWAVE", x)
  x <- gsub("\\bHs\\b", "Hs", x)
  x <- gsub("\\bTp\\b", "Tp", x)
  x <- gsub("\\bMglob\\b", "Mglob", x)
  x <- gsub("\\bNglob\\b", "Nglob", x)
  x <- gsub("\\bDx\\b", "DX", x)
  x <- gsub("\\bDy\\b", "DY", x)
  x <- gsub("\\bWk\\b", "WK", x)
  x
}

transpose_summary <- function(x, digits = 2) {
  x <- as.data.frame(x, check.names = FALSE, stringsAsFactors = FALSE)
  if (nrow(x) != 1) stop("transpose_summary() expects a one-row data frame.")
  data.frame(
    Item = pretty_field_name(names(x)),
    Value = vapply(x[1, ], function(value) {
      if (is.numeric(value)) format(value, digits = digits, trim = TRUE, scientific = FALSE) else as.character(value)
    }, character(1)),
    check.names = FALSE
  )
}

report_table <- function(x, digits = 2) {
  knitr::kable(x, format = "html", digits = digits, escape = TRUE,
               table.attr = 'class="pf-report-table"')
}

# The buoys report UTC. Show the animation using Port Fairy's local civil time
# while retaining the model's elapsed time from that observation.
port_fairy_tz <- "Australia/Melbourne"
buoy_time_utc <- as.POSIXct(
  sub(" UTC$", "", forcing$time_utc[1]),
  format = "%Y-%m-%d %H:%M:%S", tz = "UTC"
)
if (is.na(buoy_time_utc)) stop("Could not parse the buoy observation time.")

# FUNWAVE writes text grids as Nglob rows (model y) by Mglob columns (model x).
# R's image() expects z[x, y], so transpose every model output exactly once.
as_model_xy <- function(z, label = "model field") {
  z <- as.matrix(z)
  if (all(dim(z) == c(grid$Nglob, grid$Mglob))) return(t(z))
  if (all(dim(z) == c(grid$Mglob, grid$Nglob))) return(z)
  stop(sprintf("%s has %.0f x %.0f values; expected %.0f x %.0f (y x x).",
               label, nrow(z), ncol(z), grid$Nglob, grid$Mglob))
}

read_text_model <- function(path, label = basename(path)) {
  as_model_xy(utils::read.table(path, header = FALSE), label)
}

read_field <- function(prefix, directory = results_dir) {
  candidates <- list.files(directory, pattern = paste0("^", prefix, "_[0-9]{5}$"),
                           full.names = TRUE)
  if (!length(candidates)) return(NULL)
  number <- as.integer(sub(paste0("^", prefix, "_"), "", basename(candidates)))
  read_text_model(candidates[which.max(number)], prefix)
}


read_all_fields <- function(prefix, directory = results_dir) {
  candidates <- list.files(directory, pattern = paste0("^", prefix, "_[0-9]{5}$"),
                           full.names = TRUE)
  if (!length(candidates)) return(list())
  number <- as.integer(sub(paste0("^", prefix, "_"), "", basename(candidates)))
  candidates <- candidates[order(number)]
  lapply(candidates, read_text_model, label = prefix)
}

maximum_model_field <- function(fields) {
  if (!length(fields)) return(NULL)
  out <- fields[[1]]
  if (length(fields) > 1) {
    for (k in 2:length(fields)) out <- pmax(out, fields[[k]], na.rm = TRUE)
  }
  out[!is.finite(out)] <- NA_real_
  out
}

read_static_model <- function(name) {
  path <- file.path(results_dir, name)
  if (!file.exists(path)) return(NULL)
  read_text_model(path, name)
}

model_axes <- function(z) {
  list(x = (seq_len(nrow(z)) - 1) * grid$dx_m,
       y = (seq_len(ncol(z)) - 1) * grid$dy_m)
}

plot_model <- function(z, main, col, zlim = range(z, finite = TRUE), ...) {
  a <- model_axes(z)
  image(a$x, a$y, z, asp = 1, col = col, zlim = zlim,
        xlab = "Model x: source edge to coast (m)",
        ylab = "Model y (m)", main = main, ...)
  invisible(a)
}

model_to_rotated_raster <- function(z, label = "model field") {
  z <- as.matrix(z)
  expected <- c(grid$Mglob, grid$Nglob)
  if (!identical(dim(z), expected)) {
    stop(sprintf("%s has %.0f x %.0f values; expected Mglob x Nglob = %.0f x %.0f.",
                 label, nrow(z), ncol(z), expected[1], expected[2]))
  }

  # z is [model x, model y]. Convert it back to terra's native
  # [north-to-south row, west-to-east column] layout using the exact inverse
  # of the transforms in setup_port_fairy_funwave.R.
  gis_rows <- t(z)
  if (source_is_low_x) {
    gis_rows <- gis_rows[nrow(gis_rows):1, , drop = FALSE]
  } else {
    gis_rows <- gis_rows[, ncol(gis_rows):1, drop = FALSE]
  }

  out <- rast(depth_raster)
  values(out) <- as.vector(t(gis_rows))
  out
}

symmetric_raster_limits <- function(r) {
  m <- global(abs(r), "max", na.rm = TRUE)[1, 1]
  if (!is.finite(m) || m == 0) m <- 1e-8
  c(-m, m)
}

write_model_output_rasters <- function(r_om, stem, method = "bilinear") {
  # Keep water-level products on cells that are water in the DEM before any
  # geographic reprojection. A second nearest-neighbour mask is applied after
  # projection to keep interpolation from bleeding values onto land pixels.
  r_om <- apply_output_visibility_mask_rotated(r_om)
  writeRaster(r_om, file.path(case_dir, paste0(stem, "_rotated_omerc.tif")),
              overwrite = TRUE)
  r_ll <- project(r_om, "EPSG:4326", method = method)
  writeRaster(r_ll, file.path(case_dir, paste0(stem, "_wgs84.tif")),
              overwrite = TRUE)
  r_ll
}


symmetric_limits <- function(z) {
  m <- max(abs(z), na.rm = TRUE)
  if (!is.finite(m) || m == 0) m <- 1e-8
  c(-m, m)
}

positive_limits <- function(z) {
  m <- max(z, na.rm = TRUE)
  if (!is.finite(m) || m <= 0) m <- 1e-8
  c(0, m)
}

depth_raster <- rast(file.path(
  case_dir, paste0("depth_", grid_tag, "_positive_water_depth.tif")
))
elevation_raster <- rast(file.path(
  case_dir, paste0("elevation_", grid_tag, "_warped.tif")
))
depth_from_file <- read_text_model(file.path(case_dir, "depth.txt"), "depth.txt")
source_is_low_x <- identical(grid$source_edge, "minimum rotated x")

# This mask comes from the warped DEM supplied to FUNWAVE, not from the
# model's dynamic wet/dry field. It deliberately limits displayed water levels
# to DEM cells with positive water depth.
dem_water_model <- is.finite(depth_from_file) & depth_from_file > 0
dem_water_mask_rotated <- ifel(depth_raster > 0, 1, 0)
dem_water_mask_wgs84 <- project(dem_water_mask_rotated, "EPSG:4326", method = "near")

apply_dem_water_mask_model <- function(z) {
  z <- as.matrix(z)
  z[!dem_water_model] <- NA_real_
  z
}

apply_dem_water_mask_rotated <- function(r_om) {
  r_om[is.na(dem_water_mask_rotated) | dem_water_mask_rotated <= 0] <- NA_real_
  r_om
}

# The internal wavemaker emits energy in both directions. The buoy-side side
# is intentionally hidden from coastal forecast products before projection and
# again afterwards, preventing interpolation from bleeding across the paddle.
shoreward_of_paddle_model <- outer(
  (seq_len(grid$Mglob[1]) - 1) * grid$dx_m[1], seq_len(grid$Nglob[1]),
  function(x, y) x >= grid$x_wk_m[1]
)
shoreward_of_paddle_rotated <- model_to_rotated_raster(
  ifelse(shoreward_of_paddle_model, 1, 0), "shoreward-of-wavemaker mask"
)
shoreward_of_paddle_wgs84 <- project(shoreward_of_paddle_rotated, "EPSG:4326", method = "near")

apply_output_visibility_mask_model <- function(z) {
  z <- apply_dem_water_mask_model(z)
  z[!shoreward_of_paddle_model] <- NA_real_
  z
}

apply_output_visibility_mask_rotated <- function(r_om) {
  r_om <- apply_dem_water_mask_rotated(r_om)
  r_om[is.na(shoreward_of_paddle_rotated) | shoreward_of_paddle_rotated <= 0] <- NA_real_
  r_om
}

# Reconstruct depth.txt on the exact native Omerc raster, then derive maps
# from that raster. This is the same pathway used for FUNWAVE outputs below.
depth_from_file_rotated <- model_to_rotated_raster(depth_from_file, "depth.txt")
depth_file_minus_raster <- depth_from_file_rotated - depth_raster
max_depth_file_delta <- global(abs(depth_file_minus_raster), "max", na.rm = TRUE)[1, 1]

buoy_ll <- vect(matrix(c(forcing$buoy_lon, forcing$buoy_lat), ncol = 2),
                type = "points", crs = "EPSG:4326")
buoy_om <- project(buoy_ll, crs(depth_raster))

# Plot all map products in WGS84, but only after reconstructing them on the
# native rotated grid. The black outline is the rectangular FUNWAVE domain.
outline_from_raster <- function(r) {
  vect(rbind(
    c(xmin(r), ymin(r)), c(xmax(r), ymin(r)),
    c(xmax(r), ymax(r)), c(xmin(r), ymax(r)),
    c(xmin(r), ymin(r))
  ), type = "polygons", crs = crs(r))
}
model_outline_om <- outline_from_raster(depth_raster)
model_outline_ll_xy <- crds(project(model_outline_om, "EPSG:4326"))

# Model x is always written from the buoy-side source edge toward the coast.
# Draw the actual internal-wavemaker line on every map so the reconstructed
# fields can be checked against their physical source location.
source_x_om <- if (source_is_low_x) {
  xmin(depth_raster) + grid$x_wk_m
} else {
  xmax(depth_raster) - grid$x_wk_m
}
source_line_om <- vect(rbind(
  c(source_x_om, ymin(depth_raster)),
  c(source_x_om, ymax(depth_raster))
), type = "lines", crs = crs(depth_raster))
source_line_ll_xy <- crds(project(source_line_om, "EPSG:4326"))

plot_geographic <- function(r_ll, ..., show_legend = FALSE, range = NULL) {
  # terra::plot uses range= to fix its colour scale; zlim= is for image().
  if (is.null(range)) plot(r_ll, ...) else plot(r_ll, ..., range = range)
  lines(model_outline_ll_xy[, 1], model_outline_ll_xy[, 2],
        col = "black", lwd = 2)
  lines(source_line_ll_xy[, 1], source_line_ll_xy[, 2],
        col = "dodgerblue3", lwd = 2, lty = 2)
  points(crds(buoy_ll)[1, 1], crds(buoy_ll)[1, 2],
         pch = 16, col = "red", cex = 1.25)
  if (show_legend) {
    legend("bottomleft",
           legend = c("FUNWAVE grid", "internal wavemaker", "buoy"),
           col = c("black", "dodgerblue3", "red"), lwd = c(2, 2, NA),
           lty = c(1, 2, NA), pch = c(NA, NA, 16),
           bty = "n", cex = 0.75)
  }
}

local_map_bounds <- function(r_ll, padding = 0.04) {
  e <- ext(r_ll)
  x_pad <- (xmax(e) - xmin(e)) * padding
  y_pad <- (ymax(e) - ymin(e)) * padding
  c(west = xmin(e) - x_pad, south = ymin(e) - y_pad,
    east = xmax(e) + x_pad, north = ymax(e) + y_pad)
}

apply_water_masks_wgs84 <- function(r_ll) {
  dem_mask <- dem_water_mask_wgs84
  if (!compareGeom(dem_mask, r_ll, stopOnError = FALSE)) {
    dem_mask <- resample(dem_mask, r_ll, method = "near")
  }
  r_ll[is.na(dem_mask) | dem_mask <= 0] <- NA_real_

  paddle_mask <- shoreward_of_paddle_wgs84
  if (!compareGeom(paddle_mask, r_ll, stopOnError = FALSE)) {
    paddle_mask <- resample(paddle_mask, r_ll, method = "near")
  }
  r_ll[is.na(paddle_mask) | paddle_mask <= 0] <- NA_real_

  if (is.null(mask_wgs84)) return(r_ll)
  wet_mask <- mask_wgs84
  if (!compareGeom(wet_mask, r_ll, stopOnError = FALSE)) {
    wet_mask <- resample(wet_mask, r_ll, method = "near")
  }
  r_ll[is.na(wet_mask) | wet_mask <= 0] <- NA_real_
  r_ll
}

leaflet_model_map <- function(r_ll, palette, limits, legend_title,
                              station_plot_url = NULL) {
  b <- local_map_bounds(r_ll)
  pal <- colorNumeric(palette = palette, domain = limits, na.color = "transparent")
  m <- leaflet(width = "100%", height = 480,
               options = leafletOptions(minZoom = 14, maxZoom = 18,
                                        maxBoundsViscosity = 1))
  # These public sources require no project key.
  m <- addProviderTiles(m, "OpenStreetMap.Mapnik", group = "OpenStreetMap")
  m <- addProviderTiles(m, "Esri.WorldImagery", group = "Satellite imagery")
  m <- setMaxBounds(m, b[["west"]], b[["south"]], b[["east"]], b[["north"]])
  m <- fitBounds(m, b[["west"]], b[["south"]], b[["east"]], b[["north"]])
  m <- addRasterImage(m, r_ll, colors = pal, opacity = 0.88, project = FALSE,
                      group = "Model output")
  m <- addPolylines(m, lng = model_outline_ll_xy[, 1], lat = model_outline_ll_xy[, 2],
                    color = "black", weight = 2, group = "FUNWAVE grid")
  m <- addPolylines(m, lng = source_line_ll_xy[, 1], lat = source_line_ll_xy[, 2],
                    color = "dodgerblue3", weight = 2, dashArray = "6 6",
                    group = "Internal wavemaker")
  m <- addCircleMarkers(m, lng = crds(buoy_ll)[1, 1], lat = crds(buoy_ll)[1, 2],
                        radius = 5, color = "red", fillOpacity = 1,
                        label = "Port Fairy wave buoy", group = "Buoy")
  if (exists("station_metadata", inherits = TRUE) && nrow(station_metadata)) {
    m <- addCircleMarkers(
      m, lng = station_metadata$lon, lat = station_metadata$lat,
      radius = 4, color = station_colours[seq_len(nrow(station_metadata))],
      fillOpacity = 1, label = station_metadata$station_id,
      group = "Transect stations"
    )
  }
  m <- addLayersControl(
    m,
    baseGroups = c("OpenStreetMap", "Satellite imagery"),
    overlayGroups = c("Model output", "FUNWAVE grid", "Internal wavemaker", "Buoy", "Transect stations"),
    options = layersControlOptions(collapsed = FALSE)
  )
  m <- addLegend(m, position = "bottomright", pal = pal, values = limits,
                 title = legend_title, opacity = 0.88)
  if (!is.null(station_plot_url)) {
    m <- addControl(
      m,
      html = htmltools::tags$details(
        style = "background: rgba(255,255,255,0.94); padding: 6px; max-width: 470px;",
        htmltools::tags$summary("Show buoy-to-nearshore eta time series"),
        htmltools::tags$iframe(
          src = station_plot_url, width = "450", height = "310",
          style = "border: 0; display: block; margin-top: 6px;"
        )
      ), position = "topright"
    )
  }
  m
}

frame_local_time <- function(elapsed_s) {
  format(buoy_time_utc + elapsed_s, "%d %b %Y %H:%M %Z", tz = port_fairy_tz)
}

add_wave_frame_annotation <- function(r_ll, elapsed_s) {
  e <- ext(r_ll)
  x_span <- xmax(e) - xmin(e)
  y_span <- ymax(e) - ymin(e)
  bearing_to <- forcing$boundary_direction_to_deg_true[1]
  direction_statistic <- forcing$boundary_direction_statistic[1]
  angle <- bearing_to * pi / 180

  # The map is in longitude/latitude. Scale the longitude component by the
  # local latitude before drawing the true-north travel-direction arrow.
  arrow_y <- 0.13 * y_span
  arrow_x <- arrow_y * sin(angle) / cos(mean(c(ymin(e), ymax(e))) * pi / 180)
  arrow_dy <- arrow_y * cos(angle)
  arrow_x0 <- if (arrow_x >= 0) xmin(e) + 0.08 * x_span else xmax(e) - 0.08 * x_span
  arrow_y0 <- if (arrow_dy >= 0) ymin(e) + 0.10 * y_span else ymax(e) - 0.10 * y_span
  arrows(arrow_x0, arrow_y0, arrow_x0 + arrow_x, arrow_y0 + arrow_dy,
         col = "red", lwd = 3, length = 0.10)

  label <- c(
    paste("Local:", frame_local_time(elapsed_s)),
    sprintf("Model +%02.0f:%04.1f", floor(elapsed_s / 60), elapsed_s %% 60),
    sprintf("Buoy Hs %.2f m | Tp %.1f s", forcing$hs_m[1], forcing$tp_s[1]),
    sprintf("%s waves from %.0f° (arrow travels to %.0f°)",
            tools::toTitleCase(direction_statistic),
            forcing$boundary_direction_from_deg_true[1], bearing_to)
  )
  legend("topright", legend = label, bty = "o", bg = rgb(1, 1, 1, 0.82),
         cex = 0.74, text.col = "black")
}

elevation_wgs84 <- project(elevation_raster, "EPSG:4326", method = "bilinear")
depth_file_wgs84 <- project(depth_from_file_rotated, "EPSG:4326", method = "bilinear")
depth_delta_wgs84 <- project(depth_file_minus_raster, "EPSG:4326", method = "bilinear")

eta <- read_field("eta")
mask <- read_field("mask")
mask_wgs84 <- NULL
if (!is.null(mask)) {
  mask_rotated <- model_to_rotated_raster(mask, "mask")
  mask_wgs84 <- project(mask_rotated, "EPSG:4326", method = "near")
}

eta_paths <- list.files(results_dir, pattern = "^eta_[0-9]{5}$",
                        full.names = TRUE)
# FUNWAVE uses sequential output-file numbers in eta_00000, eta_00001, ...;
# they are not elapsed seconds. Convert those numbers using PLOT_INTV so the
# animation labels and the final-one-sixth selection use model time.
eta_file_number <- as.integer(sub("^eta_", "", basename(eta_paths)))
eta_order <- order(eta_file_number)
eta_paths <- eta_paths[eta_order]
eta_file_number <- eta_file_number[eta_order]
eta_time <- eta_file_number * grid$plot_intv_s
if (length(eta_paths) &&
    (any(!is.finite(eta_time)) || any(eta_time < 0) ||
     max(eta_time) > grid$total_time_s[1] + grid$plot_intv_s[1] ||
     max(eta_time) < grid$total_time_s[1] - grid$plot_intv_s[1])) {
  stop("eta file numbers do not map to the configured simulation period.")
}
eta_frames <- lapply(eta_paths, function(path) {
  z <- read_text_model(path, basename(path))
  if (!is.null(mask)) z[mask <= 0] <- NA_real_
  apply_output_visibility_mask_model(z)
})
eta_max <- if (length(eta_frames)) {
  do.call(pmax, c(eta_frames, na.rm = TRUE))
} else {
  NULL
}
if (!is.null(eta_max)) eta_max[!is.finite(eta_max)] <- NA_real_

# FUNWAVE's WaveHeight=T diagnostic is written by PREVIEW_MEAN as Hsig files.
# It uses the model time step internally, so it is the correct way to assess
# wave energy rather than comparing individual 15-second eta snapshots.
hsig_paths <- list.files(results_dir, pattern = "^Hsig_[0-9]{5}$",
                         full.names = TRUE)
hsig_number <- as.integer(sub("^Hsig_", "", basename(hsig_paths)))
hsig_paths <- hsig_paths[order(hsig_number)]
hsig_frames <- lapply(hsig_paths, function(path) {
  z <- read_text_model(path, basename(path))
  if (!is.null(mask)) z[mask <= 0] <- NA_real_
  apply_output_visibility_mask_model(z)
})
hsig_max <- if (length(hsig_frames)) {
  do.call(pmax, c(hsig_frames, na.rm = TRUE))
} else {
  NULL
}
if (!is.null(hsig_max)) hsig_max[!is.finite(hsig_max)] <- NA_real_

# Read the five FUNWAVE station files.  Each file is time, eta, u, v and the
# position metadata was created with stations.txt by the case setup script.
station_metadata_path <- file.path(case_dir, "station_metadata.csv")
station_metadata <- if (file.exists(station_metadata_path)) {
  read.csv(station_metadata_path, check.names = FALSE)
} else {
  data.frame()
}
station_timeseries <- data.frame()
station_plot_url <- NULL
if (nrow(station_metadata)) {
  station_series <- lapply(seq_len(nrow(station_metadata)), function(k) {
    path <- file.path(results_dir, sprintf("sta_%04d", k))
    if (!file.exists(path)) return(NULL)
    z <- utils::read.table(path, header = FALSE)
    if (ncol(z) < 2) return(NULL)
    names(z)[seq_len(min(4, ncol(z)))] <- c("time_s", "eta_m", "u_m_s", "v_m_s")[seq_len(min(4, ncol(z)))]
    z$station_id <- station_metadata$station_id[k]
    z
  })
  station_series <- Filter(Negate(is.null), station_series)
  if (length(station_series)) station_timeseries <- do.call(rbind, station_series)
}
if (nrow(station_timeseries)) {
  station_plot <- plotly::plot_ly()
  for (k in seq_len(nrow(station_metadata))) {
    station_name <- station_metadata$station_id[k]
    z <- station_timeseries[station_timeseries$station_id == station_name, , drop = FALSE]
    if (nrow(z)) {
      station_plot <- plotly::add_lines(
        station_plot, data = z, x = ~time_s, y = ~eta_m, name = station_name,
        line = list(color = station_colours[k], width = 2),
        hovertemplate = paste0(station_name, "<br>t = %{x:.1f} s<br>eta = %{y:.3f} m<extra></extra>")
      )
    }
  }
  station_plot <- plotly::layout(
    station_plot, title = "FUNWAVE eta along buoy-to-nearshore transect",
    xaxis = list(title = "Model elapsed time (s)"),
    yaxis = list(title = "Free-surface elevation, eta (m)", range = eta_display_limits),
    legend = list(orientation = "h")
  )
  site_assets_dir <- normalizePath("../site/assets", mustWork = FALSE)
  dir.create(site_assets_dir, recursive = TRUE, showWarnings = FALSE)
  htmlwidgets::saveWidget(
    station_plot, file.path(site_assets_dir, "port-fairy-transect-timeseries.html"),
    selfcontained = FALSE
  )
  station_plot_url <- "assets/port-fairy-transect-timeseries.html"
}

# A single fixed range makes every eta frame and map comparable across days.
eta_animation_limits <- eta_display_limits

# Both animations are conventional six-second looping GIFs. Their frame
# intervals adapt to the number of saved outputs.
gif_duration_s <- 6
# eta_00000 is the zero-time initial condition, not an evolved model state.
# It is retained in eta_max but deliberately excluded from the full GIF.
full_animation_indices <- which(eta_time > 0)
full_gif_interval <- gif_duration_s / max(1, length(full_animation_indices))
final_sixth_indices <- which(eta_time >= final_sixth_start_s)
final_sixth_gif_interval <- gif_duration_s / max(1, length(final_sixth_indices))
full_animation_saved_start_label <- if (length(full_animation_indices)) {
  elapsed_clock(min(eta_time[full_animation_indices]))
} else {
  "n/a"
}
full_animation_saved_end_label <- if (length(full_animation_indices)) {
  elapsed_clock(max(eta_time[full_animation_indices]))
} else {
  "n/a"
}
final_sixth_saved_start_label <- if (length(final_sixth_indices)) {
  elapsed_clock(min(eta_time[final_sixth_indices]))
} else {
  "n/a"
}
final_sixth_saved_end_label <- if (length(final_sixth_indices)) {
  elapsed_clock(max(eta_time[final_sixth_indices]))
} else {
  "n/a"
}
```

```{r report-table-style, results='asis', echo=FALSE}
cat('<style>
.pf-report-table { width: 100%; margin: 0.8rem 0 1.5rem; border-collapse: collapse; font-size: 0.93rem; }
.pf-report-table th { background: #2c3e50; color: #fff; text-align: left; padding: 0.55rem 0.7rem; }
.pf-report-table td { padding: 0.45rem 0.7rem; border-bottom: 1px solid #dfe6e9; vertical-align: top; }
.pf-report-table td:first-child { width: 34%; color: #2c3e50; font-weight: 600; }
.pf-report-table tbody tr:nth-child(even) { background: #f6f8fa; }
</style>')
```

## Interpretation first

This is a `r sprintf("%.0f m", grid$dx_m[1])`, `r simulation_label` grid and
forcing diagnostic—not a validated harbour
forecast. Each raw FUNWAVE text field is first rebuilt on the exact rotated
Oblique Mercator grid, DEM-masked to water cells, and masked seaward of the
internal wavemaker, then reprojected to WGS84
for geographic plotting. The interactive maps are limited to the local model
extent and use a compact display height.

Live report: <https://julianog.github.io/run_surf_wave_forecast/>. Source:
<https://github.com/JulianOG/run_surf_wave_forecast>.

```{r forcing-status}
direction_status <- if (forcing$inward_source_component > 0) {
  "The wave vector has an inward component from the selected source edge."
} else {
  "Warning: the wave vector points out of the selected source edge. Do not interpret this run as a coastal-wave prediction."
}
report_table(transpose_summary(data.frame(
  source_edge = grid$source_edge,
  boundary_direction_statistic = forcing$boundary_direction_statistic,
  requested_model_x_bearing_deg_true = forcing$requested_model_x_bearing_deg_true,
  realised_model_x_bearing_deg_true = forcing$model_x_bearing_deg_true,
  funwave_theta_peak_deg = forcing$funwave_theta_peak_deg,
  inward_source_component = forcing$inward_source_component,
  interpretation = direction_status
), digits = 2))
```

## Latest buoy forcing

The irregular `WK_IRR` source is a TMA/JONSWAP-style parameterisation made from
the newest usable IMOS Spotter observation. It uses the buoy **mean** direction
and matching mean directional spread when present (`SSWMD` and `WMDS`), falling
back to the peak values only when the mean values are unavailable. The buoy
provides integral Hs and period parameters—not a phase-resolved spectrum.

```{r forcing}
report_table(transpose_summary(forcing, digits = 2))
```

## Latest free-surface elevation

```{r eta, echo=FALSE}
if (is.null(eta)) {
  plot.new(); text(0.5, 0.5, "No eta output was produced.")
} else {
  eta_plot <- apply_output_visibility_mask_model(eta)
  if (!is.null(mask)) eta_plot[mask <= 0] <- NA_real_
  eta_rotated <- model_to_rotated_raster(eta_plot, "eta")
  eta_wgs84 <- write_model_output_rasters(eta_rotated, "eta_latest")
  eta_wgs84 <- apply_water_masks_wgs84(eta_wgs84)
  writeRaster(eta_wgs84, file.path(case_dir, "eta_latest_wgs84.tif"), overwrite = TRUE)
  leaflet_model_map(eta_wgs84, ygnbu_palette,
                    eta_display_limits, "Free-surface elevation (m)",
                    station_plot_url = station_plot_url)
}
```

## Free-surface animations

Both GIFs use a shared colour scale. Knitr's gifski hook creates them directly
from this Rmd, after each raw output has been reconstructed on the rotated
FUNWAVE grid and projected to WGS84. The zero-time initial condition is
intentionally omitted from the full animation. Each GIF loops for six seconds:
one represents the complete `r simulation_label` run (saved frames
`r full_animation_saved_start_label`--`r full_animation_saved_end_label`), and
the other shows the final one-sixth (configured `r final_sixth_start_label`--`r simulation_end_label`;
saved frames `r final_sixth_saved_start_label`--`r final_sixth_saved_end_label`).

### Full `r simulation_label` simulation: saved `r full_animation_saved_start_label`--`r full_animation_saved_end_label`

```{r eta-animation-full, fig.show='animate', animation.hook='gifski', interval=full_gif_interval, aniopts='loop', fig.width=8, fig.height=6, dpi=100}
if (!length(full_animation_indices)) {
  plot.new(); text(0.5, 0.5, "No evolved eta output was produced.")
} else {
  for (k in full_animation_indices) {
    r_om <- model_to_rotated_raster(eta_frames[[k]], basename(eta_paths[k]))
    r_om <- apply_output_visibility_mask_rotated(r_om)
    r_ll <- project(r_om, "EPSG:4326", method = "bilinear")
    r_ll <- apply_water_masks_wgs84(r_ll)
    plot_geographic(r_ll,
                    main = sprintf("Free-surface elevation: t = %.1f s (WGS84)", eta_time[k]),
                    col = ygnbu_palette,
                    range = eta_animation_limits, show_legend = TRUE)
    add_wave_frame_annotation(r_ll, eta_time[k])
  }
}
```

### Final one-sixth: saved `r final_sixth_saved_start_label`--`r final_sixth_saved_end_label`

```{r eta-animation-final-sixth, fig.show='animate', animation.hook='gifski', interval=final_sixth_gif_interval, aniopts='loop', fig.width=8, fig.height=6, dpi=100}
if (!length(final_sixth_indices)) {
  plot.new(); text(0.5, 0.5, "No final-one-sixth eta output was produced.")
} else {
  for (k in final_sixth_indices) {
    r_om <- model_to_rotated_raster(eta_frames[[k]], basename(eta_paths[k]))
    r_om <- apply_output_visibility_mask_rotated(r_om)
    r_ll <- project(r_om, "EPSG:4326", method = "bilinear")
    r_ll <- apply_water_masks_wgs84(r_ll)
    plot_geographic(r_ll,
                    main = sprintf("Free-surface elevation: t = %.1f s (WGS84)", eta_time[k]),
                    col = ygnbu_palette,
                    range = eta_animation_limits, show_legend = TRUE)
    add_wave_frame_annotation(r_ll, eta_time[k])
  }
}
```

## Maximum free-surface elevation

This is the maximum positive instantaneous `eta` at each cell. It is useful for
seeing crest reach, but it is **not** a wave-height statistic.

```{r hmax-eta, echo=FALSE}
if (is.null(eta_max)) {
  plot.new(); text(0.5, 0.5, "No eta output was produced.")
} else {
  eta_max_rotated <- model_to_rotated_raster(eta_max, "eta maximum")
  eta_max_wgs84 <- write_model_output_rasters(eta_max_rotated, "eta_max")
  eta_max_wgs84 <- apply_water_masks_wgs84(eta_max_wgs84)
  writeRaster(eta_max_wgs84, file.path(case_dir, "eta_max_wgs84.tif"), overwrite = TRUE)
  leaflet_model_map(eta_max_wgs84, ygnbu_palette,
                    eta_max_display_limits, "Maximum free-surface elevation (m)")
}
```

## Peak simulated significant wave height

`WaveHeight = T` is evaluated by FUNWAVE's mean-wave diagnostic after the
configured spin-up. The map below is the largest `Hsig` value at each cell from
the saved mean-wave outputs, rather than a maximum of individual `eta` frames.

```{r hsig-max, echo=FALSE}
if (is.null(hsig_max)) {
  plot.new(); text(0.5, 0.5, "No Hsig output was produced. Check STEADY_TIME and T_INTV_mean.")
} else {
  hsig_max_rotated <- model_to_rotated_raster(hsig_max, "Hsig maximum")
  hsig_max_wgs84 <- write_model_output_rasters(hsig_max_rotated, "hsig_max")
  hsig_max_wgs84 <- apply_water_masks_wgs84(hsig_max_wgs84)
  writeRaster(hsig_max_wgs84, file.path(case_dir, "hsig_max_wgs84.tif"), overwrite = TRUE)
  leaflet_model_map(hsig_max_wgs84, ygnbu_palette,
                    hsig_display_limits, "Peak simulated significant wave height (m)")
}
```

## Data acknowledgement and citation

Buoy data were sourced from the **Integrated Marine Observing System (IMOS)**,
which is enabled by the National Collaborative Research Infrastructure Strategy
(NCRIS). The wave data were collected and quality controlled by the University
of Western Australia. This dataset is an output from the Catching Oz Waves
project, which received investment from the Australian Research Data Commons
(ARDC): <https://doi.org/10.47486/DP748>. ARDC is funded by NCRIS.

Suggested data citation: *Deakin University (year of data downloaded), Wave
buoys Observations -- Australia -- near real-time, downloaded from the IMOS URL
on the date of download.*

## Provenance

```{r provenance}
report_table(transpose_summary(grid, digits = 3))
```
