---
title: "Port Fairy near-real-time FUNWAVE test"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 11, fig.height = 6)

suppressPackageStartupMessages({
  library(terra)
  library(leaflet)
})

case_dir <- normalizePath("../port_fairy/output", mustWork = TRUE)
# The normal forecast writes support files under site/assets. Historical runs
# set PORT_FAIRY_SITE_DIR to an isolated report folder so their maps/assets do
# not overwrite the live report's equivalents.
report_site_dir <- normalizePath(
  file.path("..", Sys.getenv("PORT_FAIRY_SITE_DIR", unset = "site")),
  mustWork = FALSE
)
results_dir <- file.path(case_dir, "results")
forcing <- read.csv(file.path(case_dir, "latest_buoy_forcing.csv"),
                    check.names = FALSE)
grid <- read.csv(file.path(case_dir, "grid_metadata.csv"), check.names = FALSE)
if (nrow(forcing) != 1 || nrow(grid) != 1) {
  stop("Expected one row in forcing and grid metadata files.")
}
if (!is.finite(grid$dx_m[1]) || !is.finite(grid$total_time_s[1]) ||
    !is.finite(grid$plot_intv_s[1]) || grid$dx_m[1] <= 0 ||
    grid$total_time_s[1] <= 0 || grid$plot_intv_s[1] <= 0) {
  stop("Grid metadata must contain positive dx_m, total_time_s and plot_intv_s values.")
}
metres_label <- function(x) sub("\\.?0+$", "", format(x, trim = TRUE, scientific = FALSE))
public_output_dx_m <- if ("public_output_dx_m" %in% names(grid) &&
                          is.finite(grid$public_output_dx_m[1])) {
  grid$public_output_dx_m[1]
} else {
  max(10, grid$dx_m[1])
}
public_output_factor <- public_output_dx_m / grid$dx_m[1]
if (public_output_factor < 1 ||
    abs(public_output_factor - round(public_output_factor)) > 1e-8) {
  stop("The requested public-output resolution must be an integer multiple of the model resolution.")
}
public_output_factor <- as.integer(round(public_output_factor))
grid_tag <- paste0(metres_label(grid$dx_m[1]), "m")
public_grid_tag <- paste0(metres_label(public_output_dx_m), "m")
simulation_minutes <- grid$total_time_s[1] / 60
final_sixth_start_s <- (5 / 6) * grid$total_time_s[1]
elapsed_clock <- function(elapsed_s) {
  sprintf("%02.0f:%02.0f", floor(elapsed_s / 60), round(elapsed_s %% 60))
}
simulation_label <- paste0(sprintf("%.0f", simulation_minutes), "-minute")
final_sixth_start_label <- elapsed_clock(final_sixth_start_s)
simulation_end_label <- elapsed_clock(grid$total_time_s[1])
eta_display_limits <- c(-5, 5)
eta_max_display_limits <- c(0, 5)
hsig_display_limits <- c(0, 4)
ygnbu_palette <- RColorBrewer::brewer.pal(9, "YlGnBu")
# Station 1 is the buoy/offshore end; use deep blue offshore and yellow at
# the nearshore end of the transect.
station_colours <- ygnbu_palette[c(9, 8, 6, 4, 2)]
eta_animation_palette <- grDevices::colorRampPalette(ygnbu_palette)(100)

pretty_field_name <- function(x) {
  x <- gsub("_", " ", x)
  x <- tools::toTitleCase(x)
  x <- gsub("\\bFunwave\\b", "FUNWAVE", x)
  x <- gsub("\\bHs\\b", "Hs", x)
  x <- gsub("\\bTp\\b", "Tp", x)
  x <- gsub("\\bMglob\\b", "Mglob", x)
  x <- gsub("\\bNglob\\b", "Nglob", x)
  x <- gsub("\\bDx\\b", "DX", x)
  x <- gsub("\\bDy\\b", "DY", x)
  x <- gsub("\\bWk\\b", "WK", x)
  x
}

transpose_summary <- function(x, digits = 2) {
  x <- as.data.frame(x, check.names = FALSE, stringsAsFactors = FALSE)
  if (nrow(x) != 1) stop("transpose_summary() expects a one-row data frame.")
  data.frame(
    Item = pretty_field_name(names(x)),
    Value = vapply(x[1, ], function(value) {
      if (is.numeric(value)) format(value, digits = digits, trim = TRUE, scientific = FALSE) else as.character(value)
    }, character(1)),
    check.names = FALSE
  )
}

report_table <- function(x, digits = 2) {
  knitr::kable(x, format = "html", digits = digits, escape = TRUE,
               table.attr = 'class="pf-report-table"')
}

# The buoys report UTC. Show the animation using Port Fairy's local civil time
# while retaining the model's elapsed time from that observation.
port_fairy_tz <- "Australia/Melbourne"
buoy_time_utc <- as.POSIXct(
  sub(" UTC$", "", forcing$time_utc[1]),
  format = "%Y-%m-%d %H:%M:%S", tz = "UTC"
)
if (is.na(buoy_time_utc)) stop("Could not parse the buoy observation time.")

# FUNWAVE writes text grids as Nglob rows (model y) by Mglob columns (model x).
# R's image() expects z[x, y], so transpose every model output exactly once.
as_model_xy <- function(z, label = "model field") {
  z <- as.matrix(z)
  if (all(dim(z) == c(grid$Nglob, grid$Mglob))) return(t(z))
  if (all(dim(z) == c(grid$Mglob, grid$Nglob))) return(z)
  stop(sprintf("%s has %.0f x %.0f values; expected %.0f x %.0f (y x x).",
               label, nrow(z), ncol(z), grid$Nglob, grid$Mglob))
}

read_text_model <- function(path, label = basename(path)) {
  as_model_xy(utils::read.table(path, header = FALSE), label)
}

read_field <- function(prefix, directory = results_dir) {
  candidates <- list.files(directory, pattern = paste0("^", prefix, "_[0-9]{5}$"),
                           full.names = TRUE)
  if (!length(candidates)) return(NULL)
  number <- as.integer(sub(paste0("^", prefix, "_"), "", basename(candidates)))
  read_text_model(candidates[which.max(number)], prefix)
}


read_all_fields <- function(prefix, directory = results_dir) {
  candidates <- list.files(directory, pattern = paste0("^", prefix, "_[0-9]{5}$"),
                           full.names = TRUE)
  if (!length(candidates)) return(list())
  number <- as.integer(sub(paste0("^", prefix, "_"), "", basename(candidates)))
  candidates <- candidates[order(number)]
  lapply(candidates, read_text_model, label = prefix)
}

maximum_model_field <- function(fields) {
  if (!length(fields)) return(NULL)
  out <- fields[[1]]
  if (length(fields) > 1) {
    for (k in 2:length(fields)) out <- pmax(out, fields[[k]], na.rm = TRUE)
  }
  out[!is.finite(out)] <- NA_real_
  out
}

read_static_model <- function(name) {
  path <- file.path(results_dir, name)
  if (!file.exists(path)) return(NULL)
  read_text_model(path, name)
}

model_axes <- function(z) {
  list(x = (seq_len(nrow(z)) - 1) * grid$dx_m,
       y = (seq_len(ncol(z)) - 1) * grid$dy_m)
}

plot_model <- function(z, main, col, zlim = range(z, finite = TRUE), ...) {
  a <- model_axes(z)
  image(a$x, a$y, z, asp = 1, col = col, zlim = zlim,
        xlab = "Model x: source edge to coast (m)",
        ylab = "Model y (m)", main = main, ...)
  invisible(a)
}

model_to_rotated_raster <- function(z, label = "model field") {
  z <- as.matrix(z)
  expected <- c(grid$Mglob, grid$Nglob)
  if (!identical(dim(z), expected)) {
    stop(sprintf("%s has %.0f x %.0f values; expected Mglob x Nglob = %.0f x %.0f.",
                 label, nrow(z), ncol(z), expected[1], expected[2]))
  }

  # z is [model x, model y]. Convert it back to terra's native
  # [north-to-south row, west-to-east column] layout using the exact inverse
  # of the transforms in setup_port_fairy_funwave.R.
  gis_rows <- t(z)
  if (source_is_low_x) {
    gis_rows <- gis_rows[nrow(gis_rows):1, , drop = FALSE]
  } else {
    gis_rows <- gis_rows[, ncol(gis_rows):1, drop = FALSE]
  }

  out <- rast(depth_raster)
  values(out) <- as.vector(t(gis_rows))
  out
}

symmetric_raster_limits <- function(r) {
  m <- global(abs(r), "max", na.rm = TRUE)[1, 1]
  if (!is.finite(m) || m == 0) m <- 1e-8
  c(-m, m)
}

write_model_output_rasters <- function(r_om, stem, method = "bilinear",
                                       already_public = FALSE) {
  # `make_public_output_rotated()` first masks native-grid values then
  # aggregates to the requested public resolution before any geographic
  # reprojection. This keeps a 2.5 m historical simulation numerical while
  # publishing compact 10 m maps and GeoTIFFs.
  if (!already_public) r_om <- make_public_output_rotated(r_om)
  writeRaster(r_om, file.path(case_dir, paste0(stem, "_rotated_omerc.tif")),
              overwrite = TRUE)
  r_ll <- project(r_om, "EPSG:4326", method = method)
  writeRaster(r_ll, file.path(case_dir, paste0(stem, "_wgs84.tif")),
              overwrite = TRUE)
  r_ll
}


symmetric_limits <- function(z) {
  m <- max(abs(z), na.rm = TRUE)
  if (!is.finite(m) || m == 0) m <- 1e-8
  c(-m, m)
}

positive_limits <- function(z) {
  m <- max(z, na.rm = TRUE)
  if (!is.finite(m) || m <= 0) m <- 1e-8
  c(0, m)
}

depth_raster <- rast(file.path(
  case_dir, paste0("depth_", grid_tag, "_positive_water_depth.tif")
))
elevation_raster <- rast(file.path(
  case_dir, paste0("elevation_", grid_tag, "_warped.tif")
))
depth_from_file <- read_text_model(file.path(case_dir, "depth.txt"), "depth.txt")
source_is_low_x <- identical(grid$source_edge, "minimum rotated x")

# This mask comes from the warped DEM supplied to FUNWAVE, not from the
# model's dynamic wet/dry field. It deliberately limits displayed water levels
# to DEM cells with positive water depth.
dem_water_model <- is.finite(depth_from_file) & depth_from_file > 0
dem_water_mask_rotated <- ifel(depth_raster > 0, 1, 0)
dem_water_mask_wgs84 <- project(dem_water_mask_rotated, "EPSG:4326", method = "near")

apply_dem_water_mask_model <- function(z) {
  z <- as.matrix(z)
  z[!dem_water_model] <- NA_real_
  z
}

apply_dem_water_mask_rotated <- function(r_om) {
  r_om[is.na(dem_water_mask_rotated) | dem_water_mask_rotated <= 0] <- NA_real_
  r_om
}

# The internal wavemaker emits energy in both directions. The buoy-side side
# is intentionally hidden from coastal forecast products before projection and
# again afterwards, preventing interpolation from bleeding across the paddle.
shoreward_of_paddle_model <- outer(
  (seq_len(grid$Mglob[1]) - 1) * grid$dx_m[1], seq_len(grid$Nglob[1]),
  function(x, y) x >= grid$x_wk_m[1]
)
shoreward_of_paddle_rotated <- model_to_rotated_raster(
  ifelse(shoreward_of_paddle_model, 1, 0), "shoreward-of-wavemaker mask"
)
shoreward_of_paddle_wgs84 <- project(shoreward_of_paddle_rotated, "EPSG:4326", method = "near")

apply_output_visibility_mask_model <- function(z) {
  z <- apply_dem_water_mask_model(z)
  z[!shoreward_of_paddle_model] <- NA_real_
  z
}

apply_output_visibility_mask_rotated <- function(r_om) {
  r_om <- apply_dem_water_mask_rotated(r_om)
  r_om[is.na(shoreward_of_paddle_rotated) | shoreward_of_paddle_rotated <= 0] <- NA_real_
  r_om
}

make_public_output_rotated <- function(r_om) {
  r_om <- apply_output_visibility_mask_rotated(r_om)
  if (public_output_factor <= 1L) return(r_om)

  # A public cell is retained only where all contributing native cells are
  # water and shoreward of the wavemaker. This prevents a 10 m aggregation
  # from leaking displayed values across the coast or source-side mask.
  visibility <- ifel(
    is.na(dem_water_mask_rotated) | is.na(shoreward_of_paddle_rotated) |
      dem_water_mask_rotated <= 0 | shoreward_of_paddle_rotated <= 0,
    0, 1
  )
  public_visibility <- aggregate(
    visibility, fact = public_output_factor, fun = min, na.rm = TRUE
  )
  r_om <- aggregate(r_om, fact = public_output_factor, fun = mean, na.rm = TRUE)
  r_om[is.na(public_visibility) | public_visibility <= 0] <- NA_real_
  r_om
}

# Reconstruct depth.txt on the exact native Omerc raster, then derive maps
# from that raster. This is the same pathway used for FUNWAVE outputs below.
depth_from_file_rotated <- model_to_rotated_raster(depth_from_file, "depth.txt")
depth_file_minus_raster <- depth_from_file_rotated - depth_raster
max_depth_file_delta <- global(abs(depth_file_minus_raster), "max", na.rm = TRUE)[1, 1]

buoy_ll <- vect(matrix(c(forcing$buoy_lon, forcing$buoy_lat), ncol = 2),
                type = "points", crs = "EPSG:4326")
buoy_om <- project(buoy_ll, crs(depth_raster))

# Plot all map products in WGS84, but only after reconstructing them on the
# native rotated grid. The black outline is the rectangular FUNWAVE domain.
outline_from_raster <- function(r) {
  vect(rbind(
    c(xmin(r), ymin(r)), c(xmax(r), ymin(r)),
    c(xmax(r), ymax(r)), c(xmin(r), ymax(r)),
    c(xmin(r), ymin(r))
  ), type = "polygons", crs = crs(r))
}
model_outline_om <- outline_from_raster(depth_raster)
model_outline_ll_xy <- crds(project(model_outline_om, "EPSG:4326"))

# Model x is always written from the buoy-side source edge toward the coast.
# Draw the actual internal-wavemaker line on every map so the reconstructed
# fields can be checked against their physical source location.
source_x_om <- if (source_is_low_x) {
  xmin(depth_raster) + grid$x_wk_m
} else {
  xmax(depth_raster) - grid$x_wk_m
}
source_line_om <- vect(rbind(
  c(source_x_om, ymin(depth_raster)),
  c(source_x_om, ymax(depth_raster))
), type = "lines", crs = crs(depth_raster))
source_line_ll_xy <- crds(project(source_line_om, "EPSG:4326"))

plot_geographic <- function(r_ll, ..., show_legend = FALSE, range = NULL) {
  # terra::plot uses range= to fix its colour scale; zlim= is for image().
  if (is.null(range)) plot(r_ll, ...) else plot(r_ll, ..., range = range)
  lines(model_outline_ll_xy[, 1], model_outline_ll_xy[, 2],
        col = "black", lwd = 2)
  lines(source_line_ll_xy[, 1], source_line_ll_xy[, 2],
        col = "dodgerblue3", lwd = 2, lty = 2)
  points(crds(buoy_ll)[1, 1], crds(buoy_ll)[1, 2],
         pch = 16, col = "red", cex = 1.25)
  if (show_legend) {
    legend("bottomleft",
           legend = c("FUNWAVE grid", "internal wavemaker", "buoy"),
           col = c("black", "dodgerblue3", "red"), lwd = c(2, 2, NA),
           lty = c(1, 2, NA), pch = c(NA, NA, 16),
           bty = "n", cex = 0.75)
  }
}

local_map_bounds <- function(r_ll, padding = 0.04) {
  e <- ext(r_ll)
  x_pad <- (xmax(e) - xmin(e)) * padding
  y_pad <- (ymax(e) - ymin(e)) * padding
  c(west = xmin(e) - x_pad, south = ymin(e) - y_pad,
    east = xmax(e) + x_pad, north = ymax(e) + y_pad)
}

apply_water_masks_wgs84 <- function(r_ll) {
  dem_mask <- dem_water_mask_wgs84
  if (!compareGeom(dem_mask, r_ll, stopOnError = FALSE)) {
    dem_mask <- resample(dem_mask, r_ll, method = "near")
  }
  r_ll[is.na(dem_mask) | dem_mask <= 0] <- NA_real_

  paddle_mask <- shoreward_of_paddle_wgs84
  if (!compareGeom(paddle_mask, r_ll, stopOnError = FALSE)) {
    paddle_mask <- resample(paddle_mask, r_ll, method = "near")
  }
  r_ll[is.na(paddle_mask) | paddle_mask <= 0] <- NA_real_

  if (is.null(mask_wgs84)) return(r_ll)
  wet_mask <- mask_wgs84
  if (!compareGeom(wet_mask, r_ll, stopOnError = FALSE)) {
    wet_mask <- resample(wet_mask, r_ll, method = "near")
  }
  r_ll[is.na(wet_mask) | wet_mask <= 0] <- NA_real_
  r_ll
}

leaflet_model_map <- function(r_ll, palette, limits, legend_title) {
  b <- local_map_bounds(r_ll)
  pal <- colorNumeric(palette = palette, domain = limits, na.color = "transparent")
  m <- leaflet(width = "100%", height = 480,
               options = leafletOptions(minZoom = 14, maxZoom = 18,
                                        maxBoundsViscosity = 1))
  # These public sources require no project key.
  m <- addProviderTiles(m, "OpenStreetMap.Mapnik", group = "OpenStreetMap")
  m <- addProviderTiles(m, "Esri.WorldImagery", group = "Satellite imagery")
  m <- setMaxBounds(m, b[["west"]], b[["south"]], b[["east"]], b[["north"]])
  m <- fitBounds(m, b[["west"]], b[["south"]], b[["east"]], b[["north"]])
  m <- addRasterImage(m, r_ll, colors = pal, opacity = 0.88, project = FALSE,
                      group = "Model output")
  m <- addPolylines(m, lng = model_outline_ll_xy[, 1], lat = model_outline_ll_xy[, 2],
                    color = "black", weight = 2, group = "FUNWAVE grid")
  m <- addPolylines(m, lng = source_line_ll_xy[, 1], lat = source_line_ll_xy[, 2],
                    color = "dodgerblue3", weight = 2, dashArray = "6 6",
                    group = "Internal wavemaker")
  m <- addCircleMarkers(m, lng = crds(buoy_ll)[1, 1], lat = crds(buoy_ll)[1, 2],
                        radius = 5, color = "red", fillOpacity = 1,
                        label = "Port Fairy wave buoy", group = "Buoy")
  if (exists("station_metadata", inherits = TRUE) && nrow(station_metadata)) {
    m <- addCircleMarkers(
      m, lng = station_metadata$lon, lat = station_metadata$lat,
      radius = 4, color = station_colours[seq_len(nrow(station_metadata))],
      fillOpacity = 1, label = station_metadata$station_id,
      group = "Transect stations"
    )
  }
  m <- addLayersControl(
    m,
    baseGroups = c("OpenStreetMap", "Satellite imagery"),
    overlayGroups = c("Model output", "FUNWAVE grid", "Internal wavemaker", "Buoy", "Transect stations"),
    options = layersControlOptions(collapsed = FALSE)
  )
  m <- addLegend(m, position = "bottomright", pal = pal, values = limits,
                 title = legend_title, opacity = 0.88)
  m
}

frame_local_time <- function(elapsed_s) {
  format(buoy_time_utc + elapsed_s, "%d %b %Y %H:%M %Z", tz = port_fairy_tz)
}

add_wave_frame_annotation <- function(r_ll, elapsed_s) {
  e <- ext(r_ll)
  x_span <- xmax(e) - xmin(e)
  y_span <- ymax(e) - ymin(e)
  bearing_to <- forcing$boundary_direction_to_deg_true[1]
  direction_statistic <- forcing$boundary_direction_statistic[1]
  angle <- bearing_to * pi / 180

  # The map is in longitude/latitude. Scale the longitude component by the
  # local latitude before drawing the true-north travel-direction arrow.
  arrow_y <- 0.13 * y_span
  arrow_x <- arrow_y * sin(angle) / cos(mean(c(ymin(e), ymax(e))) * pi / 180)
  arrow_dy <- arrow_y * cos(angle)
  arrow_x0 <- if (arrow_x >= 0) xmin(e) + 0.08 * x_span else xmax(e) - 0.08 * x_span
  arrow_y0 <- if (arrow_dy >= 0) ymin(e) + 0.10 * y_span else ymax(e) - 0.10 * y_span
  arrows(arrow_x0, arrow_y0, arrow_x0 + arrow_x, arrow_y0 + arrow_dy,
         col = "red", lwd = 3, length = 0.10)

  label <- c(
    paste("Local:", frame_local_time(elapsed_s)),
    sprintf("Model +%02.0f:%04.1f", floor(elapsed_s / 60), elapsed_s %% 60),
    sprintf("Buoy Hs %.2f m | Tp %.1f s", forcing$hs_m[1], forcing$tp_s[1]),
    sprintf("%s waves from %.0f° (arrow travels to %.0f°)",
            tools::toTitleCase(direction_statistic),
            forcing$boundary_direction_from_deg_true[1], bearing_to)
  )
  legend("topright", legend = label, bty = "o", bg = rgb(1, 1, 1, 0.82),
         cex = 0.74, text.col = "black")
}

elevation_wgs84 <- project(elevation_raster, "EPSG:4326", method = "bilinear")
depth_file_wgs84 <- project(depth_from_file_rotated, "EPSG:4326", method = "bilinear")
depth_delta_wgs84 <- project(depth_file_minus_raster, "EPSG:4326", method = "bilinear")

eta <- read_field("eta")
mask <- read_field("mask")
mask_wgs84 <- NULL
if (!is.null(mask)) {
  mask_rotated <- model_to_rotated_raster(mask, "mask")
  mask_wgs84 <- project(mask_rotated, "EPSG:4326", method = "near")
}

eta_paths <- list.files(results_dir, pattern = "^eta_[0-9]{5}$",
                        full.names = TRUE)
# FUNWAVE uses sequential output-file numbers in eta_00000, eta_00001, ...;
# they are not elapsed seconds. Convert those numbers using PLOT_INTV so the
# animation labels and the final-one-sixth selection use model time.
eta_file_number <- as.integer(sub("^eta_", "", basename(eta_paths)))
eta_order <- order(eta_file_number)
eta_paths <- eta_paths[eta_order]
eta_file_number <- eta_file_number[eta_order]
eta_time <- eta_file_number * grid$plot_intv_s
if (length(eta_paths) &&
    (any(!is.finite(eta_time)) || any(eta_time < 0) ||
     max(eta_time) > grid$total_time_s[1] + grid$plot_intv_s[1] ||
     max(eta_time) < grid$total_time_s[1] - grid$plot_intv_s[1])) {
  stop("eta file numbers do not map to the configured simulation period.")
}
read_public_output_raster <- function(path, label = basename(path)) {
  z <- read_text_model(path, basename(path))
  if (!is.null(mask)) z[mask <= 0] <- NA_real_
  make_public_output_rotated(model_to_rotated_raster(z, label))
}
maximum_public_raster <- function(fields) {
  if (!length(fields)) return(NULL)
  out <- fields[[1]]
  if (length(fields) > 1) {
    for (k in 2:length(fields)) {
      out <- terra::app(c(out, fields[[k]]), fun = max, na.rm = TRUE)
    }
  }
  out
}

# Retain report fields only at the requested public resolution. This is
# essential for 2.5 m historical cases: FUNWAVE still writes native 2.5 m
# files, but the report never holds all of those large matrices in memory.
eta_public_frames <- lapply(eta_paths, read_public_output_raster)
eta_max <- maximum_public_raster(eta_public_frames)
if (!is.null(eta_max)) {
  eta_max[!is.finite(eta_max)] <- NA_real_
}

# FUNWAVE's WaveHeight=T diagnostic is written by PREVIEW_MEAN as Hsig files.
# It uses the model time step internally, so it is the correct way to assess
# wave energy rather than comparing individual 15-second eta snapshots.
hsig_paths <- list.files(results_dir, pattern = "^Hsig_[0-9]{5}$",
                         full.names = TRUE)
hsig_number <- as.integer(sub("^Hsig_", "", basename(hsig_paths)))
hsig_paths <- hsig_paths[order(hsig_number)]
hsig_public_frames <- lapply(hsig_paths, read_public_output_raster)
hsig_max <- maximum_public_raster(hsig_public_frames)
if (!is.null(hsig_max)) {
  hsig_max[!is.finite(hsig_max)] <- NA_real_
}

# `eta_max` and `hsig_max` above are already rotated public-resolution rasters,
# not model-index matrices. The original raw field files remain untouched.
eta_frames <- NULL
hsig_frames <- NULL

# Read the five FUNWAVE station files.  Each file is time, eta, u, v and the
# position metadata was created with stations.txt by the case setup script.
station_metadata_path <- file.path(case_dir, "station_metadata.csv")
station_metadata <- if (file.exists(station_metadata_path)) {
  read.csv(station_metadata_path, check.names = FALSE)
} else {
  data.frame()
}
station_timeseries <- data.frame()
if (nrow(station_metadata)) {
  station_series <- lapply(seq_len(nrow(station_metadata)), function(k) {
    path <- file.path(results_dir, sprintf("sta_%04d", k))
    if (!file.exists(path)) return(NULL)
    z <- utils::read.table(path, header = FALSE)
    if (ncol(z) < 2) return(NULL)
    names(z)[seq_len(min(4, ncol(z)))] <- c("time_s", "eta_m", "u_m_s", "v_m_s")[seq_len(min(4, ncol(z)))]
    z$station_id <- station_metadata$station_id[k]
    z
  })
  station_series <- Filter(Negate(is.null), station_series)
  if (length(station_series)) station_timeseries <- do.call(rbind, station_series)
}
if (nrow(station_timeseries)) {
  station_plot <- plotly::plot_ly()
  for (k in seq_len(nrow(station_metadata))) {
    station_name <- station_metadata$station_id[k]
    z <- station_timeseries[station_timeseries$station_id == station_name, , drop = FALSE]
    if (nrow(z)) {
      station_plot <- plotly::add_lines(
        station_plot, data = z, x = ~time_s, y = ~eta_m, name = station_name,
        line = list(color = station_colours[k], width = 2),
        hovertemplate = paste0(station_name, "<br>t = %{x:.1f} s<br>eta = %{y:.3f} m<extra></extra>")
      )
    }
  }
  station_plot <- plotly::layout(
    station_plot, title = "FUNWAVE eta along buoy-to-nearshore transect",
    xaxis = list(title = "Model elapsed time (s)"),
    yaxis = list(title = "Free-surface elevation, eta (m)", range = eta_display_limits),
    legend = list(orientation = "h")
  )
  site_assets_dir <- file.path(report_site_dir, "assets")
  dir.create(site_assets_dir, recursive = TRUE, showWarnings = FALSE)
  htmlwidgets::saveWidget(
    station_plot, file.path(site_assets_dir, "port-fairy-transect-timeseries.html"),
    selfcontained = FALSE
  )
}

# Current diagnostics use the final 60 seconds of saved u/v output. If the
# phase-resolved fields largely cancel in that mean, use the final saved field
# and label it honestly as a snapshot. Option 2 is a grid-advection particle
# map; option 1 remains as a separately inspectable streamline diagnostic.
current_particle_map_url <- NULL
current_streamline_debug_url <- NULL
u_paths <- list.files(results_dir, pattern = "^[uU]_[0-9]{5}$", full.names = TRUE)
v_paths <- list.files(results_dir, pattern = "^[vV]_[0-9]{5}$", full.names = TRUE)
current_file_number <- function(paths) {
  as.integer(sub("^[[:alpha:]]_", "", basename(paths)))
}
u_numbers <- current_file_number(u_paths)
v_numbers <- current_file_number(v_paths)
common_current_numbers <- sort(intersect(u_numbers, v_numbers))
current_numbers <- integer()
if (length(common_current_numbers)) {
  current_times <- common_current_numbers * grid$plot_intv_s[1]
  current_numbers <- tail(common_current_numbers[current_times >= max(current_times) - 60], 5)
}
current_label <- "No paired U/V output was produced"
current_time_label <- "n/a"
current_peak_speed <- NA_real_
current_speed_floor <- NA_real_
current_seed_count <- 0L
current_streamline_count <- 0L
current_public_cell_count <- 0L
current_public_grid_shape <- "n/a"
current_selected_files <- "none"

model_xy_to_wgs84 <- function(xy_model) {
  xy_om <- if (source_is_low_x) {
    cbind(xmin(depth_raster) + xy_model[, 1], ymin(depth_raster) + xy_model[, 2])
  } else {
    cbind(xmax(depth_raster) - xy_model[, 1], ymax(depth_raster) - xy_model[, 2])
  }
  crds(project(vect(xy_om, type = "points", crs = crs(depth_raster)), "EPSG:4326"))
}

trace_current_streamline <- function(seed_xy, u_flow, v_flow, speed_floor,
                                     step_m = 25, n_steps = 100) {
  advance <- function(direction) {
    xy <- seed_xy
    out <- matrix(xy, ncol = 2)
    for (step in seq_len(n_steps)) {
      ij <- pmax(c(1, 1), pmin(c(nrow(u_flow), ncol(u_flow)),
                               round(xy / grid$dx_m[1]) + 1L))
      ux <- u_flow[ij[1], ij[2]]
      vy <- v_flow[ij[1], ij[2]]
      speed <- sqrt(ux^2 + vy^2)
      if (!is.finite(speed) || speed < speed_floor) break
      xy <- xy + direction * step_m * c(ux, vy) / speed
      if (xy[1] < grid$x_wk_m[1] || xy[1] > (nrow(u_flow) - 1) * grid$dx_m[1] ||
          xy[2] < 0 || xy[2] > (ncol(u_flow) - 1) * grid$dy_m[1]) break
      out <- rbind(out, xy)
    }
    out
  }
  backward <- advance(-1)
  forward <- advance(1)
  rbind(backward[nrow(backward):1, , drop = FALSE], forward[-1, , drop = FALSE])
}

if (length(current_numbers)) {
  read_current_component <- function(paths, numbers) {
    path_numbers <- current_file_number(paths)
    fields <- lapply(numbers, function(number) {
      path <- paths[match(number, path_numbers)]
      apply_output_visibility_mask_model(read_text_model(path, basename(path)))
    })
    total <- matrix(0, nrow = nrow(fields[[1]]), ncol = ncol(fields[[1]]))
    count <- matrix(0, nrow = nrow(fields[[1]]), ncol = ncol(fields[[1]]))
    for (field in fields) {
      keep <- is.finite(field)
      total[keep] <- total[keep] + field[keep]
      count[keep] <- count[keep] + 1
    }
    total[count == 0] <- NA_real_
    total[count > 0] <- total[count > 0] / count[count > 0]
    total
  }

  current_selected_files <- paste(sprintf("%05d", current_numbers), collapse = ", ")
  u_flow <- read_current_component(u_paths, current_numbers)
  v_flow <- read_current_component(v_paths, current_numbers)
  current_speed <- sqrt(u_flow^2 + v_flow^2)
  current_label <- "Mean depth-averaged velocity: final 60 s"
  current_time_label <- paste(elapsed_clock(min(current_numbers * grid$plot_intv_s[1])),
                              "to", elapsed_clock(max(current_numbers * grid$plot_intv_s[1])))
  valid_speed <- current_speed[is.finite(current_speed) & current_speed > 0]
  mean_peak_speed <- if (length(valid_speed)) max(valid_speed) else 0
  speed_floor <- if (length(valid_speed)) {
    max(0.005, 0.02 * as.numeric(stats::quantile(valid_speed, 0.15, na.rm = TRUE)))
  } else {
    Inf
  }
  if (!is.finite(mean_peak_speed) || mean_peak_speed < 0.005) {
    last_number <- tail(current_numbers, 1)
    u_flow <- read_current_component(u_paths, last_number)
    v_flow <- read_current_component(v_paths, last_number)
    current_speed <- sqrt(u_flow^2 + v_flow^2)
    valid_speed <- current_speed[is.finite(current_speed) & current_speed > 0]
    speed_floor <- if (length(valid_speed)) {
      max(0.005, 0.02 * as.numeric(stats::quantile(valid_speed, 0.15, na.rm = TRUE)))
    } else {
      Inf
    }
    current_label <- "Depth-averaged velocity: final saved snapshot"
    current_time_label <- elapsed_clock(last_number * grid$plot_intv_s[1])
    current_selected_files <- sprintf("%05d", last_number)
  }
  current_peak_speed <- if (length(valid_speed)) max(valid_speed) else NA_real_
  current_speed_floor <- speed_floor

  x_margin <- max(2 * grid$dx_m[1], 25)
  y_margin <- max(2 * grid$dy_m[1], 25)
  x_start <- grid$x_wk_m[1] + x_margin
  x_end <- (nrow(u_flow) - 1) * grid$dx_m[1] - x_margin
  y_start <- y_margin
  y_end <- (ncol(u_flow) - 1) * grid$dy_m[1] - y_margin
  seeds <- if (x_start < x_end && y_start < y_end && is.finite(speed_floor)) {
    as.matrix(expand.grid(
      seq(x_start, x_end, length.out = 15),
      seq(y_start, y_end, length.out = 11)
    ))
  } else {
    matrix(numeric(), ncol = 2)
  }
  current_seed_count <- nrow(seeds)
  streamlines_model <- lapply(seq_len(nrow(seeds)), function(k) {
    ij <- pmax(c(1, 1), pmin(c(nrow(u_flow), ncol(u_flow)),
                             round(seeds[k, ] / grid$dx_m[1]) + 1L))
    if (!is.finite(current_speed[ij[1], ij[2]]) || current_speed[ij[1], ij[2]] < speed_floor) return(NULL)
    trace_current_streamline(seeds[k, ], u_flow, v_flow, speed_floor)
  })
  streamlines_model <- Filter(function(z) !is.null(z) && nrow(z) >= 4, streamlines_model)
  streamlines_ll <- lapply(streamlines_model, model_xy_to_wgs84)
  streamlines_ll <- Filter(function(z) {
    is.matrix(z) && nrow(z) >= 4 && all(is.finite(z))
  }, streamlines_ll)
  current_streamline_count <- length(streamlines_ll)

  site_assets_dir <- file.path(report_site_dir, "assets")
  dir.create(site_assets_dir, recursive = TRUE, showWarnings = FALSE)
  west <- min(model_outline_ll_xy[, 1]); east <- max(model_outline_ll_xy[, 1])
  south <- min(model_outline_ll_xy[, 2]); north <- max(model_outline_ll_xy[, 2])
  x_pad <- (east - west) * 0.04; y_pad <- (north - south) * 0.04
  current_bounds <- list(c(south - y_pad, west - x_pad), c(north + y_pad, east + x_pad))
  outline_latlng <- lapply(seq_len(nrow(model_outline_ll_xy)), function(k) {
    unname(as.numeric(model_outline_ll_xy[k, c(2, 1)]))
  })
  source_latlng <- lapply(seq_len(nrow(source_line_ll_xy)), function(k) {
    unname(as.numeric(source_line_ll_xy[k, c(2, 1)]))
  })
  buoy_latlng <- unname(as.numeric(c(crds(buoy_ll)[1, 2], crds(buoy_ll)[1, 1])))
  debug_prefix <- paste(
    "Paired U/V files:", length(common_current_numbers),
    "Selected file number(s):", current_selected_files,
    "Field:", current_label,
    "Model time:", current_time_label,
    "Peak speed (m/s):", ifelse(is.finite(current_peak_speed), format(round(current_peak_speed, 3), nsmall = 3), "n/a"),
    "Trace floor (m/s):", ifelse(is.finite(current_speed_floor), format(round(current_speed_floor, 3), nsmall = 3), "n/a"),
    sep = "\n"
  )

  # Option 1: retain a small, independently inspectable streamline map. The
  # debug panel exposes every failure point instead of silently drawing no
  # paths. It is intentionally kept separate from the grid-particle map.
  option1_data <- list(
    bounds = current_bounds,
    lines = lapply(streamlines_ll, function(z) {
      lapply(seq_len(nrow(z)), function(k) unname(as.numeric(z[k, c(2, 1)])))
    }),
    outline = outline_latlng,
    source = source_latlng,
    buoy = buoy_latlng,
    label = paste(current_label, paste0("Model time: ", current_time_label), sep = "\n"),
    debug = paste(debug_prefix, "Option 1 seeds:", current_seed_count,
                  "Option 1 completed paths:", current_streamline_count,
                  sep = "\n")
  )
  option1_js <- paste0(
    "function(el, x, data) {",
    "var map=this,bounds=L.latLngBounds(data.bounds),lines=data.lines||[];",
    "var osm=L.tileLayer('https://{s}.tile.openstreetmap.org/{z}/{x}/{y}.png',{attribution:'&copy; OpenStreetMap contributors'}).addTo(map);",
    "var satellite=L.tileLayer('https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}',{attribution:'Tiles &copy; Esri'});",
    "var paths=L.layerGroup().addTo(map),particles=L.layerGroup().addTo(map);",
    "var grid=L.polyline(data.outline,{color:'black',weight:2}),paddle=L.polyline(data.source,{color:'dodgerblue',weight:2,dashArray:'6 6'}),buoy=L.circleMarker(data.buoy,{radius:5,color:'red',fillOpacity:1});",
    "grid.addTo(map);paddle.addTo(map);buoy.addTo(map);map.fitBounds(bounds,{padding:[4,4]});map.setMaxBounds(bounds);",
    "L.control.layers({'OpenStreetMap':osm,'Satellite imagery':satellite},{'Option 1 flow paths':paths,'Option 1 moving dots':particles,'FUNWAVE grid':grid,'Internal wavemaker':paddle,'Buoy':buoy},{collapsed:false}).addTo(map);",
    "function panel(position,text){var c=L.control({position:position});c.onAdd=function(){var d=L.DomUtil.create('div');d.style.background='rgba(255,255,255,.92)';d.style.padding='6px';d.style.font='12px Arial';d.style.whiteSpace='pre-line';d.textContent=text;return d;};c.addTo(map);}",
    "panel('bottomleft',data.label);panel('topright',data.debug);",
    "var dots=[];lines.forEach(function(line){if(line&&line.length>1){L.polyline(line,{color:'#2c7fb8',weight:1.2,opacity:.72}).addTo(paths);dots.push({line:line,marker:L.circleMarker(line[0],{radius:3,color:'#08306b',weight:0,fillOpacity:.95}).addTo(particles)});}});",
    "if(dots.length){function animate(t){dots.forEach(function(item,i){var q=((t/6500)+(i/dots.length))%1*(item.line.length-1),j=Math.floor(q),a=item.line[j],b=item.line[Math.min(j+1,item.line.length-1)],f=q-j;item.marker.setLatLng([a[0]+f*(b[0]-a[0]),a[1]+f*(b[1]-a[1])]);});window.requestAnimationFrame(animate);}window.requestAnimationFrame(animate);}",
    "window.setTimeout(function(){map.invalidateSize();map.fitBounds(bounds,{padding:[4,4]});},250);",
    "}"
  )
  option1_widget <- leaflet::leaflet(
    width = "100%", height = 560,
    options = leaflet::leafletOptions(minZoom = 14, maxZoom = 18, maxBoundsViscosity = 1)
  )
  option1_widget <- htmlwidgets::onRender(option1_widget, option1_js, data = option1_data)
  htmlwidgets::saveWidget(
    option1_widget,
    file.path(site_assets_dir, "port-fairy-current-streamline-diagnostic.html"),
    selfcontained = FALSE
  )
  current_streamline_debug_url <- "assets/port-fairy-current-streamline-diagnostic.html"

  # Convert model-axis u/v to true east/north components, aggregate to the
  # public grid in the rotated CRS, then project to a regular WGS84 field for
  # browser-side particle advection. No extra R or Docker package is needed.
  x_bearing <- grid$model_x_bearing_deg_true[1] * pi / 180
  y_bearing <- grid$model_y_bearing_deg_true[1] * pi / 180
  u_east_model <- u_flow * sin(x_bearing) + v_flow * sin(y_bearing)
  v_north_model <- u_flow * cos(x_bearing) + v_flow * cos(y_bearing)
  u_east_om <- make_public_output_rotated(model_to_rotated_raster(u_east_model, "eastward current"))
  v_north_om <- make_public_output_rotated(model_to_rotated_raster(v_north_model, "northward current"))
  u_east_ll <- apply_water_masks_wgs84(project(u_east_om, "EPSG:4326", method = "bilinear"))
  v_north_ll <- apply_water_masks_wgs84(project(v_north_om, "EPSG:4326", method = "bilinear"))
  if (!compareGeom(u_east_ll, v_north_ll, stopOnError = FALSE)) {
    v_north_ll <- resample(v_north_ll, u_east_ll, method = "bilinear")
  }
  u_east <- as.matrix(u_east_ll, wide = TRUE)
  v_north <- as.matrix(v_north_ll, wide = TRUE)
  velocity_keep <- is.finite(u_east) & is.finite(v_north) &
    sqrt(u_east^2 + v_north^2) >= current_speed_floor
  u_east[!velocity_keep] <- NA_real_
  v_north[!velocity_keep] <- NA_real_
  current_public_cell_count <- sum(velocity_keep)
  current_public_grid_shape <- paste(nrow(u_east), "x", ncol(u_east))
  particle_count <- if (current_public_cell_count) {
    min(900L, max(250L, as.integer(floor(current_public_cell_count / 75))))
  } else {
    0L
  }
  option2_data <- list(
    bounds = current_bounds,
    outline = outline_latlng,
    source = source_latlng,
    buoy = buoy_latlng,
    label = paste(current_label, paste0("Model time: ", current_time_label),
                  paste0("Public vector grid: ", public_grid_tag),
                  "Particle speed is visually accelerated; directions use the simulated mean field.",
                  sep = "\n"),
    debug = paste(debug_prefix, "Option 2 public grid:", current_public_grid_shape,
                  "Option 2 finite vector cells:", current_public_cell_count,
                  sep = "\n"),
    field = list(
      west = xmin(u_east_ll), south = ymin(u_east_ll),
      east = xmax(u_east_ll), north = ymax(u_east_ll),
      nrows = nrow(u_east), ncols = ncol(u_east),
      uEast = as.vector(t(u_east)), vNorth = as.vector(t(v_north)),
      particleCount = particle_count, visualSpeedMultiplier = 300
    )
  )
  option2_js <- paste0(
    "function(el, x, data) {",
    "var map=this,field=data.field,bounds=L.latLngBounds(data.bounds);",
    "var osm=L.tileLayer('https://{s}.tile.openstreetmap.org/{z}/{x}/{y}.png',{attribution:'&copy; OpenStreetMap contributors'}).addTo(map);",
    "var satellite=L.tileLayer('https://server.arcgisonline.com/ArcGIS/rest/services/World_Imagery/MapServer/tile/{z}/{y}/{x}',{attribution:'Tiles &copy; Esri'});",
    "var grid=L.polyline(data.outline,{color:'black',weight:2}),paddle=L.polyline(data.source,{color:'dodgerblue',weight:2,dashArray:'6 6'}),buoy=L.circleMarker(data.buoy,{radius:5,color:'red',fillOpacity:1});",
    "grid.addTo(map);paddle.addTo(map);buoy.addTo(map);map.fitBounds(bounds,{padding:[4,4]});map.setMaxBounds(bounds);",
    "function number(v){return typeof v==='number'&&Number.isFinite(v);}",
    "function cellOK(row,col){var a=row*field.ncols+col,b=a+1,c=a+field.ncols,d=c+1;return number(field.uEast[a])&&number(field.uEast[b])&&number(field.uEast[c])&&number(field.uEast[d])&&number(field.vNorth[a])&&number(field.vNorth[b])&&number(field.vNorth[c])&&number(field.vNorth[d]);}",
    "function sample(lat,lng){var gx=(lng-field.west)/(field.east-field.west)*(field.ncols-1),gy=(field.north-lat)/(field.north-field.south)*(field.nrows-1);if(!number(gx)||!number(gy)||gx<0||gy<0||gx>=field.ncols-1||gy>=field.nrows-1)return null;var col=Math.floor(gx),row=Math.floor(gy),fx=gx-col,fy=gy-row;if(!cellOK(row,col))return null;var a=row*field.ncols+col,b=a+1,c=a+field.ncols,d=c+1;var interp=function(values){return values[a]*(1-fx)*(1-fy)+values[b]*fx*(1-fy)+values[c]*(1-fx)*fy+values[d]*fx*fy;};return {east:interp(field.uEast),north:interp(field.vNorth)};}",
    "var validCells=[];for(var row=0;row<field.nrows-1;row++){for(var col=0;col<field.ncols-1;col++){if(cellOK(row,col))validCells.push([row,col]);}}",
    "function resetParticle(p){if(!validCells.length)return false;var q=validCells[Math.floor(Math.random()*validCells.length)];p.lat=field.north-(q[0]+Math.random())/(field.nrows-1)*(field.north-field.south);p.lng=field.west+(q[1]+Math.random())/(field.ncols-1)*(field.east-field.west);p.trail=[[p.lat,p.lng]];return true;}",
    "function moveParticle(p,dt){var velocity=sample(p.lat,p.lng);if(!velocity)return false;var seconds=dt*field.visualSpeedMultiplier,nextLat=p.lat+velocity.north*seconds/110540,nextLng=p.lng+velocity.east*seconds/(111320*Math.cos(p.lat*Math.PI/180));if(!sample(nextLat,nextLng))return false;p.lat=nextLat;p.lng=nextLng;p.trail.push([p.lat,p.lng]);if(p.trail.length>7)p.trail.shift();return true;}",
    "var ParticleFlow=L.Layer.extend({initialize:function(){this._particles=[];this._running=false;},onAdd:function(m){this._map=m;this._canvas=L.DomUtil.create('canvas','leaflet-zoom-animated',m.getPanes().overlayPane);this._canvas.style.pointerEvents='none';this._canvas.style.zIndex=450;this._resize();this._particles=[];for(var i=0;i<field.particleCount;i++){var p={trail:[]};if(resetParticle(p))this._particles.push(p);}m.on('resize moveend zoomend',this._resize,this);this._running=true;this._last=0;this._frame=window.requestAnimationFrame(this._animate.bind(this));},onRemove:function(m){this._running=false;window.cancelAnimationFrame(this._frame);m.off('resize moveend zoomend',this._resize,this);L.DomUtil.remove(this._canvas);this._map=null;},_resize:function(){if(!this._map||!this._canvas)return;var size=this._map.getSize(),dpr=window.devicePixelRatio||1;this._canvas.width=Math.max(1,Math.round(size.x*dpr));this._canvas.height=Math.max(1,Math.round(size.y*dpr));this._canvas.style.width=size.x+'px';this._canvas.style.height=size.y+'px';this._topLeft=this._map.containerPointToLayerPoint([0,0]);L.DomUtil.setPosition(this._canvas,this._topLeft);this._ctx=this._canvas.getContext('2d');this._ctx.setTransform(dpr,0,0,dpr,0,0);},_animate:function(timestamp){if(!this._running)return;var dt=this._last?Math.min(0.05,(timestamp-this._last)/1000):0.016;this._last=timestamp;var ctx=this._ctx,size=this._map.getSize();ctx.clearRect(0,0,size.x,size.y);ctx.strokeStyle='rgba(8,48,107,0.68)';ctx.lineWidth=1.15;ctx.lineCap='round';for(var i=0;i<this._particles.length;i++){var p=this._particles[i];if(!moveParticle(p,dt)){resetParticle(p);continue;}if(p.trail.length<2)continue;ctx.beginPath();for(var j=0;j<p.trail.length;j++){var point=this._map.latLngToLayerPoint(p.trail[j]),px=point.x-this._topLeft.x,py=point.y-this._topLeft.y;if(j===0)ctx.moveTo(px,py);else ctx.lineTo(px,py);}ctx.stroke();}this._frame=window.requestAnimationFrame(this._animate.bind(this));}});",
    "var particleFlow=new ParticleFlow();particleFlow.addTo(map);",
    "L.control.layers({'OpenStreetMap':osm,'Satellite imagery':satellite},{'Option 2 particle current field':particleFlow,'FUNWAVE grid':grid,'Internal wavemaker':paddle,'Buoy':buoy},{collapsed:false}).addTo(map);",
    "function panel(position,text){var c=L.control({position:position});c.onAdd=function(){var d=L.DomUtil.create('div');d.style.background='rgba(255,255,255,.92)';d.style.padding='6px';d.style.font='12px Arial';d.style.whiteSpace='pre-line';d.textContent=text;return d;};c.addTo(map);}",
    "panel('bottomleft',data.label);panel('topright',data.debug);",
    "window.setTimeout(function(){map.invalidateSize();map.fitBounds(bounds,{padding:[4,4]});},250);",
    "}"
  )
  option2_widget <- leaflet::leaflet(
    width = "100%", height = 620,
    options = leaflet::leafletOptions(minZoom = 14, maxZoom = 18, maxBoundsViscosity = 1)
  )
  option2_widget <- htmlwidgets::onRender(option2_widget, option2_js, data = option2_data)
  htmlwidgets::saveWidget(
    option2_widget,
    file.path(site_assets_dir, "port-fairy-current-particles.html"),
    selfcontained = FALSE
  )
  current_particle_map_url <- "assets/port-fairy-current-particles.html"
}

current_debug <- data.frame(
  Diagnostic = c("U field files found", "V field files found", "Paired U/V field files",
                 "Selected file number(s)", "Current field used", "Model time represented",
                 "Peak vector speed (m/s)", "Trace / particle floor (m/s)",
                 "Option 1 seeds", "Option 1 completed paths",
                 paste0("Option 2 public vector grid (", public_grid_tag, ")"),
                 "Option 2 finite vector cells"),
  Value = c(length(u_paths), length(v_paths), length(common_current_numbers),
            current_selected_files, current_label, current_time_label,
            ifelse(is.finite(current_peak_speed), format(round(current_peak_speed, 3), nsmall = 3), "n/a"),
            ifelse(is.finite(current_speed_floor), format(round(current_speed_floor, 3), nsmall = 3), "n/a"),
            current_seed_count, current_streamline_count,
            current_public_grid_shape, current_public_cell_count),
  check.names = FALSE, stringsAsFactors = FALSE
)

# A single fixed range makes every eta frame and map comparable across days.
eta_animation_limits <- eta_display_limits

# Both animations are conventional six-second looping GIFs. Their frame
# intervals adapt to the number of saved outputs.
gif_duration_s <- 6
# eta_00000 is the zero-time initial condition, not an evolved model state.
# It is retained in eta_max but deliberately excluded from the full GIF.
full_animation_indices <- which(eta_time > 0)
full_gif_interval <- gif_duration_s / max(1, length(full_animation_indices))
final_sixth_indices <- which(eta_time >= final_sixth_start_s)
final_sixth_gif_interval <- gif_duration_s / max(1, length(final_sixth_indices))
full_animation_saved_start_label <- if (length(full_animation_indices)) {
  elapsed_clock(min(eta_time[full_animation_indices]))
} else {
  "n/a"
}
full_animation_saved_end_label <- if (length(full_animation_indices)) {
  elapsed_clock(max(eta_time[full_animation_indices]))
} else {
  "n/a"
}
final_sixth_saved_start_label <- if (length(final_sixth_indices)) {
  elapsed_clock(min(eta_time[final_sixth_indices]))
} else {
  "n/a"
}
final_sixth_saved_end_label <- if (length(final_sixth_indices)) {
  elapsed_clock(max(eta_time[final_sixth_indices]))
} else {
  "n/a"
}

# Write real GIF files before the HTML is assembled. This is more reliable than
# Knitr's animation hook, which can leave an empty figure placeholder when the
# report is moved to the Pages output directory.
site_assets_dir <- file.path(report_site_dir, "assets")
dir.create(site_assets_dir, recursive = TRUE, showWarnings = FALSE)
full_gif_url <- "assets/port-fairy-full.gif"
final_sixth_gif_url <- "assets/port-fairy-final-sixth.gif"

draw_eta_frame <- function(k) {
  r_om <- eta_public_frames[[k]]
  r_ll <- project(r_om, "EPSG:4326", method = "bilinear")
  r_ll <- apply_water_masks_wgs84(r_ll)
  plot_geographic(r_ll,
                  main = sprintf("Free-surface elevation: t = %.1f s (WGS84)", eta_time[k]),
                  col = eta_animation_palette,
                  range = eta_animation_limits, show_legend = TRUE)
  add_wave_frame_annotation(r_ll, eta_time[k])
}

write_eta_gif <- function(indices, file_name, interval) {
  if (!length(indices)) return(NULL)
  gif_path <- file.path(site_assets_dir, file_name)
  gifski::save_gif(
    expr = for (k in indices) draw_eta_frame(k),
    gif_file = gif_path, width = 800, height = 600, res = 100,
    delay = interval, loop = TRUE, progress = FALSE
  )
  if (!file.exists(gif_path) || file.info(gif_path)$size < 1000) {
    stop("GIF was not written: ", gif_path)
  }
  gif_path
}

full_gif_path <- write_eta_gif(full_animation_indices, "port-fairy-full.gif",
                               full_gif_interval)
final_sixth_gif_path <- write_eta_gif(final_sixth_indices,
                                      "port-fairy-final-sixth.gif",
                                      final_sixth_gif_interval)
```

```{r report-table-style, results='asis', echo=FALSE}
cat('<style>
.pf-report-table { width: 100%; margin: 0.8rem 0 1.5rem; border-collapse: collapse; font-size: 0.93rem; }
.pf-report-table th { background: #2c3e50; color: #fff; text-align: left; padding: 0.55rem 0.7rem; }
.pf-report-table td { padding: 0.45rem 0.7rem; border-bottom: 1px solid #dfe6e9; vertical-align: top; }
.pf-report-table td:first-child { width: 34%; color: #2c3e50; font-weight: 600; }
.pf-report-table tbody tr:nth-child(even) { background: #f6f8fa; }
</style>')
```

## Interpretation first

This is a `r paste0(metres_label(grid$dx_m[1]), " m")`, `r simulation_label`
grid and forcing diagnostic—not a validated harbour forecast. Each raw
FUNWAVE text field is first rebuilt on the exact rotated Oblique Mercator grid,
DEM-masked to water cells, and masked seaward of the internal wavemaker. The
historical workflow aggregates report maps, animations, GeoTIFFs and displayed
velocity fields to `r paste0(public_grid_tag, " cells")` in that rotated CRS,
then reprojects them to WGS84 for geographic display. The interactive maps are
limited to the local model extent and use a compact display height.

Live report: <https://julianog.github.io/run_surf_wave_forecast/>. Source:
<https://github.com/JulianOG/run_surf_wave_forecast>.

```{r forcing-status}
direction_status <- if (forcing$inward_source_component > 0) {
  "The wave vector has an inward component from the selected source edge."
} else {
  "Warning: the wave vector points out of the selected source edge. Do not interpret this run as a coastal-wave prediction."
}
report_table(transpose_summary(data.frame(
  source_edge = grid$source_edge,
  boundary_direction_statistic = forcing$boundary_direction_statistic,
  requested_model_x_bearing_deg_true = forcing$requested_model_x_bearing_deg_true,
  realised_model_x_bearing_deg_true = forcing$model_x_bearing_deg_true,
  funwave_theta_peak_deg = forcing$funwave_theta_peak_deg,
  inward_source_component = forcing$inward_source_component,
  interpretation = direction_status
), digits = 2))
```

## Latest buoy forcing

The irregular `WK_IRR` source is a TMA/JONSWAP-style parameterisation made from
the newest usable IMOS Spotter observation. It uses the buoy **mean** direction
and matching mean directional spread when present (`SSWMD` and `WMDS`), falling
back to the peak values only when the mean values are unavailable. The buoy
provides integral Hs and period parameters—not a phase-resolved spectrum.

```{r forcing}
report_table(transpose_summary(forcing, digits = 2))
```

## Latest free-surface elevation

```{r eta, echo=FALSE}
if (is.null(eta)) {
  plot.new(); text(0.5, 0.5, "No eta output was produced.")
} else {
  eta_plot <- apply_output_visibility_mask_model(eta)
  if (!is.null(mask)) eta_plot[mask <= 0] <- NA_real_
  eta_rotated <- model_to_rotated_raster(eta_plot, "eta")
  eta_wgs84 <- write_model_output_rasters(eta_rotated, "eta_latest")
  eta_wgs84 <- apply_water_masks_wgs84(eta_wgs84)
  writeRaster(eta_wgs84, file.path(case_dir, "eta_latest_wgs84.tif"), overwrite = TRUE)
  leaflet_model_map(eta_wgs84, eta_animation_palette,
                    eta_display_limits, "Free-surface elevation (m)")
}
```

## Free-surface animations

Both GIFs use a shared colour scale and are written as standalone assets from
this Rmd, after each raw output has been reconstructed on the rotated FUNWAVE
grid and projected to WGS84. The zero-time initial condition is
intentionally omitted from the full animation. Each GIF loops for six seconds:
one represents the complete `r simulation_label` run (saved frames
`r full_animation_saved_start_label`--`r full_animation_saved_end_label`), and
the other shows the final one-sixth (configured `r final_sixth_start_label`--`r simulation_end_label`;
saved frames `r final_sixth_saved_start_label`--`r final_sixth_saved_end_label`).

### Full `r simulation_label` simulation: saved `r full_animation_saved_start_label`--`r full_animation_saved_end_label`

```{r eta-animation-full}
if (is.null(full_gif_path)) {
  plot.new(); text(0.5, 0.5, "No evolved eta output was produced.")
} else {
  htmltools::tags$img(src = full_gif_url, alt = "Full Port Fairy free-surface animation",
                       style = "max-width: 100%; height: auto; display: block;")
}
```

### Final one-sixth: saved `r final_sixth_saved_start_label`--`r final_sixth_saved_end_label`

```{r eta-animation-final-sixth}
if (is.null(final_sixth_gif_path)) {
  plot.new(); text(0.5, 0.5, "No final-one-sixth eta output was produced.")
} else {
  htmltools::tags$img(src = final_sixth_gif_url,
                       alt = "Final one-sixth Port Fairy free-surface animation",
                       style = "max-width: 100%; height: auto; display: block;")
}
```

## Maximum free-surface elevation

This is the maximum positive instantaneous `eta` at each cell. It is useful for
seeing crest reach, but it is **not** a wave-height statistic.

```{r hmax-eta, echo=FALSE}
if (is.null(eta_max)) {
  plot.new(); text(0.5, 0.5, "No eta output was produced.")
} else {
  eta_max_wgs84 <- write_model_output_rasters(eta_max, "eta_max", already_public = TRUE)
  eta_max_wgs84 <- apply_water_masks_wgs84(eta_max_wgs84)
  writeRaster(eta_max_wgs84, file.path(case_dir, "eta_max_wgs84.tif"), overwrite = TRUE)
  leaflet_model_map(eta_max_wgs84, ygnbu_palette,
                    eta_max_display_limits, "Maximum free-surface elevation (m)")
}
```

## Peak simulated significant wave height

`WaveHeight = T` is evaluated by FUNWAVE's mean-wave diagnostic after the
configured spin-up. The map below is the largest `Hsig` value at each cell from
the saved mean-wave outputs, rather than a maximum of individual `eta` frames.

```{r hsig-max, echo=FALSE}
if (is.null(hsig_max)) {
  plot.new(); text(0.5, 0.5, "No Hsig output was produced. Check STEADY_TIME and T_INTV_mean.")
} else {
  hsig_max_wgs84 <- write_model_output_rasters(hsig_max, "hsig_max", already_public = TRUE)
  hsig_max_wgs84 <- apply_water_masks_wgs84(hsig_max_wgs84)
  writeRaster(hsig_max_wgs84, file.path(case_dir, "hsig_max_wgs84.tif"), overwrite = TRUE)
  leaflet_model_map(hsig_max_wgs84, ygnbu_palette,
                    hsig_display_limits, "Peak simulated significant wave height (m)")
}
```

## Data acknowledgement and citation

Buoy data were sourced from the **Integrated Marine Observing System (IMOS)**,
which is enabled by the National Collaborative Research Infrastructure Strategy
(NCRIS). The wave data were collected and quality controlled by the University
of Western Australia. This dataset is an output from the Catching Oz Waves
project, which received investment from the Australian Research Data Commons
(ARDC): <https://doi.org/10.47486/DP748>. ARDC is funded by NCRIS.

Suggested data citation: *Deakin University (year of data downloaded), Wave
buoys Observations -- Australia -- near real-time, downloaded from the IMOS URL
on the date of download.*

## Provenance

```{r provenance}
report_table(transpose_summary(grid, digits = 3))
```

## Buoy-to-nearshore eta time series

These are native FUNWAVE station outputs from five selected integer grid cells:
the buoy end is dark blue and the nearshore end is yellow. They are written
every second from the model state—there is no spatial or temporal interpolation.

```{r station-timeseries, echo=FALSE}
if (!nrow(station_timeseries)) {
  plot.new(); text(0.5, 0.5, "No FUNWAVE station output was produced.")
} else {
  htmltools::tags$iframe(
    src = "assets/port-fairy-transect-timeseries.html", width = "100%", height = "560",
    style = "border: 1px solid #dfe6e9; display: block;"
  )
}
```

## Animated current particles

This earth.nullschool-style map advects particles through the final 60-second
mean of FUNWAVE's depth-averaged velocity field. If that mean is too weak, it
uses the final saved velocity snapshot and labels it accordingly. Raw FUNWAVE
`u` and `v` are components along the rotated model axes, so the report first
rotates them to geographic east/north using the true x/y bearings recorded in
`grid_metadata.csv`, then aggregates and projects the vectors for display.
Particle speed is visually accelerated only; particle direction and spatial
variation come from the simulated vector field. OpenStreetMap and satellite
imagery are switchable without an API key.

```{r current-particles, echo=FALSE}
if (is.null(current_particle_map_url)) {
  plot.new(); text(0.5, 0.5, "No paired u/v output was produced for the current map.")
} else {
  htmltools::tags$iframe(
    src = current_particle_map_url, width = "100%", height = "620",
    style = "border: 1px solid #dfe6e9; display: block;"
  )
}
```

## Streamline diagnostic

Option 1 is retained as a separate, inspectable streamline diagnostic. The
table exposes the U/V files, selected model times, velocity threshold, seed
count, and completed-path count so an empty result can be diagnosed rather
than silently displayed as a blank map.

```{r current-streamline-diagnostic, echo=FALSE}
report_table(current_debug)
if (!is.null(current_streamline_debug_url)) {
  htmltools::tags$details(
    htmltools::tags$summary("Show option 1 streamline diagnostic"),
    htmltools::tags$iframe(
      src = current_streamline_debug_url, width = "100%", height = "560",
      style = "border: 1px solid #dfe6e9; display: block; margin-top: 0.7rem;"
    )
  )
}
```
