---
title: "Port Fairy near-real-time FUNWAVE test"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 11, fig.height = 6)

suppressPackageStartupMessages({
  library(terra)
  library(leaflet)
})

case_dir <- normalizePath("../port_fairy/output", mustWork = TRUE)
results_dir <- file.path(case_dir, "results")
forcing <- read.csv(file.path(case_dir, "latest_buoy_forcing.csv"),
                    check.names = FALSE)
grid <- read.csv(file.path(case_dir, "grid_metadata.csv"), check.names = FALSE)

# The buoys report UTC. Show the animation using Port Fairy's local civil time
# while retaining the model's elapsed time from that observation.
port_fairy_tz <- "Australia/Melbourne"
buoy_time_utc <- as.POSIXct(
  sub(" UTC$", "", forcing$time_utc[1]),
  format = "%Y-%m-%d %H:%M:%S", tz = "UTC"
)
if (is.na(buoy_time_utc)) stop("Could not parse the buoy observation time.")

# FUNWAVE writes text grids as Nglob rows (model y) by Mglob columns (model x).
# R's image() expects z[x, y], so transpose every model output exactly once.
as_model_xy <- function(z, label = "model field") {
  z <- as.matrix(z)
  if (all(dim(z) == c(grid$Nglob, grid$Mglob))) return(t(z))
  if (all(dim(z) == c(grid$Mglob, grid$Nglob))) return(z)
  stop(sprintf("%s has %.0f x %.0f values; expected %.0f x %.0f (y x x).",
               label, nrow(z), ncol(z), grid$Nglob, grid$Mglob))
}

read_text_model <- function(path, label = basename(path)) {
  as_model_xy(utils::read.table(path, header = FALSE), label)
}

read_field <- function(prefix, directory = results_dir) {
  candidates <- list.files(directory, pattern = paste0("^", prefix, "_[0-9]{5}$"),
                           full.names = TRUE)
  if (!length(candidates)) return(NULL)
  number <- as.integer(sub(paste0("^", prefix, "_"), "", basename(candidates)))
  read_text_model(candidates[which.max(number)], prefix)
}


read_all_fields <- function(prefix, directory = results_dir) {
  candidates <- list.files(directory, pattern = paste0("^", prefix, "_[0-9]{5}$"),
                           full.names = TRUE)
  if (!length(candidates)) return(list())
  number <- as.integer(sub(paste0("^", prefix, "_"), "", basename(candidates)))
  candidates <- candidates[order(number)]
  lapply(candidates, read_text_model, label = prefix)
}

maximum_model_field <- function(fields) {
  if (!length(fields)) return(NULL)
  out <- fields[[1]]
  if (length(fields) > 1) {
    for (k in 2:length(fields)) out <- pmax(out, fields[[k]], na.rm = TRUE)
  }
  out[!is.finite(out)] <- NA_real_
  out
}

read_static_model <- function(name) {
  path <- file.path(results_dir, name)
  if (!file.exists(path)) return(NULL)
  read_text_model(path, name)
}

model_axes <- function(z) {
  list(x = (seq_len(nrow(z)) - 1) * grid$dx_m,
       y = (seq_len(ncol(z)) - 1) * grid$dy_m)
}

plot_model <- function(z, main, col, zlim = range(z, finite = TRUE), ...) {
  a <- model_axes(z)
  image(a$x, a$y, z, asp = 1, col = col, zlim = zlim,
        xlab = "Model x: source edge to coast (m)",
        ylab = "Model y (m)", main = main, ...)
  invisible(a)
}

model_to_rotated_raster <- function(z, label = "model field") {
  z <- as.matrix(z)
  expected <- c(grid$Mglob, grid$Nglob)
  if (!identical(dim(z), expected)) {
    stop(sprintf("%s has %.0f x %.0f values; expected Mglob x Nglob = %.0f x %.0f.",
                 label, nrow(z), ncol(z), expected[1], expected[2]))
  }

  # z is [model x, model y]. Convert it back to terra's native
  # [north-to-south row, west-to-east column] layout using the exact inverse
  # of the transforms in setup_port_fairy_funwave.R.
  gis_rows <- t(z)
  if (source_is_low_x) {
    gis_rows <- gis_rows[nrow(gis_rows):1, , drop = FALSE]
  } else {
    gis_rows <- gis_rows[, ncol(gis_rows):1, drop = FALSE]
  }

  out <- rast(depth_raster)
  values(out) <- as.vector(t(gis_rows))
  out
}

symmetric_raster_limits <- function(r) {
  m <- global(abs(r), "max", na.rm = TRUE)[1, 1]
  if (!is.finite(m) || m == 0) m <- 1e-8
  c(-m, m)
}

write_model_output_rasters <- function(r_om, stem, method = "bilinear") {
  writeRaster(r_om, file.path(case_dir, paste0(stem, "_rotated_omerc.tif")),
              overwrite = TRUE)
  r_ll <- project(r_om, "EPSG:4326", method = method)
  writeRaster(r_ll, file.path(case_dir, paste0(stem, "_wgs84.tif")),
              overwrite = TRUE)
  r_ll
}


symmetric_limits <- function(z) {
  m <- max(abs(z), na.rm = TRUE)
  if (!is.finite(m) || m == 0) m <- 1e-8
  c(-m, m)
}

positive_limits <- function(z) {
  m <- max(z, na.rm = TRUE)
  if (!is.finite(m) || m <= 0) m <- 1e-8
  c(0, m)
}

depth_raster <- rast(file.path(case_dir, "depth_20m_positive_water_depth.tif"))
elevation_raster <- rast(file.path(case_dir, "elevation_20m_warped.tif"))
depth_from_file <- read_text_model(file.path(case_dir, "depth.txt"), "depth.txt")
source_is_low_x <- identical(grid$source_edge, "minimum rotated x")

# Reconstruct depth.txt on the exact native Omerc raster, then derive maps
# from that raster. This is the same pathway used for FUNWAVE outputs below.
depth_from_file_rotated <- model_to_rotated_raster(depth_from_file, "depth.txt")
depth_file_minus_raster <- depth_from_file_rotated - depth_raster
max_depth_file_delta <- global(abs(depth_file_minus_raster), "max", na.rm = TRUE)[1, 1]

buoy_ll <- vect(matrix(c(forcing$buoy_lon, forcing$buoy_lat), ncol = 2),
                type = "points", crs = "EPSG:4326")
buoy_om <- project(buoy_ll, crs(depth_raster))

# Plot all map products in WGS84, but only after reconstructing them on the
# native rotated grid. The black outline is the rectangular FUNWAVE domain.
outline_from_raster <- function(r) {
  vect(rbind(
    c(xmin(r), ymin(r)), c(xmax(r), ymin(r)),
    c(xmax(r), ymax(r)), c(xmin(r), ymax(r)),
    c(xmin(r), ymin(r))
  ), type = "polygons", crs = crs(r))
}
model_outline_om <- outline_from_raster(depth_raster)
model_outline_ll_xy <- crds(project(model_outline_om, "EPSG:4326"))

# Model x is always written from the buoy-side source edge toward the coast.
# Draw the actual internal-wavemaker line on every map so the reconstructed
# fields can be checked against their physical source location.
source_x_om <- if (source_is_low_x) {
  xmin(depth_raster) + grid$x_wk_m
} else {
  xmax(depth_raster) - grid$x_wk_m
}
source_line_om <- vect(rbind(
  c(source_x_om, ymin(depth_raster)),
  c(source_x_om, ymax(depth_raster))
), type = "lines", crs = crs(depth_raster))
source_line_ll_xy <- crds(project(source_line_om, "EPSG:4326"))

plot_geographic <- function(r_ll, ..., show_legend = FALSE, range = NULL) {
  # terra::plot uses range= to fix its colour scale; zlim= is for image().
  if (is.null(range)) plot(r_ll, ...) else plot(r_ll, ..., range = range)
  lines(model_outline_ll_xy[, 1], model_outline_ll_xy[, 2],
        col = "black", lwd = 2)
  lines(source_line_ll_xy[, 1], source_line_ll_xy[, 2],
        col = "dodgerblue3", lwd = 2, lty = 2)
  points(crds(buoy_ll)[1, 1], crds(buoy_ll)[1, 2],
         pch = 16, col = "red", cex = 1.25)
  if (show_legend) {
    legend("bottomleft",
           legend = c("FUNWAVE grid", "internal wavemaker", "buoy"),
           col = c("black", "dodgerblue3", "red"), lwd = c(2, 2, NA),
           lty = c(1, 2, NA), pch = c(NA, NA, 16),
           bty = "n", cex = 0.75)
  }
}

local_map_bounds <- function(r_ll, padding = 0.12) {
  e <- ext(r_ll)
  x_pad <- (xmax(e) - xmin(e)) * padding
  y_pad <- (ymax(e) - ymin(e)) * padding
  c(west = xmin(e) - x_pad, south = ymin(e) - y_pad,
    east = xmax(e) + x_pad, north = ymax(e) + y_pad)
}

apply_wet_mask_wgs84 <- function(r_ll) {
  if (is.null(mask_wgs84)) return(r_ll)
  r_ll[is.na(mask_wgs84) | mask_wgs84 <= 0] <- NA_real_
  r_ll
}

leaflet_model_map <- function(r_ll, palette, limits, legend_title) {
  b <- local_map_bounds(r_ll)
  pal <- colorNumeric(palette = palette, domain = limits, na.color = "transparent")
  m <- leaflet(options = leafletOptions(minZoom = 14, maxZoom = 18,
                                        maxBoundsViscosity = 1))
  # CARTO's public tile endpoint now returns an API-key warning in the report.
  # Use standard OpenStreetMap tiles instead; they require no project key.
  m <- addProviderTiles(m, "OpenStreetMap.Mapnik", group = "OpenStreetMap")
  m <- setMaxBounds(m, b[["west"]], b[["south"]], b[["east"]], b[["north"]])
  m <- fitBounds(m, b[["west"]], b[["south"]], b[["east"]], b[["north"]])
  m <- addRasterImage(m, r_ll, colors = pal, opacity = 0.88, project = FALSE,
                      group = "Model output")
  m <- addPolylines(m, lng = model_outline_ll_xy[, 1], lat = model_outline_ll_xy[, 2],
                    color = "black", weight = 2, group = "FUNWAVE grid")
  m <- addPolylines(m, lng = source_line_ll_xy[, 1], lat = source_line_ll_xy[, 2],
                    color = "dodgerblue3", weight = 2, dashArray = "6 6",
                    group = "Internal wavemaker")
  m <- addCircleMarkers(m, lng = crds(buoy_ll)[1, 1], lat = crds(buoy_ll)[1, 2],
                        radius = 5, color = "red", fillOpacity = 1,
                        label = "Port Fairy wave buoy", group = "Buoy")
  m <- addLayersControl(
    m,
    baseGroups = "OpenStreetMap",
    overlayGroups = c("Model output", "FUNWAVE grid", "Internal wavemaker", "Buoy"),
    options = layersControlOptions(collapsed = FALSE)
  )
  addLegend(m, position = "bottomright", pal = pal, values = limits,
            title = legend_title, opacity = 0.88)
}

frame_local_time <- function(elapsed_s) {
  format(buoy_time_utc + elapsed_s, "%d %b %Y %H:%M %Z", tz = port_fairy_tz)
}

add_wave_frame_annotation <- function(r_ll, elapsed_s) {
  e <- ext(r_ll)
  x_span <- xmax(e) - xmin(e)
  y_span <- ymax(e) - ymin(e)
  bearing_to <- forcing$boundary_direction_to_deg_true[1]
  direction_statistic <- forcing$boundary_direction_statistic[1]
  angle <- bearing_to * pi / 180

  # The map is in longitude/latitude. Scale the longitude component by the
  # local latitude before drawing the true-north travel-direction arrow.
  arrow_y <- 0.13 * y_span
  arrow_x <- arrow_y * sin(angle) / cos(mean(c(ymin(e), ymax(e))) * pi / 180)
  arrow_dy <- arrow_y * cos(angle)
  arrow_x0 <- if (arrow_x >= 0) xmin(e) + 0.08 * x_span else xmax(e) - 0.08 * x_span
  arrow_y0 <- if (arrow_dy >= 0) ymin(e) + 0.10 * y_span else ymax(e) - 0.10 * y_span
  arrows(arrow_x0, arrow_y0, arrow_x0 + arrow_x, arrow_y0 + arrow_dy,
         col = "red", lwd = 3, length = 0.10)

  label <- c(
    paste("Local:", frame_local_time(elapsed_s)),
    sprintf("Model +%02.0f:%04.1f", floor(elapsed_s / 60), elapsed_s %% 60),
    sprintf("Buoy Hs %.2f m | Tp %.1f s", forcing$hs_m[1], forcing$tp_s[1]),
    sprintf("%s waves from %.0f° (arrow travels to %.0f°)",
            tools::toTitleCase(direction_statistic),
            forcing$boundary_direction_from_deg_true[1], bearing_to)
  )
  legend("topright", legend = label, bty = "o", bg = rgb(1, 1, 1, 0.82),
         cex = 0.74, text.col = "black")
}

elevation_wgs84 <- project(elevation_raster, "EPSG:4326", method = "bilinear")
depth_file_wgs84 <- project(depth_from_file_rotated, "EPSG:4326", method = "bilinear")
depth_delta_wgs84 <- project(depth_file_minus_raster, "EPSG:4326", method = "bilinear")

eta <- read_field("eta")
mask <- read_field("mask")
mask_wgs84 <- NULL
if (!is.null(mask)) {
  mask_rotated <- model_to_rotated_raster(mask, "mask")
  mask_wgs84 <- project(mask_rotated, "EPSG:4326", method = "near")
}

eta_paths <- list.files(results_dir, pattern = "^eta_[0-9]{5}$",
                        full.names = TRUE)
# FUNWAVE uses sequential output-file numbers in eta_00000, eta_00001, ...;
# they are not elapsed seconds. Convert those numbers using PLOT_INTV so the
# animation labels and the final-one-sixth selection use model time.
eta_file_number <- as.integer(sub("^eta_", "", basename(eta_paths)))
eta_order <- order(eta_file_number)
eta_paths <- eta_paths[eta_order]
eta_file_number <- eta_file_number[eta_order]
eta_time <- eta_file_number * grid$plot_intv_s
eta_frames <- lapply(eta_paths, function(path) {
  z <- read_text_model(path, basename(path))
  if (!is.null(mask)) z[mask <= 0] <- NA_real_
  z
})
eta_max <- if (length(eta_frames)) {
  do.call(pmax, c(eta_frames, na.rm = TRUE))
} else {
  NULL
}
if (!is.null(eta_max)) eta_max[!is.finite(eta_max)] <- NA_real_

# FUNWAVE's WaveHeight=T diagnostic is written by PREVIEW_MEAN as Hsig files.
# It uses the model time step internally, so it is the correct way to assess
# wave energy rather than comparing individual 7.5-second eta snapshots.
hsig_paths <- list.files(results_dir, pattern = "^Hsig_[0-9]{5}$",
                         full.names = TRUE)
hsig_number <- as.integer(sub("^Hsig_", "", basename(hsig_paths)))
hsig_paths <- hsig_paths[order(hsig_number)]
hsig_frames <- lapply(hsig_paths, function(path) {
  z <- read_text_model(path, basename(path))
  if (!is.null(mask)) z[mask <= 0] <- NA_real_
  z
})
hsig_max <- if (length(hsig_frames)) {
  do.call(pmax, c(hsig_frames, na.rm = TRUE))
} else {
  NULL
}
if (!is.null(hsig_max)) hsig_max[!is.finite(hsig_max)] <- NA_real_
eta_amplitude <- if (length(eta_frames)) max(vapply(eta_frames, function(z) {
  value <- max(abs(z), na.rm = TRUE)
  if (is.finite(value)) value else 0
}, numeric(1))) else 0
if (!is.finite(eta_amplitude) || eta_amplitude == 0) eta_amplitude <- 1e-8
# One common scale is deliberately used for both GIFs and every frame. The
# final static eta map uses it too, so amplitudes are comparable at a glance.
eta_animation_limits <- c(-eta_amplitude, eta_amplitude)

# Both animations are conventional six-second looping GIFs. Their frame
# intervals adapt to the number of saved outputs.
gif_duration_s <- 6
full_gif_interval <- gif_duration_s / max(1, length(eta_frames))
final_sixth_indices <- which(eta_time >= (5 / 6) * grid$total_time_s)
final_sixth_gif_interval <- gif_duration_s / max(1, length(final_sixth_indices))
depth_out <- read_static_model("dep.out")
```

## Interpretation first

This is a 20 m, 30-minute grid and forcing diagnostic—not a validated harbour
forecast. Each raw FUNWAVE text field is first rebuilt on the exact rotated
Oblique Mercator grid, then reprojected to WGS84 for geographic plotting.

```{r forcing-status}
direction_status <- if (forcing$inward_source_component > 0) {
  "The wave vector has an inward component from the selected source edge."
} else {
  "Warning: the wave vector points out of the selected source edge. Do not interpret this run as a coastal-wave prediction."
}
knitr::kable(data.frame(
  source_edge = grid$source_edge,
  boundary_direction_statistic = forcing$boundary_direction_statistic,
  requested_model_x_bearing_deg_true = forcing$requested_model_x_bearing_deg_true,
  realised_model_x_bearing_deg_true = forcing$model_x_bearing_deg_true,
  funwave_theta_peak_deg = forcing$funwave_theta_peak_deg,
  inward_source_component = forcing$inward_source_component,
  interpretation = direction_status
), digits = 2)
```

## Latest buoy forcing

The irregular `WK_IRR` source is a TMA/JONSWAP-style parameterisation made from
the newest usable IMOS Spotter observation. It uses the buoy **mean** direction
and matching mean directional spread when present (`SSWMD` and `WMDS`), falling
back to the peak values only when the mean values are unavailable. The buoy
provides integral Hs and period parameters—not a phase-resolved spectrum.

```{r forcing}
knitr::kable(forcing, digits = 2)
```

## Input bathymetry and depth-file checks

The black outline shows the rectangular FUNWAVE grid. All three maps below are
displayed in WGS84 only after the native rotated grid has been reconstructed.

```{r input-bathy, fig.width=15, fig.height=5}
par(mfrow = c(1, 3), mar = c(4, 4, 3, 4))
plot_geographic(elevation_wgs84, main = "Warped 20 m elevation input (WGS84)",
                col = hcl.colors(45, "Terrain", rev = TRUE), show_legend = TRUE)
plot_geographic(depth_file_wgs84, main = "FUNWAVE depth.txt reconstructed (WGS84)",
                col = hcl.colors(45, "YlGnBu", rev = TRUE))
plot_geographic(depth_delta_wgs84, main = "depth.txt minus warped raster (WGS84)",
                col = hcl.colors(41, "Blue-Red 3"),
                range = symmetric_raster_limits(depth_delta_wgs84))
```

```{r bathy-check-table}
knitr::kable(data.frame(
  check = c("FUNWAVE depth-file dimensions", "maximum |depth.txt - warped raster|", "wet cells in depth.txt"),
  value = c(
    sprintf("%.0f x %.0f; expected Mglob x Nglob = %.0f x %.0f", nrow(depth_from_file),
            ncol(depth_from_file), grid$Mglob, grid$Nglob),
    sprintf("%.8f m", max_depth_file_delta),
    sprintf("%.1f%%", 100 * mean(depth_from_file > 0))
  )
))
```

## Source orientation diagnostic

The dashed line is the internal wavemaker location. The red arrow is the
selected buoy wave-travel direction (mean when available; peak fallback) after
conversion from the buoy's reported **from** direction into the actual rotated
model axes. The grid is rebuilt daily so model +x is aligned to that travel
direction. This avoids losing most of an oblique sea state into a lateral
sponge; it does not artificially amplify the imposed wave height.

```{r forcing-geometry}
theta <- forcing$funwave_theta_peak_deg * pi / 180
a <- plot_model(depth_from_file, "Wavemaker and transformed buoy direction",
                hcl.colors(45, "YlGnBu", rev = TRUE))
abline(v = grid$x_wk_m, lwd = 3, lty = 2)
arrow_length <- min(max(a$x), max(a$y)) * 0.22
x0 <- max(grid$x_wk_m + arrow_length * 0.7, arrow_length)
y0 <- mean(a$y)
arrows(x0, y0, x0 + arrow_length * cos(theta), y0 + arrow_length * sin(theta),
       lwd = 3, col = "red", length = 0.12)
text(x0, y0, sprintf("ThetaPeak = %.1f°", forcing$funwave_theta_peak_deg),
     pos = 3, col = "red")
```

## FUNWAVE depth read-back

`DEPTH_OUT = T` asks FUNWAVE to write the depth it read. The difference panel
is the direct end-to-end check; values near zero confirm that the model did not
transpose or flip the supplied bathymetry.

```{r depth-readback, fig.width=15, fig.height=5}
if (is.null(depth_out)) {
  plot.new(); text(0.5, 0.5, "No dep.out file was found in this run.")
} else {
  depth_out_minus_input <- depth_out - depth_from_file
  par(mfrow = c(1, 3), mar = c(4, 4, 3, 4))
  plot_model(depth_from_file, "depth.txt supplied to FUNWAVE",
             hcl.colors(45, "YlGnBu", rev = TRUE))
  plot_model(depth_out, "dep.out read back from FUNWAVE",
             hcl.colors(45, "YlGnBu", rev = TRUE))
  plot_model(depth_out_minus_input, "dep.out minus depth.txt",
             hcl.colors(41, "Blue-Red 3"), zlim = symmetric_limits(depth_out_minus_input))
}
```

## Latest free-surface elevation

```{r eta, echo=FALSE}
if (is.null(eta)) {
  plot.new(); text(0.5, 0.5, "No eta output was produced.")
} else {
  eta_plot <- eta
  if (!is.null(mask)) eta_plot[mask <= 0] <- NA_real_
  eta_rotated <- model_to_rotated_raster(eta_plot, "eta")
  eta_wgs84 <- write_model_output_rasters(eta_rotated, "eta_latest")
  eta_wgs84 <- apply_wet_mask_wgs84(eta_wgs84)
  writeRaster(eta_wgs84, file.path(case_dir, "eta_latest_wgs84.tif"), overwrite = TRUE)
  leaflet_model_map(eta_wgs84, hcl.colors(40, "Blue-Red 3"),
                    eta_animation_limits, "Free-surface elevation (m)")
}
```

## Free-surface animations

Both GIFs use all available eta snapshots at a shared colour scale. Knitr's
gifski hook creates them directly from this Rmd, after each raw output has been
reconstructed on the rotated FUNWAVE grid and projected to WGS84. Each GIF
loops for six seconds: one shows the complete 30-minute run and the other
shows the final one-sixth (25:00--30:00).

### Full 30-minute simulation

```{r eta-animation-full, fig.show='animate', animation.hook='gifski', interval=full_gif_interval, aniopts='loop', fig.width=8, fig.height=6, dpi=100}
if (!length(eta_frames)) {
  plot.new(); text(0.5, 0.5, "No eta output was produced.")
} else {
  for (k in seq_along(eta_frames)) {
    r_om <- model_to_rotated_raster(eta_frames[[k]], basename(eta_paths[k]))
    r_ll <- project(r_om, "EPSG:4326", method = "bilinear")
    r_ll <- apply_wet_mask_wgs84(r_ll)
    plot_geographic(r_ll,
                    main = sprintf("Free-surface elevation: t = %.1f s (WGS84)", eta_time[k]),
                    col = hcl.colors(40, "Blue-Red 3"),
                    range = eta_animation_limits, show_legend = TRUE)
    add_wave_frame_annotation(r_ll, eta_time[k])
  }
}
```

### Final one-sixth of the simulation

```{r eta-animation-final-sixth, fig.show='animate', animation.hook='gifski', interval=final_sixth_gif_interval, aniopts='loop', fig.width=8, fig.height=6, dpi=100}
if (!length(final_sixth_indices)) {
  plot.new(); text(0.5, 0.5, "No final-one-sixth eta output was produced.")
} else {
  for (k in final_sixth_indices) {
    r_om <- model_to_rotated_raster(eta_frames[[k]], basename(eta_paths[k]))
    r_ll <- project(r_om, "EPSG:4326", method = "bilinear")
    r_ll <- apply_wet_mask_wgs84(r_ll)
    plot_geographic(r_ll,
                    main = sprintf("Free-surface elevation: t = %.1f s (WGS84)", eta_time[k]),
                    col = hcl.colors(40, "Blue-Red 3"),
                    range = eta_animation_limits, show_legend = TRUE)
    add_wave_frame_annotation(r_ll, eta_time[k])
  }
}
```

## Maximum free-surface elevation

This is the maximum positive instantaneous `eta` at each cell. It is useful for
seeing crest reach, but it is **not** a wave-height statistic.

```{r hmax-eta, echo=FALSE}
if (is.null(eta_max)) {
  plot.new(); text(0.5, 0.5, "No eta output was produced.")
} else {
  eta_max_rotated <- model_to_rotated_raster(eta_max, "eta maximum")
  eta_max_wgs84 <- write_model_output_rasters(eta_max_rotated, "eta_max")
  eta_max_wgs84 <- apply_wet_mask_wgs84(eta_max_wgs84)
  writeRaster(eta_max_wgs84, file.path(case_dir, "eta_max_wgs84.tif"), overwrite = TRUE)
  leaflet_model_map(eta_max_wgs84, hcl.colors(40, "YlOrRd", rev = TRUE),
                    positive_limits(eta_max), "Maximum free-surface elevation (m)")
}
```

## Peak simulated significant wave height

`WaveHeight = T` is evaluated by FUNWAVE's mean-wave diagnostic after the
10-minute spin-up. The map below is the largest `Hsig` value at each cell from
the saved mean-wave outputs, rather than a maximum of individual `eta` frames.

```{r hsig-max, echo=FALSE}
if (is.null(hsig_max)) {
  plot.new(); text(0.5, 0.5, "No Hsig output was produced. Check STEADY_TIME and T_INTV_mean.")
} else {
  hsig_max_rotated <- model_to_rotated_raster(hsig_max, "Hsig maximum")
  hsig_max_wgs84 <- write_model_output_rasters(hsig_max_rotated, "hsig_max")
  hsig_max_wgs84 <- apply_wet_mask_wgs84(hsig_max_wgs84)
  writeRaster(hsig_max_wgs84, file.path(case_dir, "hsig_max_wgs84.tif"), overwrite = TRUE)
  leaflet_model_map(hsig_max_wgs84, hcl.colors(40, "YlOrRd", rev = TRUE),
                    positive_limits(hsig_max), "Peak simulated significant wave height (m)")
}
```

## Data acknowledgement and citation

Buoy data were sourced from the **Integrated Marine Observing System (IMOS)**,
which is enabled by the National Collaborative Research Infrastructure Strategy
(NCRIS). The wave data were collected and quality controlled by the University
of Western Australia. This dataset is an output from the Catching Oz Waves
project, which received investment from the Australian Research Data Commons
(ARDC): <https://doi.org/10.47486/DP748>. ARDC is funded by NCRIS.

Suggested data citation: *Deakin University (year of data downloaded), Wave
buoys Observations -- Australia -- near real-time, downloaded from the IMOS URL
on the date of download.*

## Provenance

```{r provenance}
knitr::kable(grid)
```
