---
title: "Port Fairy near-real-time FUNWAVE test"
output:
  html_document:
    toc: true
    toc_float: true
    theme: flatly
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      fig.width = 11, fig.height = 6)

suppressPackageStartupMessages(library(terra))

case_dir <- normalizePath("../port_fairy/output", mustWork = TRUE)
results_dir <- file.path(case_dir, "results")
forcing <- read.csv(file.path(case_dir, "latest_buoy_forcing.csv"),
                    check.names = FALSE)
grid <- read.csv(file.path(case_dir, "grid_metadata.csv"), check.names = FALSE)

# The buoys report UTC. Show the animation using Port Fairy's local civil time
# while retaining the model's elapsed time from that observation.
port_fairy_tz <- "Australia/Melbourne"
buoy_time_utc <- as.POSIXct(
  sub(" UTC$", "", forcing$time_utc[1]),
  format = "%Y-%m-%d %H:%M:%S", tz = "UTC"
)
if (is.na(buoy_time_utc)) stop("Could not parse the buoy observation time.")

# FUNWAVE writes text grids as Nglob rows (model y) by Mglob columns (model x).
# R's image() expects z[x, y], so transpose every model output exactly once.
as_model_xy <- function(z, label = "model field") {
  z <- as.matrix(z)
  if (all(dim(z) == c(grid$Nglob, grid$Mglob))) return(t(z))
  if (all(dim(z) == c(grid$Mglob, grid$Nglob))) return(z)
  stop(sprintf("%s has %d x %d values; expected %d x %d (y x x).",
               label, nrow(z), ncol(z), grid$Nglob, grid$Mglob))
}

read_text_model <- function(path, label = basename(path)) {
  as_model_xy(utils::read.table(path, header = FALSE), label)
}

read_field <- function(prefix, directory = results_dir) {
  candidates <- list.files(directory, pattern = paste0("^", prefix, "_[0-9]{5}$"),
                           full.names = TRUE)
  if (!length(candidates)) return(NULL)
  number <- as.integer(sub(paste0("^", prefix, "_"), "", basename(candidates)))
  read_text_model(candidates[which.max(number)], prefix)
}


read_all_fields <- function(prefix, directory = results_dir) {
  candidates <- list.files(directory, pattern = paste0("^", prefix, "_[0-9]{5}$"),
                           full.names = TRUE)
  if (!length(candidates)) return(list())
  number <- as.integer(sub(paste0("^", prefix, "_"), "", basename(candidates)))
  candidates <- candidates[order(number)]
  lapply(candidates, read_text_model, label = prefix)
}

maximum_model_field <- function(fields) {
  if (!length(fields)) return(NULL)
  out <- fields[[1]]
  if (length(fields) > 1) {
    for (k in 2:length(fields)) out <- pmax(out, fields[[k]], na.rm = TRUE)
  }
  out[!is.finite(out)] <- NA_real_
  out
}

read_static_model <- function(name) {
  path <- file.path(results_dir, name)
  if (!file.exists(path)) return(NULL)
  read_text_model(path, name)
}

model_axes <- function(z) {
  list(x = (seq_len(nrow(z)) - 1) * grid$dx_m,
       y = (seq_len(ncol(z)) - 1) * grid$dy_m)
}

plot_model <- function(z, main, col, zlim = range(z, finite = TRUE), ...) {
  a <- model_axes(z)
  image(a$x, a$y, z, asp = 1, col = col, zlim = zlim,
        xlab = "Model x: source edge to coast (m)",
        ylab = "Model y (m)", main = main, ...)
  invisible(a)
}

model_to_rotated_raster <- function(z, label = "model field") {
  z <- as.matrix(z)
  expected <- c(grid$Mglob, grid$Nglob)
  if (!identical(dim(z), expected)) {
    stop(sprintf("%s has %d x %d values; expected Mglob x Nglob = %d x %d.",
                 label, nrow(z), ncol(z), expected[1], expected[2]))
  }

  # z is [model x, model y]. Convert it back to terra's native
  # [north-to-south row, west-to-east column] layout using the exact inverse
  # of the transforms in setup_port_fairy_funwave.R.
  gis_rows <- t(z)
  if (source_is_low_x) {
    gis_rows <- gis_rows[nrow(gis_rows):1, , drop = FALSE]
  } else {
    gis_rows <- gis_rows[, ncol(gis_rows):1, drop = FALSE]
  }

  out <- rast(depth_raster)
  values(out) <- as.vector(t(gis_rows))
  out
}

symmetric_raster_limits <- function(r) {
  m <- global(abs(r), "max", na.rm = TRUE)[1, 1]
  if (!is.finite(m) || m == 0) m <- 1e-8
  c(-m, m)
}

write_model_output_rasters <- function(r_om, stem, method = "bilinear") {
  writeRaster(r_om, file.path(case_dir, paste0(stem, "_rotated_omerc.tif")),
              overwrite = TRUE)
  r_ll <- project(r_om, "EPSG:4326", method = method)
  writeRaster(r_ll, file.path(case_dir, paste0(stem, "_wgs84.tif")),
              overwrite = TRUE)
  r_ll
}


symmetric_limits <- function(z) {
  m <- max(abs(z), na.rm = TRUE)
  if (!is.finite(m) || m == 0) m <- 1e-8
  c(-m, m)
}

depth_raster <- rast(file.path(case_dir, "depth_20m_positive_water_depth.tif"))
elevation_raster <- rast(file.path(case_dir, "elevation_20m_warped.tif"))
depth_from_file <- read_text_model(file.path(case_dir, "depth.txt"), "depth.txt")
source_is_low_x <- identical(grid$source_edge, "minimum rotated x")

# Reconstruct depth.txt on the exact native Omerc raster, then derive maps
# from that raster. This is the same pathway used for FUNWAVE outputs below.
depth_from_file_rotated <- model_to_rotated_raster(depth_from_file, "depth.txt")
depth_file_minus_raster <- depth_from_file_rotated - depth_raster
max_depth_file_delta <- global(abs(depth_file_minus_raster), "max", na.rm = TRUE)[1, 1]

buoy_ll <- vect(matrix(c(forcing$buoy_lon, forcing$buoy_lat), ncol = 2),
                type = "points", crs = "EPSG:4326")
buoy_om <- project(buoy_ll, crs(depth_raster))

# Plot all map products in WGS84, but only after reconstructing them on the
# native rotated grid. The black outline is the rectangular FUNWAVE domain.
outline_from_raster <- function(r) {
  vect(rbind(
    c(xmin(r), ymin(r)), c(xmax(r), ymin(r)),
    c(xmax(r), ymax(r)), c(xmin(r), ymax(r)),
    c(xmin(r), ymin(r))
  ), type = "polygons", crs = crs(r))
}
model_outline_om <- outline_from_raster(depth_raster)
model_outline_ll_xy <- crds(project(model_outline_om, "EPSG:4326"))

plot_geographic <- function(r_ll, ..., show_legend = FALSE) {
  plot(r_ll, ...)
  lines(model_outline_ll_xy[, 1], model_outline_ll_xy[, 2],
        col = "black", lwd = 2)
  points(crds(buoy_ll)[1, 1], crds(buoy_ll)[1, 2],
         pch = 16, col = "red", cex = 1.25)
  if (show_legend) {
    legend("bottomleft",
           legend = c("FUNWAVE grid", "buoy"),
           col = c("black", "red"), lwd = c(2, NA), pch = c(NA, 16),
           bty = "n", cex = 0.75)
  }
}

frame_local_time <- function(elapsed_s) {
  format(buoy_time_utc + elapsed_s, "%d %b %Y %H:%M %Z", tz = port_fairy_tz)
}

add_wave_frame_annotation <- function(r_ll, elapsed_s) {
  e <- ext(r_ll)
  x_span <- xmax(e) - xmin(e)
  y_span <- ymax(e) - ymin(e)
  bearing_to <- forcing$peak_direction_to_deg_true[1]
  angle <- bearing_to * pi / 180

  # The map is in longitude/latitude. Scale the longitude component by the
  # local latitude before drawing the true-north travel-direction arrow.
  arrow_y <- 0.13 * y_span
  arrow_x <- arrow_y * sin(angle) / cos(mean(c(ymin(e), ymax(e))) * pi / 180)
  arrow_dy <- arrow_y * cos(angle)
  arrow_x0 <- if (arrow_x >= 0) xmin(e) + 0.08 * x_span else xmax(e) - 0.08 * x_span
  arrow_y0 <- if (arrow_dy >= 0) ymin(e) + 0.10 * y_span else ymax(e) - 0.10 * y_span
  arrows(arrow_x0, arrow_y0, arrow_x0 + arrow_x, arrow_y0 + arrow_dy,
         col = "red", lwd = 3, length = 0.10)

  label <- c(
    paste("Local:", frame_local_time(elapsed_s)),
    sprintf("Model +%02d:%02d", elapsed_s %/% 60, elapsed_s %% 60),
    sprintf("Buoy Hs %.2f m | Tp %.1f s", forcing$hs_m[1], forcing$tp_s[1]),
    sprintf("Waves from %.0f° (arrow travels to %.0f°)",
            forcing$peak_direction_from_deg_true[1], bearing_to)
  )
  legend("topright", legend = label, bty = "o", bg = rgb(1, 1, 1, 0.82),
         cex = 0.74, text.col = "black")
}

elevation_wgs84 <- project(elevation_raster, "EPSG:4326", method = "bilinear")
depth_file_wgs84 <- project(depth_from_file_rotated, "EPSG:4326", method = "bilinear")
depth_delta_wgs84 <- project(depth_file_minus_raster, "EPSG:4326", method = "bilinear")

eta <- read_field("eta")
hmax <- read_field("Hmax")
waveheight_max <- maximum_model_field(read_all_fields("WaveHeight"))
mask <- read_field("mask")

eta_paths <- list.files(results_dir, pattern = "^eta_[0-9]{5}$",
                        full.names = TRUE)
eta_time <- as.integer(sub("^eta_", "", basename(eta_paths)))
eta_paths <- eta_paths[order(eta_time)]
eta_time <- sort(eta_time)
eta_frames <- lapply(eta_paths, function(path) {
  z <- read_text_model(path, basename(path))
  if (!is.null(mask)) z[mask <= 0] <- NA_real_
  z
})
eta_amplitude <- if (length(eta_frames)) max(vapply(eta_frames, function(z) {
  value <- max(abs(z), na.rm = TRUE)
  if (is.finite(value)) value else 0
}, numeric(1))) else 0
if (!is.finite(eta_amplitude) || eta_amplitude == 0) eta_amplitude <- 1e-8
# One common scale is deliberately used for both GIFs and every frame. The
# final static eta map uses it too, so amplitudes are comparable at a glance.
eta_animation_limits <- c(-eta_amplitude, eta_amplitude)

# Both animations are conventional six-second looping GIFs. Their frame
# intervals adapt to the number of saved outputs.
gif_duration_s <- 6
full_gif_interval <- gif_duration_s / max(1, length(eta_frames))
final_sixth_indices <- which(eta_time >= (5 / 6) * grid$total_time_s)
final_sixth_gif_interval <- gif_duration_s / max(1, length(final_sixth_indices))
depth_out <- read_static_model("dep.out")
```

## Interpretation first

This is a 20 m, 30-minute grid and forcing diagnostic—not a validated harbour
forecast. Each raw FUNWAVE text field is first rebuilt on the exact rotated
Oblique Mercator grid, then reprojected to WGS84 for geographic plotting.

```{r forcing-status}
direction_status <- if (forcing$inward_source_component > 0) {
  "The wave vector has an inward component from the selected source edge."
} else {
  "Warning: the wave vector points out of the selected source edge. Do not interpret this run as a coastal-wave prediction."
}
knitr::kable(data.frame(
  source_edge = grid$source_edge,
  requested_cross_shore_bearing_deg_true = forcing$requested_cross_shore_bearing_deg_true,
  realised_model_x_bearing_deg_true = forcing$model_x_bearing_deg_true,
  funwave_theta_peak_deg = forcing$funwave_theta_peak_deg,
  inward_source_component = forcing$inward_source_component,
  interpretation = direction_status
), digits = 2)
```

## Latest buoy forcing

The irregular `WK_IRR` source is a TMA/JONSWAP-style parameterisation made from
the newest usable IMOS Spotter observation. The buoy provides integral Hs, peak
period and direction—not a phase-resolved spectrum.

```{r forcing}
knitr::kable(forcing, digits = 2)
```

## Input bathymetry and depth-file checks

The black outline shows the rectangular FUNWAVE grid. All three maps below are
displayed in WGS84 only after the native rotated grid has been reconstructed.

```{r input-bathy, fig.width=15, fig.height=5}
par(mfrow = c(1, 3), mar = c(4, 4, 3, 4))
plot_geographic(elevation_wgs84, main = "Warped 20 m elevation input (WGS84)",
                col = hcl.colors(45, "Terrain", rev = TRUE), show_legend = TRUE)
plot_geographic(depth_file_wgs84, main = "FUNWAVE depth.txt reconstructed (WGS84)",
                col = hcl.colors(45, "YlGnBu", rev = TRUE))
plot_geographic(depth_delta_wgs84, main = "depth.txt minus warped raster (WGS84)",
                col = hcl.colors(41, "Blue-Red 3"),
                zlim = symmetric_raster_limits(depth_delta_wgs84))
```

```{r bathy-check-table}
knitr::kable(data.frame(
  check = c("FUNWAVE depth-file dimensions", "maximum |depth.txt - warped raster|", "wet cells in depth.txt"),
  value = c(
    sprintf("%d x %d; expected Mglob x Nglob = %d x %d", nrow(depth_from_file),
            ncol(depth_from_file), grid$Mglob, grid$Nglob),
    sprintf("%.8f m", max_depth_file_delta),
    sprintf("%.1f%%", 100 * mean(depth_from_file > 0))
  )
))
```

## Source orientation diagnostic

The dashed line is the internal wavemaker location. The red arrow is the
observed peak-wave **travel** direction after conversion from the buoy's
reported **from** direction into the actual rotated model axes. An arrow with a
leftward component leaves the source side and is a forcing-geometry problem,
not a bathymetry result.

```{r forcing-geometry}
theta <- forcing$funwave_theta_peak_deg * pi / 180
a <- plot_model(depth_from_file, "Wavemaker and transformed buoy direction",
                hcl.colors(45, "YlGnBu", rev = TRUE))
abline(v = grid$x_wk_m, lwd = 3, lty = 2)
arrow_length <- min(max(a$x), max(a$y)) * 0.22
x0 <- max(grid$x_wk_m + arrow_length * 0.7, arrow_length)
y0 <- mean(a$y)
arrows(x0, y0, x0 + arrow_length * cos(theta), y0 + arrow_length * sin(theta),
       lwd = 3, col = "red", length = 0.12)
text(x0, y0, sprintf("ThetaPeak = %.1f°", forcing$funwave_theta_peak_deg),
     pos = 3, col = "red")
```

## FUNWAVE depth read-back

`DEPTH_OUT = T` asks FUNWAVE to write the depth it read. The difference panel
is the direct end-to-end check; values near zero confirm that the model did not
transpose or flip the supplied bathymetry.

```{r depth-readback, fig.width=15, fig.height=5}
if (is.null(depth_out)) {
  plot.new(); text(0.5, 0.5, "No dep.out file was found in this run.")
} else {
  depth_out_minus_input <- depth_out - depth_from_file
  par(mfrow = c(1, 3), mar = c(4, 4, 3, 4))
  plot_model(depth_from_file, "depth.txt supplied to FUNWAVE",
             hcl.colors(45, "YlGnBu", rev = TRUE))
  plot_model(depth_out, "dep.out read back from FUNWAVE",
             hcl.colors(45, "YlGnBu", rev = TRUE))
  plot_model(depth_out_minus_input, "dep.out minus depth.txt",
             hcl.colors(41, "Blue-Red 3"), zlim = symmetric_limits(depth_out_minus_input))
}
```

## Latest free-surface elevation

```{r eta, fig.cap='Latest modelled water-surface elevation reconstructed on the rotated grid and mapped in WGS84. Land/dry cells are masked.'}
if (is.null(eta)) {
  plot.new(); text(0.5, 0.5, "No eta output was produced.")
} else {
  eta_plot <- eta
  if (!is.null(mask)) eta_plot[mask <= 0] <- NA_real_
  eta_rotated <- model_to_rotated_raster(eta_plot, "eta")
  eta_wgs84 <- write_model_output_rasters(eta_rotated, "eta_latest")
  plot_geographic(eta_wgs84, main = "Latest water-surface elevation (WGS84)",
                  col = hcl.colors(40, "Blue-Red 3"),
                  zlim = eta_animation_limits, show_legend = TRUE)
}
```

## Free-surface animations

Both GIFs use all available eta snapshots at a shared colour scale. Knitr's
gifski hook creates them directly from this Rmd, after each raw output has been
reconstructed on the rotated FUNWAVE grid and projected to WGS84. Each GIF
loops for six seconds: one shows the complete 30-minute run and the other
shows the final one-sixth (25:00--30:00).

### Full 30-minute simulation

```{r eta-animation-full, fig.show='animate', animation.hook='gifski', interval=full_gif_interval, aniopts='loop', fig.width=8, fig.height=6, dpi=100}
if (!length(eta_frames)) {
  plot.new(); text(0.5, 0.5, "No eta output was produced.")
} else {
  for (k in seq_along(eta_frames)) {
    r_om <- model_to_rotated_raster(eta_frames[[k]], basename(eta_paths[k]))
    r_ll <- project(r_om, "EPSG:4326", method = "bilinear")
    plot_geographic(r_ll,
                    main = sprintf("Free-surface elevation: t = %d s (WGS84)", eta_time[k]),
                    col = hcl.colors(40, "Blue-Red 3"),
                    zlim = eta_animation_limits, show_legend = TRUE)
    add_wave_frame_annotation(r_ll, eta_time[k])
  }
}
```

### Final one-sixth of the simulation

```{r eta-animation-final-sixth, fig.show='animate', animation.hook='gifski', interval=final_sixth_gif_interval, aniopts='loop', fig.width=8, fig.height=6, dpi=100}
if (!length(final_sixth_indices)) {
  plot.new(); text(0.5, 0.5, "No final-one-sixth eta output was produced.")
} else {
  for (k in final_sixth_indices) {
    r_om <- model_to_rotated_raster(eta_frames[[k]], basename(eta_paths[k]))
    r_ll <- project(r_om, "EPSG:4326", method = "bilinear")
    plot_geographic(r_ll,
                    main = sprintf("Free-surface elevation: t = %d s (WGS84)", eta_time[k]),
                    col = hcl.colors(40, "Blue-Red 3"),
                    zlim = eta_animation_limits, show_legend = TRUE)
    add_wave_frame_annotation(r_ll, eta_time[k])
  }
}
```

## Maximum wave height over the 30-minute run

FUNWAVE's Hmax is maximum free-surface elevation, not wave height. This
diagnostic is therefore the cellwise maximum of every WaveHeight output written
during the run.

```{r hmax, fig.cap='Maximum WaveHeight across all saved 30-minute outputs, reconstructed on the rotated grid and mapped in WGS84.'}
if (is.null(waveheight_max)) {
  plot.new(); text(0.5, 0.5, "No WaveHeight output was produced.")
} else {
  if (!is.null(mask)) waveheight_max[mask <= 0] <- NA_real_
  wave_rotated <- model_to_rotated_raster(waveheight_max, "maximum WaveHeight")
  wave_wgs84 <- write_model_output_rasters(wave_rotated, "waveheight_max_30min")
  plot_geographic(wave_wgs84, main = "Maximum WaveHeight over 30 minutes (WGS84)",
                  col = hcl.colors(40, "YlOrRd", rev = TRUE),
                  show_legend = TRUE)
}
```

## Maximum free-surface elevation

```{r hmax-eta, fig.cap='FUNWAVE Hmax: maximum free-surface elevation (not maximum wave height), shown for context.'}
if (!is.null(hmax)) {
  if (!is.null(mask)) hmax[mask <= 0] <- NA_real_
  hmax_rotated <- model_to_rotated_raster(hmax, "Hmax")
  hmax_wgs84 <- write_model_output_rasters(hmax_rotated, "eta_max")
  plot_geographic(hmax_wgs84, main = "Maximum free-surface elevation, Hmax (WGS84)",
                  col = hcl.colors(40, "Blue-Red 3"),
                  zlim = symmetric_limits(hmax), show_legend = TRUE)
}
```

## Provenance

```{r provenance}
knitr::kable(grid)
```
